### 2019-03-05 README.Rmd
First commit

---
  output: github_document
---

```{r , include = FALSE}

```

# gosset

## Overview

The **gosset** package provides an environment for modelling metadata and experimental citizen science data. It has implementations of cross-validation and stepwise variable selection for glm, lm, Bradley-Terry and Plackett-Luce models.

The package is under development, so be patient and have fun! We appreciate that issues to be reported in our issues section on Github.

### 2019-03-05 R/deviance.R

@@ -1,0 +1,377 @@
+# Compute deviance on a validation sample
+#
+# Glue code for calling btmodel()
+# code from psychotree repository
+# https://github.com/cran/psychotree/blob/master/R/bttree.R
+# Strobl et al (2011)
+.btfit <- function(y, x = NULL, start = NULL, weights = NULL, offset = NULL,
+                   cluster = NULL, ..., estfun = FALSE, object = FALSE)
+{
+  if(!(is.null(x) || NCOL(x) == 0L)) warning("x not used")
+  if(!is.null(offset)) warning("offset not used")
+  rval <- btmodel(y, weights = weights, ..., vcov = object)
+  rval <- list(
+    coefficients = rval$coefficients,
+    objfun = -rval$loglik,
+    estfun = if(estfun) estfun.btmodel(rval) else NULL,
+    object = if(object) rval else NULL
+  )
+  return(rval)
+}
+
+# AIC from a Bradley-Terry model
+# code adpeted from PlackettLuce repository
+# Turner et al (2018)
+# https://github.com/hturner/PlackettLuce/blob/master/R/pltree.R
+#' @method AIC bttree
+#' @importFrom stats formula logLik model.frame model.response model.weights
+#' @export
+AIC.bttree <- function(object, newdata = NULL, ...) {
+  if (is.null(newdata)) {
+    return(NextMethod(object, ...))
+  }
+  # create model.frame from newdata
+  response <- as.character(formula(object)[[2]])
+  if (!response %in% colnames(newdata))
+    stop("`newdata` must include response")
+  f <- formula(object)
+  environment(f) <- parent.frame()
+  newdata <- model.frame(f, data = newdata, ...)
+  # predict node for each grouped ranking
+  node <- partykit::predict.modelparty(object,
+                                       newdata = newdata,
+                                       type = "node")
+  # set up to refit models based on newdata
+  cf <- itempar(object)
+  if (is.null(dim(cf))) cf <- t(as.matrix(cf))
+  nodes <- partykit::nodeids(object, terminal = TRUE)
+  dots <- object$info$dots
+  G <- model.response(newdata)
+  w <- model.weights(newdata)
+  if (is.null(w)) w <- rep.int(1, length(G))
+  LL <- df <- numeric(length(nodes))
+  for (i in seq_along(nodes)){
+    # fit model with coef fixed to get logLik
+    # suppress warning due to fixing maxit
+    id <- node == nodes[i]
[…]

### 2019-03-05 R/environmental_indices.R

@@ -1,0 +1,652 @@
+#' Temperature indices
+#'
+#' Compute temperature indices over a timespan 
+#'
+#' @param object a numeric vector of geographic coordinates (lonlat) or
+#' an array with two dimensions containing the temperature data;
+#' 1st dimension contains the day temperature and 2nd dimension the night temperature.
+#' When lonlat is used, the function makes a call to
+#' \code{nasapower::get_power} to fetch and concatenate environmental data from NASA POWER
+#' (\url{https://power.larc.nasa.gov/})
+#' @param day.one a vector of class 'Date' for the starting date to capture the environmental data
+#' @param span an integer or a vector with integers for the duration of the timespan to be captured
+#' @param index optional, a character or a vector with characters for the indices to be computed. 
+#' All indices are given by default
+#' @param ... additional arguments passed to methods
+#' @return A dataframe with selected indices. Options are:
+#' \item{maxDT}{maximun day temperature (degree Celsius)}
+#' \item{minDT}{minimum day temperature (degree Celsius)}
+#' \item{maxNT}{maximun night temperature (degree Celsius)}
+#' \item{minNT}{minimum night temperature (degree Celsius) }
+#' \item{DTR}{diurnal temperature range (mean difference between DT and NT (degree Celsius)) }
+#' \item{SU}{summer days, number of days with maximum temperature > 30 (degree Celsius)}
+#' \item{TR}{tropical nights, number of nights with maximum temperature > 25 (degree Celsius) }
+#' @examples
+#' # Compute indices using temperature data from MODIS MYD11A1
+#' # (http://dx.doi.org/10.5067/MODIS/MYD11A2.006)
+#' # each row in breadwheat corresponds to the same rows in breadwheat_temp,
+#' # first layer is the day temperature and second layer is the night temperature
+#' 
+#' data("breadwheat_temp", package = "gosset")
+#' data("breadwheat", package = "gosset")
+#' 
+#' # compute all temperature indices for the first 30 days in the plots
+#' temperature(breadwheat_temp,
+#'             day.one = breadwheat["planting_date"],
+#'             span = 30)
+#' 
+#' 
[…]

### 2019-03-05 R/get_climmob.R

@@ -1,0 +1,29 @@
+# library(httr)
+# library(rvest)
+# library(jsonlite)
+# 
+# 
+# 
+# list_projects <- function(key = NULL){
+# 
+#   url <- "https://climmob.net/climmob3/api/readProjects?Apikey="
+# 
+#   result <- httr::GET(url = url, query = list(Apikey = key), httr::accept_json())
+# 
+#   result <- httr::content(result, "parsed", encoding = "UTF-8")
+# 
+#   nodes <- rvest::html_node(result, css = "body")
+# 
+#   text <- rvest::html_text(nodes)
+# 
+#   data <- jsonlite::fromJSON(text)
+# 
+#   data
+# 
+# }
+# 
+# x <- list_projects("d39a3c66-5822-4930-a9d4-50e7da041e77")
+# str(x)
+
+
+

### 2019-03-05 R/plot_nodes.R

@@ -1,0 +1,164 @@
+#' Plot nodes from recursive partitioning trees
+#'
+#' Make a ggplot2 chart from model-based recursive partitioning trees with quasi-variance
+#'
+#' @param object an object of class modelparty
+#' @param labels optional, new labels for the axis X in the plot
+#' by default
+#' @param ... additional arguments passed to ggplot2
+#' @return a list of plots with probabilities of winning and 
+#' intervals based on quasi-standard errors
+#' @seealso \code{\link[qvcalc]{qvcalc}}
+#' @examples
+#' library("PlackettLuce")
+#' library("qvcalc")
+#' library("ggplot2")
+#' 
+#' data("breadwheat", package = "gosset")
+#' 
+#' # convert the tricot rankings from breadwheat data
+#' # into a object of class 'rankings'
+#' names(breadwheat)
+#' 
+#' G <- to_rankings(breadwheat,
+#'                  items = c(1:3),
+#'                  rankings = c(7:8),
+#'                  type = "tricot",
+#'                  grouped.rankings = TRUE)
+#' 
+#' # combine grouped rankings with lon lat from wheat dataset
+#' mydata <- cbind(G, breadwheat[c("lon","lat")])
+#' 
+#' # fit a pltree model using lon and lat
+#' mod <- pltree(G ~ ., data = mydata)
+#' 
+#' # plot nodes
+#' plot_nodes(mod)
+#' @importFrom psychotools itempar
+#' @importFrom partykit nodeids
+#' @import qvcalc
+#' @export
+
+plot_nodes <- function(object, labels = NULL, ...){
+
+  # Extract ids from terminal nodes
+  node_id <- partykit::nodeids(object, terminal = TRUE)
+
+  dots <- list(...)
+  
+  if ("font.size" %in% names(dots)) {
+    font.size <- dots[["font.size"]]
+  } else {
+    font.size <- NULL
+  }
+  
+  
+  # get node information
+  nodes <- list()
+  for (i in seq_along(node_id)) {
+    nodes[[i]] <- object[[ node_id[i] ]]$node$info$object
+  }
+  
+  # get number of observers in each node
+  nobs <- list()
+  for (i in seq_along(node_id)) {
+    nobs[[i]] <- as.integer(object[[ node_id[i] ]]$node$info$nobs) 
+  }
+  
+  # get item parameters from model
+  coeffs <- lapply(nodes, psychotools::itempar)
+
[…]

### 2019-03-05 R/pseudoR2.R

@@ -1,0 +1,241 @@
+#' Pseudo R-squared 
+#' 
+#' Regression coefficient to evaluate goodness-of-fit in a given model  
+#' 
+#' @param object a model object for which pseudo R-squared is desired
+#' @param ... additional arguments affecting the R-squared produced
+#' @return A tidy object containing the pseudo R-squared coefficients:
+#' \item{logLik}{log-likelihood from the fitted model}
+#' \item{logLikNull}{log-likelihood from the null model}
+#' \item{MaxLik}{maximum likelihood pseudo R-squared}
+#' \item{CraggUhler}{Cragg and Uhler's pseudo R-squared}
+#' @examples
+#' 
+#' data("airquality")
+#' 
+#' mod <- glm(Temp ~ Wind + Solar.R, 
+#'            data = airquality, 
+#'            family = poisson())
+#' 
+#' pseudoR2(mod)
+#'  
+#' #################################
+#' 
+#' # Compute pseudo R-squared on a validation sample
+#' 
+#' n <- nrow(airquality)
+#' 
+#' s <- sample(1:n, n * 0.7)
+#' 
+#' train <- airquality[s,]
+#' test <- airquality[-s,]
+#' 
+#' mod <- glm(Temp ~ Wind + Solar.R,
+#'            data = train,
+#'            family = poisson())
+#' 
+#' pseudoR2(mod, newdata = test)
+#' 
+#' @export
+pseudoR2 <- function(object, ...){
+  
+  UseMethod("pseudoR2")
+  
+}
+
+#' @export
+pseudoR2.default <- function(object, newdata = NULL){
+  
+  # update the model as a null model 
+  # model without covariates 
+  mod_null <- update(object, ~ 1)
+  
+  # pseudo R squared 
+  if(is.null(newdata)){
+    # get the logLik of the null model
+    LLNull <- deviance(mod_null)[1] / -2
+    
+    # get the logLik of the original model
+    LL <- deviance(object)[1] / -2
+    
+    # number of observations in the original model
+    # to avoid issues with the structure of different model aproaches
+    # we get the number of rows of a prediction matrix which will
+    # have the same dimensions as the response variable 
+    n <- nrow(as.matrix(predict(object)))
+  }
+  
+  # pseudo R squared on a validation sample
+  if(!is.null(newdata)){
[…]

### 2019-03-05 R/to_paircomp.R

@@ -1,0 +1,101 @@
+#' Pairwise comparison rankings
+#'
+#' Pairwise comparisons from a ranking object
+#'
+#' @param object a "rankings" or "grouped_rankings" object which is a matrix of dense rankings. 
+#' The function do not take ties into account, then they are added as NA's
+#' @return an object of class "paircomp" which is a matrix of pairwise comparisons
+#' @seealso \code{\link[PlackettLuce]{rankings}} , \code{\link[psychotools]{paircomp}}
+#' @importFrom psychotools paircomp
+#' @examples 
+#' 
+#' library("psychotools")
+#' 
+#' # a simple matrix with 4 items
+#' # ties are computed as NA's 
+#' R <- matrix(c(1, 2, 0, 0,
+#'               4, 1, 2, 3,
+#'               2, 4, 3, 1,
+#'               1, 2, 3, 0,
+#'               2, 1, 1, 0,
+#'               1, 0, 3, 2), nrow = 6, byrow = TRUE)
+#' colnames(R) <- c("apple", "banana", "orange", "pear")
+#' R <- as.rankings(R)
+#' 
+#' to_paircomp(R)
+#' 
+#' #############################################
+#' 
+#' # using breadwheat data
+#' data("breadwheat", package = "gosset")
+#' 
+#' # convert the tricot rankings from breadwheat data
+#' # into a object of class 'rankings' from PlackettLuce
+#' R <- to_rankings(breadwheat,
+#'                  items = c(1:3),
+#'                  rankings = c(7:8),
+#'                  type = "tricot")
+#' 
+#' 
+#' to_paircomp(R)
+#' 
+#' @export
+to_paircomp <- function(object){
+  
+  R <- object
+  
+  if (!class(R) %in% c("rankings", "grouped_rankings")) {
+    stop("object must be of class 'rankings' or 'grouped_rankings' \n")
+  }
+  
+  # check wich kind of input is given and convert it into a 
+  # matrix with rankings 
+  if (class(R) == "rankings") {
+    R <- R[1:length(R), , as.rankings = FALSE]
+  }
+  if (class(R) == "grouped_rankings") {
+    R <- R[1:length(R), , as.grouped_rankings = FALSE]
+  }
+  
+  # zeros into NA's
+  R[R == 0] <- NA
+  
+  # take name of items
+  items <- dimnames(R)[[2]]
+  
+  # make pairwise comparisons
+  cc <- .combn2(items, 2)
+  
[…]

### 2019-03-05 R/to_rankings.R

@@ -1,0 +1,366 @@
+#' Plackett-Luce rankings
+#'
+#' Create an object of class "rankings" from a dataframe or matrix
+#'
+#' @param data a data frame with columns specified by items and rankings. Data can be in both wide and long format.
+#' An id is required if long format.
+#' @param items a data frame or vector representing the item names
+#' @param rankings a data frame or vector representing the rankings
+#' @param type the type of data input. Options are 'tricot' and 'rank'
+#' @param ... further arguments passed to PlackettLuce::rankings
+#' @return a "rankings" object, which is a matrix of dense rankings 
+#' @seealso \code{\link[PlackettLuce]{rankings}}
+#' @examples
+#' 
+#' library("PlackettLuce")
+#' 
+#' # A matrix with 10 rankings of 5 items (A, B, C, D, E)
+#' # with numeric values as "rank"
+#' set.seed(123)
+#' df <- cbind(id = rep(1:10, each = 5),
+#'             items = rep(LETTERS[1:5], times = 10),
+#'             rankings = runif(50, 1, 3))
+#' df
+#' 
+#' # return a 'rankings' object
+#' to_rankings(df,
+#'             items = 2,
+#'             rankings = 3,
+#'             id = 1)
+#' 
+#' # return a 'grouped_rankings' object
+#' to_rankings(df,
+#'             items = 2,
+#'             rankings = 3,
+#'             id = 1,
+#'             grouped.rankings = TRUE)
+#' 
+#' ##################################
+#' 
+#' # Rankings with 5 items randomly assigned
+#' 
+#' i <- as.data.frame(matrix(NA, nrow = 10, ncol = 5))
+#' names(i) <- paste0("Item",1:5)
+#' 
+#' r <- as.data.frame(matrix(NA, nrow = 10, ncol = 5))
+#' names(r) <- paste0("Position_Item",1:5)
+#' 
+#' for(s in 1:10) {
+#'   i[s,] <- sample(LETTERS[1:5])
+#'   r[s,] <- sample(1:5)
+#' }
+#' 
+#' cbind(i, r) 
+#' 
+#' 
+#' to_rankings(items = i,
+#'             rankings = r)
+#' 
+#' ###################################
+#' 
+#' # breadwheat data, which is an object ordered in the 'tricot' format
+#' # each observer compares 3 varieties randomly distributed from a list of 16 
+#' data("breadwheat", package = "gosset")
+#' 
+#' to_rankings(breadwheat,
+#'             items = c("variety_a","variety_b","variety_c"),
[…]

### 2019-03-05 R/worst_regret.R

@@ -1,0 +1,108 @@
+#' Worst regret in regret-based models from recursive partitioning trees
+#'
+#' Compute a worst regret index from recursive partitioning trees. Regret is an important
+#' heuristic in the behavioural sciences. Minimizing worst regret (the loss under the worst
+#' possible outcome) is a criterion that takes a conservative approach to risk analysis in 
+#' diversification strategies. 
+#'
+#' @param object an object of class modelparty
+#' @param ... further arguments passed to methods
+#' @return A dataframe with probabilities of winning for each item and its worst regret values
+#' @examples
+#' 
+#' library("PlackettLuce")
+#' 
+#' data("breadwheat", package = "gosset")
+#' 
+#' # convert the tricot rankings from breadwheat data
+#' # into a object of class 'rankings'
+#' names(breadwheat)
+#' 
+#' G <- to_rankings(breadwheat,
+#'                  items = c(1:3),
+#'                  rankings = c(7:8),
+#'                  type = "tricot",
+#'                  grouped.rankings = TRUE)
+#' 
+#' 
+#' # compute temperature indices for the first 45 days in the plots
+#' temp <- temperature(breadwheat_temp,
+#'                     day.one = breadwheat["planting_date"],
+#'                     span = 45)
+#' 
+#' 
+#' # combine grouped rankings with temperature indices
+#' mydata <- cbind(G, temp)
+#' 
+#' # fit a pltree model using temperature data
+#' mod <- pltree(G ~ ., data = mydata)
+#' 
+#' 
+#' worst_regret(mod)
+#' 
+#' @export
+worst_regret <- function(object, ...){
+  
+  # get ids of terminal nodes
+  nodes <- partykit::nodeids(object, terminal = TRUE)
+
+  # get the coefficients from each terminal node
+  probs <- list()
+  for(i in seq_along(nodes)) probs[[i]] <- object[[ nodes[i] ]]$node$info$object
+
+  # get estimates from terminal nodes using qvcalc
+  probs <- lapply(probs, function(X) {
+
+    parameters <- psychotools::itempar(X, vcov = TRUE, alias = TRUE)
+    parameters <- qvcalc::qvcalc.itempar(parameters)
+    parameters <- parameters[[2]]
+    parameters <- as.data.frame(as.matrix(parameters))
+  } )
+
+  # get names of items
+  items <- unique(row.names(probs[[1]]))
+
[…]

### 2019-03-26 README.Rmd
to_rankings

@@
 ---
-
-```{r , include = FALSE}
-
-```
 
@@
 
-The package is under development, so be patient and have fun! We appreciate that issues to be reported in our issues section on Github.
+The package is under development, so be patient and have fun! If you find any issue, please report it in our [issues section](https://github.com/kauedesousa/gosset/issues).
+
+Use `devtools::install_github` to install the package
+
+```
+library("devtools")
+install_github("kauedesousa/gosset", upgrade = "never")
+```

### 2019-03-26 R/to_rankings.R

@@ -6,8 +6,8 @@
 #' An id is required if long format.
 #' @param items a data frame or vector representing the item names
 #' @param rankings a data frame or vector representing the rankings
-#' @param type the type of data input. Options are 'tricot' and 'rank'
-#' @param ... further arguments passed to PlackettLuce::rankings
+#' @param type optional, the type of data input, 'rank' is set by default
+#' @param ... other arguments passed to methods
 #' @return a "rankings" object, which is a matrix of dense rankings 
 #' @seealso \code{\link[PlackettLuce]{rankings}}
 #' @examples
@@ -77,21 +77,21 @@
 #' 
 #' # comparison with local item is added as an additional rankings, then
 #' # each of the 3 varieties are compared separately with the local item,
-#' # it return a object four times larger (in rows) than the input data
+#' # it return a object 1 + n_c (number of comparisons) larger (in rows) than the input data
 #' # combining this with covariates from other dataset is easy since 
 #' # the function keeps an internal id
-#' to_rankings(beans,
+#' # argument 'add.rank' must be passed as a dataframe
+#' to_rankings(data = beans,
 #'             items = c(1:3),
 #'             rankings = c(4:5),
-#'             local = c(6:8),
 #'             type = "tricot",
+#'             add.rank = beans[c(6:8)],
 #'             grouped.rankings = TRUE)
 #' 
 #' @import tidyverse
-#' @importFrom reshape2 melt
 #' @export
-
-to_rankings <- function(data = NULL, items = NULL, rankings = NULL, type = NULL, ...) {
+to_rankings <- function(data = NULL, items = NULL,
+                        rankings = NULL, type = NULL, ...) {
   
   if (is.null(data)) {
     data <- cbind(items, rankings)
@@ -142,7 +142,10 @@ to_rankings <- function(data = NULL, items = NULL, rankings = NULL, type = NULL,
     # make sure that rankings are numeric
     data[rankings] <- lapply(data[rankings], as.numeric)
     
-    R <- .fromrank(id, items, r = data[rankings])
+    r <- .pivot_default(id, i = items, r = data[rankings])
+    
+    # make a PlackettLuce ranking
+    R <- PlackettLuce::as.rankings(r)
     
     if ("grouped.rankings" %in% names(dots)) {
       n <- nrow(R)
[…]

### 2019-03-29 R/utility.R
fix bug in breadwheat

@@ -1,4 +1,15 @@
-# check is number is a whole number (no decimals)
+#' Whole number
+#'
+#' Find integer values in a vector with decimals
+#'
+#' @param x vector with numbers of class "numeric" or "integer"
+#' @return a logical vector 
+#' @examples
+#' 
+#' x <- c(1, 2, 1.4, 3, 4.01)
+#' 
+#' is_wholenumber(x)
+#'
 #' @export
 is_wholenumber <- function(x){
   x <- as.vector(t(x))
@@ -6,7 +17,18 @@ is_wholenumber <- function(x){
   return(x)
 }
 
-# check is number is not whole number (has decimals)
+#' Decimal number
+#'
+#' Find a decimal number in a vector with integers
+#'
+#' @param x vector with numbers of class "numeric" or "integer"
+#' @return a logical vector 
+#' @examples
+#' 
+#' x <- c(1, 2, 1.4, 3, 4.01)
+#' 
+#' is_decimal(x)
+#' 
 #' @export
 is_decimal <- function(x){
   x <- as.vector(t(x))
@@ -14,11 +36,33 @@ is_decimal <- function(x){
   return(x)
 }
 
-# group and convert numeric values into integer ranks
+#' Numbers into rankings
+#'
+#' Group and convert numeric values into integer ranks
+#'
+#' @param x vector with numbers of class "numeric" or "integer"
+#' @return a vector with ranked values where highest values are placed first
+#' @examples
+#' 
+#' x <- c(1, 2, 1.4, 3, 4.01)
+#' 
+#' is_wholenumber(x)
+#' 
 # the highest positive value is the best scored item
 # negative values are included as least scored items
 #' @export
-num2rank <- function(object){
+num2rank <- function(object, ...){
+
+  # dots <- list(...)
+  #   
+  # id <- dots[["id"]]
+  # 
+  # if(is.null(id)) {
+  #   id <- rep(1, length(object))
+  # }
+  # 
+  # object$id <- id
+  
   object <- dplyr::mutate(dplyr::group_by(object , id),
                           rank = rank((rank -1)*-1, 
                                       na.last = "keep"))
@@ -27,7 +71,7 @@ num2rank <- function(object){
   
 }
 
-# Testing a grouped_rankings object
+# Test a grouped_rankings object
 #' @export
 is_grouped_rankings <- function(object) {
   
@@ -35,7 +79,7 @@ is_grouped_rankings <- function(object) {
   
 }
 
[…]

### 2019-05-01 README.Rmd
fix bug in breadwheat
checks

--- 2019-03-05 README.Rmd
@@
 
-The package is under development, so be patient and have fun! We appreciate that issues to be reported in our issues section on Github.
+## Installation
+
+The development version can be installed via
+
+    library("devtools")
+    devtools::install_github("kauedesousa/gosset", upgrade = "never")
+
+**gosset** is under development. Use with caution.
+
+## Code of Conduct
+
+Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2019-05-01 R/crossvalidation.R

@@ -1,6 +1,6 @@
 #' Cross-validation 
 #'
-#' Methods for measuring the performance of a predictive model on test data sets. 
+#' Methods for measuring the performance of a predictive model on sets of test data. 
 #'
 #' @param formula an object of class "formula" (or one that can be coerced to that class):
 #' a symbolic description of the model to be fitted,
@@ -15,7 +15,7 @@
 #' 'foldsize', a weighted mean by the size in each fold; 
 #' 'stouffer' a weighted Z-test developed by Stouffer et al. (1949). 
 #' The two last methods are suggested for cross-validation with imbalaced folds
-#' @param ... additional arguments passed to the selected fitting model
+#' @param ... additional arguments passed to methods
 #' @return The cross-validation goodness-of-fit estimates for the best model, which are:
 #' \item{AIC}{Akaike Information Criterion}
 #' \item{deviance}{Model deviance}
@@ -69,17 +69,13 @@
 #' crossvalidation(preference ~ ., 
 #'                 data = Topmodel2007,
 #'                 k = 5)
-
-#############################
-#' 
-#' 
+#'                 
 #' @import partykit
 #' @import psychotools
 #' @import psychotree
 #' @import PlackettLuce
 #' @import gnm
 #' @export
-
 crossvalidation <- function(formula, data, k = NULL,
                             folds = NULL, mean.method = NULL, ...)
 {

### 2019-05-01 R/deviance.R

@@ -46,6 +46,8 @@ AIC.bttree <- function(object, newdata = NULL, ...) {
   # and df of original model fit
   -2 * sum(LL) + 2 * attr(logLik(object), "df")
 }
+
+
 # formula logLik model.frame model.response model.weights 
 # deviance from a Bradley-Terry model
 # code adapted from PlackettLuce repository
@@ -132,6 +134,7 @@ deviance.pltree <- function(object, newdata = NULL, ...) {
   # and the deviance 
   aic - (2 * df)
 }
+
 
 # compute AIC with a validation sample
 # code from https://freakonometrics.hypotheses.org/20158

### 2019-05-01 R/forward.R

@@ -36,47 +36,6 @@
 #'                select.by = "AIC",
 #'                family = poisson(link = "log"))
 #' 
-#' #######
-#' 
-#' \dontrun{
-#' # breadwheat data
-#' library("PlackettLuce")
-#' library("doParallel")
-#' library("abind")
-#' 
-#' data("breadwheat", package = "gosset")
-#' 
-#' # get an object of class 'grouped_rankings' 
-#' # from PlackettLuce
-#' G <- to_rankings(breadwheat, 
-#'                  items = c("variety_a","variety_b","variety_c"), 
-#'                  rankings = c("overall_best","overall_worst"),
-#'                  type = "tricot",
-#'                  grouped.rankings = TRUE)
-#' 
-#' # get some explanatory variables to include in the model
-#' # we will calculate the temperature indices
-#' # for the planting date to maturity (120 days)
-#' 
-#' data("breadwheat_modis", package = "gosset")
-#' 
-#' covars <- temperature(breadwheat_modis,
-#'                       day.one = breadwheat$planting_date,
-#'                       span = 120)
-#' 
-#' # combine the grouped_rankings and the explanatory variables
-#' data <- cbind(G, covars, breadwheat[c("lon","lat")])
-#' 
-#' # run forward selection
-#' mod <- forward(G ~ ., 
-#'                data = data, 
-#'                k = 3, 
-#'                ncores = 3,
-#'                minsize = 50, 
-#'                alpha = 0.01)
-#'
-#' }
-#'             
 #' @import doParallel
 #' @import foreach
 #' @import abind

### 2019-05-01 R/pseudoR2.R

@@ -44,8 +44,6 @@ pseudoR2 <- function(object, ...){
   
 }
 
-# part of this code from 
-# https://github.com/atahk/pscl/blob/master/R/pseudoRSq.R
 #' @export
 pseudoR2.default <- function(object, newdata = NULL){
   

### 2019-05-01 R/utility.R

@@ -11,9 +11,9 @@
 #' is_wholenumber(x)
 #'
 #' @export
-is_wholenumber <- function(x){
+is_wholenumber <- function(x) {
   x <- as.vector(t(x))
-  x <- x%%1==0
+  x <- x %% 1 == 0
   return(x)
 }
 
@@ -30,9 +30,9 @@ is_wholenumber <- function(x){
 #' is_decimal(x)
 #' 
 #' @export
-is_decimal <- function(x){
+is_decimal <- function(x) {
   x <- as.vector(t(x))
-  x <- !x%%1==0
+  x <- !x %% 1 == 0
   return(x)
 }
 

### 2019-05-01 README.Rmd
appveyor_check
pack_site

@@
 
+```{r rmd-setup, include = FALSE}
+knitr::opts_chunk$set(fig.path = "man/figures/")
+```
+
 # gosset
+
+[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/kauedesousa/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
+
+Package website: https://kauedesousa.github.io/gosset/
 
@@
 
+## Going further
+
+The full functionality of **gosset** is illustrated in the package vignette.
+The vignette can be found on the [package website](https://kauedesousa.github.io/gosset/) or from within R once the package has been installed.
+
 ## Code of Conduct

### 2019-05-03 README.Rmd
contributing_guidelines
gosset_belongs_to_agrobioinfoservices

--- 2019-05-01 README.Rmd
@@
 
+```{r rmd-setup, include = FALSE}
+knitr::opts_chunk$set(fig.path = "man/figures/")
+```
+
 # gosset
+
+[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
+
+Package website: https://agrobioinfoservices.github.io/gosset/
 
@@
     library("devtools")
-    devtools::install_github("kauedesousa/gosset", upgrade = "never")
+    devtools::install_github("agrobioinfoservices/gosset", upgrade = "never")
 
 **gosset** is under development. Use with caution.
+
+## Going further
+
+The full functionality of **gosset** is illustrated in the package vignette.
+The vignette can be found on the [package website](https://agrobioinfoservices.github.io/gosset/) or from within R once the package has been installed.
+
+## Contribution
+
+You are welcome to contribute to this project. Please read our [contribution guide lines](CONTRIBUTING.md).
 

### 2019-05-13 README.Rmd
codecov

@@
 
-```{r rmd-setup, include = FALSE}
-knitr::opts_chunk$set(fig.path = "man/figures/")
+```{r setup, include = FALSE}
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  out.width = "100%"
+)
+options(width = 120)
 ```
@@
 
+<!-- badges: start -->
+[![GPLv3 license](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.r-project.org/Licenses/GPL-3)
 [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
+<!-- badges: end -->
 

### 2019-05-13 README.Rmd
page

@@
 [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
+[![codecov](https://codecov.io/gh/agrobioinfoservices/gosset/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/agrobioinfoservices/gosset)
+[![CRAN status](https://www.r-pkg.org/badges/version/gosset){data-external="1"}](https://cran.r-project.org/package=gosset)
 <!-- badges: end -->

### 2019-05-13 README.Rmd
travis

--- 2019-05-13 README.Rmd
@@
 [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
+[![Build Status](https://travis-ci.org/agrobioinfoservices/gosset.svg?branch=master)](https://travis-ci.org/agrobioinfoservices/gosset)
+[![codecov](https://codecov.io/gh/agrobioinfoservices/gosset/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/agrobioinfoservices/gosset)
+[![CRAN status](https://www.r-pkg.org/badges/version/gosset){data-external="1"}](https://cran.r-project.org/package=gosset)
 <!-- badges: end -->

### 2019-05-14 README.Rmd
pseudoR2_not_S3method

--- 2019-05-13 README.Rmd
@@
 [![GPLv3 license](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.r-project.org/Licenses/GPL-3)
+[![CRAN status](https://www.r-pkg.org/badges/version/gosset){data-external="1"}](https://cran.r-project.org/package=gosset)
 [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
+[![Build Status](https://travis-ci.org/agrobioinfoservices/gosset.svg?branch=master)](https://travis-ci.org/agrobioinfoservices/gosset)
+[![codecov](https://codecov.io/gh/agrobioinfoservices/gosset/master.svg)](https://codecov.io/github/agrobioinfoservices/gosset?branch=master)
 <!-- badges: end -->

### 2019-05-14 R/pseudoR2.R

@@ -1,7 +1,7 @@
-#' Pseudo R-squared 
+#' Pseudo R-squared
 #' 
-#' Regression coefficient to evaluate goodness-of-fit in a given model  
-#' 
+#' Regression coefficient to evaluate goodness-of-fit in a given model
+#'
 #' @param object a model object for which pseudo R-squared is desired
 #' @param ... additional arguments affecting the R-squared produced
 #' @return A tidy object containing the pseudo R-squared coefficients:
@@ -11,50 +11,80 @@
 #' \item{CraggUhler}{Cragg and Uhler's pseudo R-squared}
 #' \item{Agresti}{Agresti pseudo R-squared}
 #' @examples
-#' 
+#'
 #' data("airquality")
-#' 
-#' mod <- glm(Temp ~ Wind + Solar.R, 
-#'            data = airquality, 
+#'
+#' mod <- glm(Temp ~ Wind + Solar.R,
+#'            data = airquality,
 #'            family = poisson())
-#' 
+#'
 #' pseudoR2(mod)
-#'  
+#'
 #' #################################
 #' \dontrun{
-#' 
+#'
 #' # Compute pseudo R-squared on a validation sample
-#' 
+#'
 #' n <- nrow(airquality)
-#' 
+#'
 #' s <- sample(1:n, n * 0.7)
-#' 
+#'
 #' train <- airquality[s,]
 #' test <- airquality[-s,]
-#' 
+#'
 #' mod <- glm(Temp ~ Wind + Solar.R,
 #'            data = train,
 #'            family = poisson())
-#' 
+#'
 #' pseudoR2(mod, newdata = test)
-#' 
+#'
 #' }
 #' @export
-pseudoR2 <- function(object, ...){
+pseudoR2 <- function(object, ...) {
   
-  UseMethod("pseudoR2")
+  cl <- class(object)[[1]]
   
+  if (cl == "pltree") {
+    pr2 <- .pseudoR2_pltree(object, ...)
+  }
+  
+  if (cl == "bttree") {
+    pr2 <- .pseudoR2_bttree(object, ...)
+  }
+  
+  if (cl != "pltree" & cl != "bttree") {
+    pr2 <- .pseudoR2_default(object, ...)
+  }
+  
+  return(pr2)
 }
 
-#' @export
-pseudoR2.default <- function(object, newdata = NULL){
-  
-  # update the model as a null model 
-  # model without covariates 
[…]

### 2019-05-14 README.Rmd
codecov_travisyml

--- 2019-05-13 README.Rmd
@@
 <!-- badges: start -->
+[![CRAN status](https://www.r-pkg.org/badges/version/gosset)](https://cran.r-project.org/package=gosset)
+[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
+[![Build Status](https://travis-ci.org/agrobioinfoservices/gosset.svg?branch=master)](https://travis-ci.org/agrobioinfoservices/gosset)
+[![codecov](https://codecov.io/gh/agrobioinfoservices/gosset/master.svg)](https://codecov.io/github/agrobioinfoservices/gosset?branch=master)
 [![GPLv3 license](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.r-project.org/Licenses/GPL-3)
-[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/agrobioinfoservices/gosset?branch=master&svg=true)](https://ci.appveyor.com/project/kauedesousa/gosset)
-[![codecov](https://codecov.io/gh/agrobioinfoservices/gosset/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/agrobioinfoservices/gosset)
-[![CRAN status](https://www.r-pkg.org/badges/version/gosset){data-external="1"}](https://cran.r-project.org/package=gosset)
 <!-- badges: end -->

### 2019-05-21 R/akaike_weights.R
akaike_weights

@@ -1,0 +1,52 @@
+#' Akaike weights
+#' 
+#' Akaike weights represent the relative likelihood of a model. It can be used in model averaging and selection.
+#' 
+#' @param object a numerical vector with models goodness of fit coefficients 
+#' @return A data frame containing the coefficients:
+#' \item{delta}{the delta overall change in the coefficients}
+#' \item{relative_logLik}{the relative log-likelihood}
+#' \item{akaike_weights}{the Akaike weights}
+#' @examples
+#' 
+#' data("airquality")
+#' 
+#' # try three model approaches
+#' mod1 <- glm(Temp ~ 1,
+#'             data = airquality,
+#'             family = poisson())
+#' 
+#' mod2 <- glm(Temp ~ Ozone,
+#'             data = airquality,
+#'             family = poisson())
+#' 
+#' mod3 <- glm(Temp ~ Ozone + Solar.R,
+#'             data = airquality,
+#'             family = poisson())
+#' 
+#' 
+#' akaike_weights(c(AIC(mod1), AIC(mod2), AIC(mod3)))
+#'  
+#' @export 
+akaike_weights <- function(object){
+
+  # take the delta value
+  delta <- object - min(object, na.rm = TRUE)
+  
+  # and the relative log likelihood
+  rel_LL <- exp(-0.5 * delta)
+  
+  # sum all relative log likelihoods
+  sum_LL <- sum(rel_LL, na.rm = TRUE)
+  
+  # get the Akaike weight
+  weights <- rel_LL / sum_LL
+
+  # combine values into a single vector
+  result <- tibble::tibble(delta = delta, 
+                           relative_logLik = rel_LL, 
+                           akaike_weights = weights)
+  
+  return(result)
+  
+}

### 2019-05-21 R/crossvalidation.R

@@ -14,7 +14,7 @@
 #' Options are: 'equal', an arithmetic mean; 
 #' 'foldsize', a weighted mean by the size in each fold; 
 #' 'stouffer' a weighted Z-test developed by Stouffer et al. (1949). 
-#' The two last methods are suggested for cross-validation with imbalaced folds
+#' The two last methods are suggested for cross-validation with imbalanced folds
 #' @param ... additional arguments passed to methods
 #' @return The cross-validation goodness-of-fit estimates for the best model, which are:
 #' \item{AIC}{Akaike Information Criterion}
@@ -178,7 +178,10 @@ crossvalidation <- function(formula, data, k = NULL,
   # estimators are then averaged weighted by 
   # number of predicted cases using selected mean method
   means <- apply(estimators, 2, function(x) {
-      .mean_crossvalidation(x, folds, method = mean.method)
+      .mean_crossvalidation(object = x, 
+                            folds = folds, 
+                            method = mean.method, 
+                            ...)
     })
   
   # means and estimates as tibble
@@ -215,23 +218,17 @@ print.crossvalidation <- function(x, ...) {
 }
 
 # Compute weighted means in cross-validation
-.mean_crossvalidation <- function(object, folds = NULL, method = NULL){
-  
-  if (is.null(method)) {method <- "foldsize"}
-  
+.mean_crossvalidation <- function(object, folds = NULL, 
+                                  method = NULL, akaike.weights = FALSE, 
+                                  ...){
   # take length of folds
   N <- length(folds)
   
-  if (method == "foldsize") {
-    # make a table of folds and take
-    # the number of observations per fold
-    foldsize <- as.vector(table(folds))
-    
-    # fold size mean is the product of multiplication of object values by 
-    # its number of observations then divided by the total number of observations
-    mean <- sum(object * foldsize, na.rm = TRUE) / sum(foldsize, na.rm = TRUE)
-  }
+  if (is.null(method)) {method <- "stouffer"}
   
+  if (akaike.weights) {method <- "stouffer"}
+  
+  # weight of imbalanced folds
   if (method == "stouffer") {
     # take the number of folds
     max_folds <- max(folds)
@@ -247,8 +244,12 @@ print.crossvalidation <- function(x, ...) {
     
     # then divide this vector by its sum
     wfold <- wfold / sum(wfold)
+    # wfold <- (max_folds * wfold) / sum(wfold)
     
-    wfold <- (max_folds * wfold) / sum(wfold)
[…]

### 2019-05-21 R/forward.R

@@ -70,7 +70,7 @@ forward <- function(formula, data, k = NULL, folds = NULL,
   }
   
   if (is.null(ncores)) {
-     ncores <- 1
+    ncores <- 1
   }
   
   # define initial parameters for forward selection
@@ -79,7 +79,7 @@ forward <- function(formula, data, k = NULL, folds = NULL,
   if (select.by %in% c("AIC","deviance")) {
     baseline <-  1e+11
   } else {
-  # if other method, take 0 as baseline
+    # if other method, take 0 as baseline
     baseline <- 0
   }
   
@@ -133,9 +133,9 @@ forward <- function(formula, data, k = NULL, folds = NULL,
                                    .combine = .comb,
                                    .packages = packages) %dopar% (.forward_dopar(as.formula(
                                      paste0(Y, " ~ ", paste(c(var_keep, exp_var[i]), collapse = " + "))
-                                     ),
-                                     args)))
-   
+                                   ),
+                                   args)))
+    
     models <- t(models)
     
     # take the model with best parameter
@@ -151,7 +151,7 @@ forward <- function(formula, data, k = NULL, folds = NULL,
       value_best <- modpar[index_best]
       
       best <- .is_lower(value_best, baseline)
-    
+      
     } else {
       
       index_best <- which.max(modpar)
@@ -200,8 +200,8 @@ forward <- function(formula, data, k = NULL, folds = NULL,
     # prevent while loop to broke when the model fits with all variables
     if(length(exp_var) == 0) {
       best <- FALSE
-      }
- 
+    }
+    
   }
   
   # Stop cluster connection
@@ -259,27 +259,3 @@ forward <- function(formula, data, k = NULL, folds = NULL,
   x < y
 }
 
-# # compute Akaike weights
-# .akaike_weights <- function(object){
-# 
-#   # remove NA values
-#   object <- object[!is.na(object)]
-# 
-#   # take the delta(AIC) value
-#   delta_aic <- object - min(object, na.rm = TRUE)
-# 
-#   # and the relative log likelihood
-#   rel_LL <- exp(-0.5 * delta_aic)
-# 
-#   # sum all relative log likelihoods
-#   sum_LL <- sum(rel_LL, na.rm = TRUE)
-# 
-#   # get the Akaike weight 
-#   weights_aic <- rel_LL / sum_LL
-# 
-#   # combine values into a single vector
-#   result <- list(deltaAIC = delta_aic, rel_LL = rel_LL, weights = weights_aic)
-# 
-#   return(result)
-# 
-# }

### 2019-05-21 R/kendallTau.R

@@ -67,7 +67,7 @@ kendallTau <- function(x, y){
                    obr <- X[1:nc]
                    prr <- X[(nc + 1):(2 * nc)]
                    
-                   keep <- !is.na(obr)
+                   keep <- !is.na(obr) & !is.na(prr)
                    
                    obr <- obr[keep]
                    

### 2019-05-24 tests/testthat/test-forward.R
forward

@@ -1,0 +1,53 @@
+context("test-forward")
+
+library("gosset")
+library("doParallel")
+library("gnm")
+
+data("airquality")
+
+k <- 3
+folds <- rep(1:3, times = nrow(airquality)/3)
+
+mod1 <- forward(Temp ~ .,
+                data = airquality, 
+                k = k,
+                folds = folds,
+                select.by = "AIC",
+                family = poisson(link = "log"))
+
+
+mod2 <- forward(Temp ~ .,
+                data = airquality, 
+                k = k,
+                folds = folds,
+                select.by = "deviance",
+                akaike.weights = TRUE,
+                family = poisson(link = "log"))
+
+mod1 <- as.vector(t(mod1$coeffs))
+coeffs1 <-
+  c(252.9602385,
+    244.9602385,
+    -122.4801192,
+    0.4125532,
+    0.4145210,
+    0.1001847)
+
+
+mod2 <- as.vector(t(mod2$coeffs))
+coeffs2 <-
+  c(265.9658315,
+    261.9658315,
+    -130.9829158,
+    0.3515344,
+    0.3528859,
+    0.0777664)
+
+test_that("model with deviance", {
+  expect_equal(mod1, coeffs1)
+})
+
+test_that("model with deviance mean equal", {
+  expect_equal(mod2, coeffs2)
+})

### 2019-08-21 R/crossvalidation.R
v0.1-1

@@ -14,7 +14,7 @@
 #' Options are: 'equal', an arithmetic mean; 
 #' 'foldsize', a weighted mean by the size in each fold; 
 #' 'stouffer' a weighted Z-test developed by Stouffer et al. (1949). 
-#' The two last methods are suggested for cross-validation with imbalanced folds
+#' @param object a model object of class \code{"crossvalidation"}.
 #' @param ... additional arguments passed to methods
 #' @return The cross-validation goodness-of-fit estimates, which are:
 #' \item{AIC}{Akaike Information Criterion}
@@ -27,8 +27,7 @@
 #' @seealso \code{\link[gnm]{gnm}}, \code{\link[PlackettLuce]{pltree}}, 
 #' \code{\link[psychotree]{bttree}}
 #' @examples
-#' \donttest{
-#' 
+#'  
 #' # Generalized Linear Models
 #' 
 #' data("airquality")
@@ -37,9 +36,11 @@
 #'                 data = airquality,
 #'                 k = 3,
 #'                 family = poisson())
+#'                 
 #' 
 #' ########################################
-#' 
+#' @examples 
+#' \dontrun{
 #' # PlackettLuce Model
 #' # beans data from PlackettLuce
 #' library("PlackettLuce")
@@ -228,7 +229,6 @@ crossvalidation <- function(formula, data, k = NULL,
   return(result)
 }
 
-#' @rdname crossvalidation
 #' @method print crossvalidation
 #' @export
 print.crossvalidation <- function(x, ...) {
@@ -241,19 +241,23 @@ print.crossvalidation <- function(x, ...) {
 #' @rdname crossvalidation
 #' @method predict crossvalidation
 #' @export
-predict.crossvalidation <- function(object, data = NULL, ...) {
+predict.crossvalidation <- function(object, ...) {
+  
+  dots <- list(...)
+  
+  newdata <- dots[["newdata"]]
+  
+  if (is.null(newdata)) {
+    newdata <- object$raw$data
+  }
   
   models <- object$raw$models
   
-  if (is.null(data)) {
-    data <- object$raw$data
-  }
-  
   preds <- lapply(models, function(x){
-    predict(x, data)
+    predict(object = x, newdata = newdata)
   })
   
-  dims <- c(dim(data)[[1]],
+  dims <- c(dim(newdata)[[1]],
             dim(preds[[1]])[[2]], 
             length(preds))
   

### 2019-08-21 R/kendallTau.R

@@ -43,7 +43,7 @@ kendallTau <- function(x, y){
   if (any(is_decimal(x))) {
     
    x <- t(apply(x, 1, function(X){
-      t(num2rank(X)["rank"])
+      t(rank_decimal(X)["rank"])
     }))
   
   }
@@ -51,7 +51,7 @@ kendallTau <- function(x, y){
   if (any(is_decimal(y))) {
     
     y <- t(apply(y, 1, function(X){
-      t(num2rank(X)["rank"])
+      t(rank_decimal(X)["rank"])
     }))
     
   }

### 2019-08-21 R/rank_PL.R

@@ -6,11 +6,18 @@
 #' An id is required if long format. 
 #' @param items a data frame or index of \code{data} specifying the column(s) containing the item names
 #' @param input a data frame or index of \code{data} specifying the column(s) containing the values to be ranked
-#' @param ... additional arguments passed to methods
+#' @param ... additional arguments passed to methods. See Details
 #' @return a PlackettLuce "rankings" object, which is a matrix of dense rankings 
 #' @seealso \code{\link[PlackettLuce]{rankings}}
+#' @details 
+#' type: a character describing the type of data, c("rank", "tricot")
+#' 
+#' grouped.rankings: logical, to return an object of class "grouped_rankings"
+#' 
+#' id: a data frame or index of \code{data} indicating the ids for "long" rankings
+#' 
+#' add.rank: a data frame for additional rankings if data of type "tricot".
 #' @examples
-#'  
 #' # A matrix with 10 rankings of 5 items (A, B, C, D, E)
 #' # with numeric values as "rank"
 #' set.seed(123)
@@ -19,27 +26,27 @@
 #'             input = runif(50, 1, 3))
 #' 
 #' # return an object of class 'rankings'
-#' R <- to_rankings(df,
-#'                  items = 2,
-#'                  input = 3,
-#'                  id = 1)
+#' R <- rank_PL(df,
+#'              items = 2,
+#'              input = 3,
+#'              id = 1)
 #' 
 #' # rankings can be computed in ascending order
-#' R <- to_rankings(df,
-#'                  items = 2,
-#'                  input = 3,
-#'                  id = 1, 
-#'                  ascending = TRUE)
+#' R <- rank_PL(df,
+#'              items = 2,
+#'              input = 3,
+#'              id = 1,
+#'              ascending = TRUE)
 #' 
 #' 
 #' # return an object of class 'grouped_rankings'
-#' R <- to_rankings(df,
-#'                  items = 2,
-#'                  input = 3,
-#'                  id = 1,
-#'                  grouped.rankings = TRUE)
-#'  
-#' ##################################
+#' R <- rank_PL(df,
+#'              items = 2,
+#'              input = 3,
+#'              id = 1,
+#'              grouped.rankings = TRUE)
+#' 
+#' ######################
 #' 
 #' # Rankings with 5 items randomly assigned
 #' 
@@ -53,20 +60,20 @@
 #'   i[s,] <- sample(LETTERS[1:5])
[…]

### 2019-08-21 R/rank_binomial.R

@@ -20,8 +20,8 @@
 #'   r[s,] <- sample(1:5)
 #' }
 #' 
-#' R <- to_rankings(items = i,
-#'                  input = r)
+#' R <- rank_PL(items = i,
+#'              input = r)
 #' 
 #' rank_binomial(R)
 #' 
@@ -30,11 +30,11 @@ rank_binomial <- function(object, drop.null = FALSE)
 {
   
   if (.is_grouped_rankings(object)) {
-    object <- to_paircomp(object)
+    object <- rank_paircomp(object)
   }
   
   if (.is_rankings(object)) {
-    object <- to_paircomp(object)
+    object <- rank_paircomp(object)
   }
   
   # take all possible combinations between items
@@ -58,13 +58,6 @@ rank_binomial <- function(object, drop.null = FALSE)
   
   # take the name of all items
   players <- factor(sort(unique(as.vector(unlist(object[,c("player1","player2")])))))
-  
-  # take all possible combinations between items
-  #comb <- t(.combn2(players, 2))
-  
-  # generate two vector with equal factors corresponding to players
-  #player1 <- factor(players[comb[, 1]], levels = players)
-  #player2 <- factor(players[comb[, 2]], levels = players)
   
   # run over object looking for corresponding combination and 
   # add the results into the new dataframe object_bin

### 2019-08-21 R/rank_paircomp.R

@@ -21,7 +21,7 @@
 #' colnames(R) <- c("apple", "banana", "orange", "pear")
 #' R <- PlackettLuce::as.rankings(R)
 #' 
-#' to_paircomp(R)
+#' rank_paircomp(R)
 #' 
 #' #############################################
 #' 
@@ -30,17 +30,17 @@
 #' 
 #' # convert the tricot rankings from breadwheat data
 #' # into a object of class 'rankings' from PlackettLuce
-#' R <- to_rankings(breadwheat,
-#'                  items = c("variety_a","variety_b","variety_c"),
-#'                  rankings = c("overall_best","overall_worst"),
-#'                  type = "tricot")
+#' R <- rank_PL(breadwheat,
+#'              items = c("variety_a","variety_b","variety_c"),
+#'              rankings = c("overall_best","overall_worst"),
+#'              type = "tricot")
+#'  
 #' 
-#' 
-#' PC <- to_paircomp(R)
+#' PC <- rank_paircomp(R)
 #' 
 #' 
 #' @export
-to_paircomp <- function(object){
+rank_paircomp <- function(object){
   
   R <- object
   

### 2019-08-21 R/utility.R

@@ -36,7 +36,7 @@ is_decimal <- function(x) {
   return(x)
 }
 
-#' Numbers into rankings
+#' Decimals into rankings
 #'
 #' Group and convert numeric values into integer ranks
 #'
@@ -52,7 +52,7 @@ is_decimal <- function(x) {
 #' 
 #' x <- c(1, 2, 1.4, 3, 4.01, -0.5)
 #' 
-#' num2rank(x)
+#' rank_decimal(x)
 #' 
 #' # passing a vector with an id
 #' # ids are used to group values and 
@@ -61,10 +61,10 @@ is_decimal <- function(x) {
 #' x <- c(1, 2, 1.4, 3, 4.01, -0.5)
 #' id <- c(rep(1, 3), rep(2, 3))
 #' 
-#' num2rank(x, id = id)
+#' rank_decimal(x, id = id)
 #' 
 #' @export
-num2rank <- function(object, id = NULL, ...){
+rank_decimal <- function(object, id = NULL, ...){
   
   dots <- list(...)
   

### 2019-08-21 tests/testthat/test-rank_PL.R

@@ -1,4 +1,4 @@
-context("test-rankings")
+context("test-rank_PL")
 
 library("gosset")
 library("PlackettLuce")
@@ -46,7 +46,7 @@ R <- grouped_rankings(R, rep(1:n, 4))
 R <- R[1:length(R), , as.grouped_rankings = FALSE]
 
 # rankings produced by to_rankings
-G <- to_rankings(data = tricot,
+G <- rank_PL(data = tricot,
                  items = c("variety_a","variety_b","variety_c"),
                  input = c("best","worst"),
                  type = "tricot",

### 2019-11-27 README.Rmd
akaike_weights
gosset

@@
   comment = "#>",
+  fig.path = "man/figures/README-",
   out.width = "100%"
@@
 <!-- badges: end -->
+
+# _gosset_: Modelling Metadata and Experimental Citizen Science <img align="right" src="man/figures/logo.png">
 
@@
 
-You are welcome to contribute to this project. Please read our [contribution guide lines](CONTRIBUTING.md).
+You are welcome to contribute to this project. Please read our [contribution guideline](CONTRIBUTING.md).
 

### 2019-11-27 R/akaike_weights.R

@@ -1,6 +1,7 @@
 #' Akaike weights
 #' 
-#' Akaike weights represent the relative likelihood of a model. It can be used in model averaging and selection.
+#' Akaike weights represent the relative likelihood of a model. 
+#' It can be used in model averaging and selection.
 #' 
 #' @param object a numerical vector with models goodness of fit coefficients 
 #' @return A data frame containing the coefficients:

### 2019-11-27 R/forward.R

@@ -1,365 +1,0 @@
-#' Forward stepwise regression for model selection
-#'
-#' Forward selection is a type of stepwise regression which begins with an empty 
-#' model and adds in variables one by one. In each step one variable is added
-#' when it makes a single best improvement to the model.
-#'
-#' @param formula a symbolic description of the model, if set as   
-#'   \eqn{ y ~ . }   all variables in data are used  
-#' @param select.by a character for the goodness-of-fit statistical parameter to
-#' select the models. Set as 'deviance' by default.
-#' @param ncores an integer for the number of cores to be used in the parallel computing.
-#' Set as 1 by default (no parallelisation)
-#' @param packages an optional character vector of packages that the parallel tasks depend on
-#' @param akaike.weights an optional logical object for averaging the goodness of fit coefficients with Akaike weights
-#' @inheritParams crossvalidation
-#' @return The cross-validation goodness-of-fit estimates for the best model, which are:
-#' \item{AIC}{Akaike Information Criterion}
-#' \item{deviance}{Model deviance}
-#' \item{logLik}{Log-Likelihood}
-#' \item{MaxLik}{Maximum likelihood pseudo R-squared}
-#' \item{CraggUhler}{Cragg and Uhler's pseudo R-squared}
-#' \item{Agresti}{Agresti pseudo R-squared}
-#' Cross-validation estimates are computed using the fitted models on the validation samples.
-#' @seealso \code{\link{crossvalidation}}
-#' @examples 
-#' \dontrun{
-#' 
-#' library("gnm")
-#' library("doParallel")
-#' library("abind")
-#' 
-#' data("airquality")
-#' 
-#' mod <- forward(Temp ~ .,
-#'                data = airquality,
-#'                k = 3,
-#'                select.by = "AIC",
-#'                family = poisson(link = "log"))
-#' 
-#' }
-#' @import abind
-#' @import foreach
-#' @import doParallel
-#' @export
-forward <- function(formula, data, k = NULL, folds = NULL, 
-                    select.by = NULL, akaike.weights = FALSE,
-                    ncores = NULL, packages = NULL, ...) {
-  
-  n <- nrow(data)
-  
-  # list of additional arguments
-  dots <- list(...)
-  
-  # number of cores to be used
-  if (is.null(ncores)) {
-    ncores <- 1
-  }
-  
-  # check/define folds before starting forward selection
-  # to make sure that all steps will use the same sample
[…]

### 2019-11-27 R/rank_PL.R

@@ -1,558 +1,0 @@
-#' Plackett-Luce rankings
-#'
-#' Create an object of class "rankings" from a dataframe or matrix
-#'
-#' @param data a data frame with columns specified by items and input values to rank. Data can be in both wide and long format.
-#' An id is required if long format. 
-#' @param items a data frame or index of \code{data} specifying the column(s) containing the item names
-#' @param input a data frame or index of \code{data} specifying the column(s) containing the values to be ranked
-#' @param ... additional arguments passed to methods. See Details
-#' @return a PlackettLuce "rankings" object, which is a matrix of dense rankings 
-#' @seealso \code{\link[PlackettLuce]{rankings}}
-#' @details 
-#' type: a character describing the type of data, c("rank", "tricot")
-#' 
-#' grouped.rankings: logical, to return an object of class "grouped_rankings"
-#' 
-#' id: a data frame or index of \code{data} indicating the ids for "long" rankings
-#' 
-#' add.rank: a data frame for additional rankings if data of type "tricot".
-#' @examples
-#' # A matrix with 10 rankings of 5 items (A, B, C, D, E)
-#' # with numeric values as "rank"
-#' set.seed(123)
-#' df <- cbind(id = rep(1:10, each = 5),
-#'             items = rep(LETTERS[1:5], times = 10),
-#'             input = runif(50, 1, 3))
-#' 
-#' # return an object of class 'rankings'
-#' R <- rank_PL(df,
-#'              items = 2,
-#'              input = 3,
-#'              id = 1)
-#' 
-#' # rankings can be computed in ascending order
-#' R <- rank_PL(df,
-#'              items = 2,
-#'              input = 3,
-#'              id = 1,
-#'              ascending = TRUE)
-#' 
-#' 
-#' # return an object of class 'grouped_rankings'
-#' R <- rank_PL(df,
-#'              items = 2,
-#'              input = 3,
-#'              id = 1,
-#'              grouped.rankings = TRUE)
-#' 
-#' ######################
-#' 
-#' # Rankings with 5 items randomly assigned
-#' 
-#' i <- as.data.frame(matrix(NA, nrow = 10, ncol = 5))
-#' names(i) <- paste0("Item",1:5)
-#' 
-#' r <- as.data.frame(matrix(NA, nrow = 10, ncol = 5))
[…]

### 2019-11-27 R/rank_binomial.R

@@ -4,7 +4,7 @@
 #'
 #' @param object an object of class "rankings", "grouped_rankings" or "paircomp".
 #' @param drop.null logical, an optional argument to remove players without comparisons 
-#' @return a binary rank
+#' @return a binary rank of pairwise contests
 #' @seealso \code{\link[PlackettLuce]{rankings}}, \code{\link[psychotools]{paircomp}}
 #' @examples 
 #' # Rankings with 5 items randomly assigned
@@ -20,8 +20,8 @@
 #'   r[s,] <- sample(1:5)
 #' }
 #' 
-#' R <- rank_PL(items = i,
-#'              input = r)
+#' R <- rank_numeric(items = i,
+#'                  input = r)
 #' 
 #' rank_binomial(R)
 #' 

### 2019-11-27 R/rank_numeric.R

@@ -1,0 +1,280 @@
+#' Plackett-Luce rankings from numeric values
+#'
+#' Group and convert numeric values into integer ranks
+#'
+#' @inheritParams rank_tricot
+#' @param id a data frame or index of \code{data} indicating the ids for "long" \code{data}
+#' @param ascending logical, to compute rankings from lower to higher values
+#' @param ... additional arguments passed to methods
+#' @return a PlackettLuce "rankings" object, which is a matrix of dense rankings 
+#' @seealso \code{\link[PlackettLuce]{rankings}} 
+#' @examples
+#' # A matrix with 10 rankings of 5 items (A, B, C, D, E)
+#' # with numeric values as "rank"
+#' set.seed(123)
+#' df <- cbind(id = rep(1:10, each = 5),
+#'             items = rep(LETTERS[1:5], times = 10),
+#'             input = runif(50, 1, 3))
+#' 
+#' # return an object of class 'rankings'
+#' R <- rank_numeric(df,
+#'                   items = 2,
+#'                   input = 3,
+#'                   id = 1)
+#' 
+#' # rankings can be computed in ascending order
+#' R <- rank_numeric(df,
+#'                   items = 2,
+#'                   input = 3,
+#'                   id = 1,
+#'                   ascending = TRUE)
+#' 
+#' 
+#' # return an object of class 'grouped_rankings'
+#' R <- rank_numeric(df,
+#'                   items = 2,
+#'                   input = 3,
+#'                   id = 1,
+#'                   grouped.rankings = TRUE)
+#' 
+#' ######################
+#' 
+#' # Rankings with 5 items randomly assigned
+#' 
+#' i <- as.data.frame(matrix(NA, nrow = 10, ncol = 5))
+#' names(i) <- paste0("Item",1:5)
+#' 
+#' r <- as.data.frame(matrix(NA, nrow = 10, ncol = 5))
+#' names(r) <- paste0("Position_Item",1:5)
+#' 
+#' for(s in 1:10) {
+#'   i[s,] <- sample(LETTERS[1:5])
+#'   r[s,] <- sample(1:5)
+#' }
+#' 
+#' R <- rank_numeric(items = i,
+#'                   input = r)
+#'  
+#' @import tibble
+#' @importFrom tidyr gather spread
+#' @export
+rank_numeric <- function(data = NULL, items = NULL,
+                    input = NULL, id = NULL, 
[…]

### 2019-11-27 R/rank_paircomp.R

@@ -30,11 +30,10 @@
 #' 
 #' # convert the tricot rankings from breadwheat data
 #' # into a object of class 'rankings' from PlackettLuce
-#' R <- rank_PL(breadwheat,
-#'              items = c("variety_a","variety_b","variety_c"),
-#'              rankings = c("overall_best","overall_worst"),
-#'              type = "tricot")
-#'  
+#' R <- rank_tricot(breadwheat,
+#'                  items = c("variety_a","variety_b","variety_c"),
+#'                  input = c("overall_best","overall_worst"))
+#' 
 #' 
 #' PC <- rank_paircomp(R)
 #' 

### 2019-11-27 R/rank_tricot.R

@@ -1,0 +1,343 @@
+#' Build Plackett-Luce rankings from tricot dataset 
+#'
+#' Create an object of class "rankings" from tricot data. Tricot stands for "triadic 
+#' comparison of technologies". Is a methodology to carry out large agronomic field experiments allowing 
+#' the comparison between many alternative technologies, in many different environments. Each participant
+#' evaluates a set of three randomised technologies from a larger set. 
+#'
+#' @param data a data frame with columns specified by items and input values
+#' @param items a data frame or index of \code{data} for the column(s) containing the item names
+#' @param input a data frame or index of \code{data} for the column(s) containing the values to be ranked
+#' @param additional.rank optional, a data frame for the comparisons between tricot items and the local item
+#' @param group logical, if TRUE return an object of class "grouped_rankings"
+#' @param ... additional arguments passed to methods. See details
+#' @return a PlackettLuce "rankings" object, which is a matrix of dense rankings 
+#' @seealso \code{\link[PlackettLuce]{rankings}}  \code{\link{breadwheat}}
+#' @details 
+#' full.output: logical, to return a list with a "rankings", a "grouped_rankings" and the ordered items
+#' @examples
+#' 
+#' # using breadwheat data
+#' data("breadwheat", package = "gosset")
+#' 
+#' # convert the tricot rankings from breadwheat data
+#' # into a object of class 'rankings' from PlackettLuce
+#' R <- rank_tricot(breadwheat,
+#'                  items = c("variety_a","variety_b","variety_c"),
+#'                  input = c("overall_best","overall_worst"))
+#'                  
+#' ############################################################
+#'   
+#' # beans data where each observer compares 3 varieties randomly distributed
+#' # from a list of 11 and additionally compares these 3 varieties
+#' # with their local variety
+#' library("PlackettLuce")
+#' data("beans", package = "PlackettLuce")
+#' 
+#' # first build rankings with only tricot items
+#' # and return an object of class 'rankings'
+#' R <- rank_tricot(data = beans,
+#'                     items = c(1:3),
+#'                     input = c(4:5))
+#' head(R)
+#' 
[…]

### 2019-11-27 dev/bt_permute.R

@@ -1,0 +1,301 @@
+# library("BradleyTerry2")
+# 
+# load("data/gossetJonathanData.rda")
+# 
+# #########
+# 
+# #items <- KenyaItems ### function permic will only work if "items" is defined beforehand! Ridiculous...
+# iter <- 100
+# contests <- KenyaChoiceData
+# predictors <- KenyaFarmerData
+# 
+# #########
+# bt_permute <- function(contests = NULL, predictors = NULL, iter = 100){
+# 
+#   #### User must specify number of iterations (permutations), (nIterations)
+#   #### a factor of the items that were ranked,                (items) ## but see the problem mentioned above
+#   #### the observer's choices in specific (pairwise) format,  (choices)
+#   #### and observer characteristics.                          (variables)
+# 
+#   # change names in contest
+#   names(contests) <- c("ObsID","Item1","Item2","Win1","Win2")
+#   # take items names
+#   items <- levels(contests$Item1)
+#   items2 <- levels(contests$Item2)
+#   
+#   not_factor <- c(!is.factor(contests$Item1), !is.factor(contests$Item2))
+#   
+#   if (any(not_factor)) {
+#     stop("Items should be a factor with the same levels")
+#   }
+#   
+#   if (items != items2) {
+#     stop("Items should be a factor with the same levels")
+#   }
+#   
+#   # take number of predictors
+#   nvar <- ncol(predictors)
+#   
+#   if(nvar > 100) {
+#     stop("Subscript out of bounds. Cannot handle more than 100 predictors")
+#   }
+#   
+#   # put input data into a list BradleyTerry2::BTm
+#   dat <- list()
+#   dat[[1]] <- contests
+#   dat[[2]] <- predictors
+#   dat[[3]] <- as.data.frame(diag(length(items)))
+# 
+#   # take original names of predictors
+#   origVarnames <- names(dat[[2]])
+#   origVarnames <- c(origVarnames, rep("Permuted variable", times = nvar))
+#   
+#   # create generalized names for predictors
+#   Vars <- sprintf("Var%1d", 1:nvar)
+#   names(dat[[2]]) <- Vars
+#   
+#   # create generalized names for permuted predictors
+#   permVars <- sprintf("PV%1d", 1:nvar)
+#   
+#   # combine both vectors
+#   allVars <- c(Vars, permVars)
+# 
+#   # make a matrix with equal rows and cols names 
+#   # for items
[…]

### 2019-11-27 dev/forward.R

@@ -1,0 +1,354 @@
+# Forward stepwise regression for model selection#Forward selection is a type of stepwise regression which begins with an empty
+# model and adds in variables one by one. In each step one variable is added
+# when it makes a single best improvement to the model.
+#
+# @param formula a symbolic description of the model, if set as   
+#   \eqn{ y ~ . }   all variables in data are used  
+# @param select.by a character for the goodness-of-fit statistical parameter to
+# select the models. Set as 'deviance' by default.
+# @param ncores an integer for the number of cores to be used in the parallel computing.
+# Set as 1 by default (no parallelisation)
+# @param packages an optional character vector of packages that the parallel tasks depend on
+# @param akaike.weights an optional logical object for averaging the goodness of fit coefficients with Akaike weights
+# @inheritParams crossvalidation
+# @return The cross-validation goodness-of-fit estimates for the best model, which are:
+# \item{AIC}{Akaike Information Criterion}
+# \item{deviance}{Model deviance}
+# \item{logLik}{Log-Likelihood}
+# \item{MaxLik}{Maximum likelihood pseudo R-squared}
+# \item{CraggUhler}{Cragg and Uhler's pseudo R-squared}
+# \item{Agresti}{Agresti pseudo R-squared}
+# Cross-validation estimates are computed using the fitted models on the validation samples.
+# @seealso \code{\link{crossvalidation}}
+# @examples 
+# \dontrun{
+# 
+# library("gnm")
+# library("doParallel")
+# library("abind")
+# 
+# data("airquality")
+# 
+# mod <- forward(Temp ~ .,
+#                data = airquality,
+#                k = 3,
+#                select.by = "AIC",
+#                family = poisson(link = "log"))
+# 
+# }
+# @import abind
+# @import foreach
+# @import doParallel
+# @export
+# forward <- function(formula, data, k = NULL, folds = NULL, 
+#                     select.by = NULL, akaike.weights = FALSE,
+#                     ncores = NULL, packages = NULL, ...) {
+#   
+#   n <- nrow(data)
+#   
+#   # list of additional arguments
+#   dots <- list(...)
+#   
+#   # number of cores to be used
+#   if (is.null(ncores)) {
+#     ncores <- 1
+#   }
+#   
+#   # check/define folds before starting forward selection
+#   # to make sure that all steps will use the same sample
[…]

### 2019-11-27 tests/testthat/test-rank_PL.R

@@ -1,35 +1,0 @@
-context("test-rank_PL")
-
-library("gosset")
-library("PlackettLuce")
-# compare it with the rankings given by to_rankings
-# load("tests/tricot.rda")
-load("../tricot.rda")
-
-testR <- beans[1:10, c(1:5)] 
-
-testR$middle <- complete(testR[c("best", "worst")],
-                         items = c("A", "B", "C"))
-
-testR <- decode(testR[,c("best", "middle", "worst")],
-                items = testR[,c(1:3)],
-                code = LETTERS[1:3])
-
-testR <- as.rankings(testR, input = "ordering")
-
-
-testR <- testR[,, as.rankings = FALSE]
-
-
-test_that("tricot ok", {
-  
-  toTest <- rank_PL(beans[1:10,],
-                    items = c("variety_a","variety_b","variety_c"),
-                    input = c("best","worst"),
-                    type = "tricot")
-  
-  toTest <- toTest[,, as.rankings = FALSE]
-  
-  
-  expect_false(any(testR != toTest))
-})

### 2019-11-27 tests/testthat/test-rank_numeric.R

@@ -1,0 +1,55 @@
+context("test-rank_numeric")
+
+library("gosset")
+library("PlackettLuce")
+# load("tests/test_data.rda")
+load("../test_data.rda")
+
+
+test_that("default ok", {
+  
+  R <- rank_numeric(data = default,
+                    items = 2,
+                    input = 3,
+                    id = 1)
+  
+  R <- R[1:length(R),, as.rankings = FALSE]
+  
+  expect_equal(!any(R == default_true), FALSE)
+  
+})
+
+
+test_that("long and ascending ok", {
+  
+  R <- rank_numeric(data = default,
+                    items = 2,
+                    input = 3,
+                    id = 1,
+                    ascending = TRUE)
+  
+  R <- R[1:length(R),, as.rankings = FALSE]
+  
+  Ras <- t(apply(default_true, 1, function(x) rank(x)))
+  
+  expect_equal(!any(R == Ras), FALSE)
+  
+})
+
+test_that("wide and group ok", {
+  
+  R <- rank_numeric(items = default_wide[[1]],
+                    input = default_wide[[2]],
+                    group = TRUE)
+
+  R <- R[1:length(R),, as.grouped_rankings = FALSE]
+  
+  expect_equal(!any(R == default_wide_true), FALSE)
+  
+})
+
+# 
+# save(lonlat, rain, tetra, tetra_true, triadic,
+#      triadic_true, triadic_true_local, default,
+#      default_true, temp, tricot, default_wide, default_wide_true,
+#      file = "tests/test_data.rda")

### 2019-11-27 tests/testthat/test-rank_tricot.R

@@ -1,0 +1,64 @@
+context("test-rank_tricot")
+
+library("gosset")
+library("PlackettLuce")
+# load("tests/test_data.rda")
+load("../test_data.rda")
+
+test_that("triadic ok", {
+  
+  R <- rank_tricot(data = triadic,
+                      items = c(1:3),
+                      input = c(4:5), 
+                      group = TRUE)
+  
+  R <- R[1:length(R),, as.grouped_rankings = FALSE]
+  
+  expect_equal(!any(R == triadic_true), FALSE)
+  
+})
+
+
+test_that("triadic with local ok", {
+  
+  R <- rank_tricot(data = triadic,
+                      items = c(1:3),
+                      input = c(4:5), 
+                      additional.rank = triadic[,6:8],
+                      group = TRUE)
+  
+  R <- R[1:length(R),, as.grouped_rankings = FALSE]
+  
+  expect_equal(!any(R == triadic_true_local), FALSE)
+  
+})
+
+
+test_that("tretra ok", {
+  
+  R <- rank_tricot(data = NULL,
+                      items = tetra[,c(1:5)],
+                      input = tetra[,c(6:10)], 
+                      group = TRUE)
+  
+  R <- R[1:length(R),, as.grouped_rankings = FALSE]
+  
+  expect_equal(!any(R == tetra_true), FALSE)
+  
+})
+
+
+test_that("full.output works", {
+  
+  R <- rank_tricot(data = triadic,
+                      items = c(1:3),
+                      input = c(4:5), 
+                      additional.rank = triadic[,6:8],
+                      full.output = TRUE)
+  
+  l <- c(!is.list(R), unlist(lapply(R, is.null)))
+
+  expect_equal(any(l), FALSE)
+  
+})
+

### 2019-11-28 README.Rmd
some

@@
 
-# _gosset_: Modelling Metadata and Experimental Citizen Science <img align="right" src="man/figures/logo.png">
-
-Package website: https://agrobioinfoservices.github.io/gosset/
+# *gosset*: Modelling Metadata and Experimental Citizen Science <img align="right" src="man/figures/logo.png">
 
@@
 The **gosset** package provides an environment for modelling metadata and experimental citizen science data.
-It has implementations of cross-validation and stepwise variable selection for glm, lm, Bradley-Terry and Plackett-Luce models.
 

### 2019-11-28 R/ETo.R

@@ -1,6 +1,7 @@
 #' Evapotranspiration
 #' 
-#' Compute evapotranspiration using the Blaney-Criddle method
+#' Compute evapotranspiration using the Blaney-Criddle method. A theoretical method used 
+#' when no measured data on pan evaporation are available locally. 
 #' 
 #' @inheritParams temperature
 #' @param lat a vector for the latitude (in Decimal degrees)
@@ -9,6 +10,9 @@
 #' representing the mean daily percentage of annual daytime hours 
 #' for different latitudes
 #' @return The evapotranspiration in mm/day
+#' @references
+#' Brouwer C. & Heibloem M. (1986). Irrigation water management: Irrigation water needs. Food and Agriculture Organization of The United Nations, Rome, Italy. http://www.fao.org/3/S2022E/s2022e00.htm
+#' 
 #' @examples
 #' \donttest{
 #'  
@@ -24,14 +28,14 @@
 #' pdates <- as.Date(pdates, origin = "1970-01-01")
 #' 
 #' # the evapotranspiration in the first 50 days after planting
-#' ETo(lonlat,
-#'     day.one = pdates,
-#'     span = 50,
-#'     lat = lonlat[["lat"]])
+#' e <- ETo(lonlat,
+#'          day.one = pdates,
+#'          span = 50,
+#'          lat = lonlat[["lat"]])
 #' }
 #' 
 #' @export
-ETo <- function(object, day.one = NULL, span = NULL, 
+ETo <- function(object, day.one = NULL, span = 150, 
                 lat = NULL, Kc = 1, p = NULL){
   
   # remove vector from a tibble object

### 2019-11-28 R/GDD.R

@@ -1,12 +1,17 @@
-#' Growing Degree Days
+#' Growing degree-days
 #' 
-#' Compute number of days required to reach growing degree days.
-#' GDD are calculated by taking the integral of warmth above a base temperature.
+#' Compute number of days required to reach Growing degree-days.
+#' Is a heuristic tool in phenology. Growing degree-days are a measure of heat accumulation 
+#' used to predict plant and animal development rates. Growing degree-days are calculated by 
+#' taking the integral of warmth above a base temperature.
 #' 
 #' @param degree.days an integer for the degree days required by the crop (look at the physiology of the focal crop)
-#' @param base an integer for the base temperature. Set 10 as default.
+#' @param base an integer for the base temperature.
 #' @inheritParams temperature
 #' @return The number of days required to reach the growing degree days.
+#' @references 
+#' Prentice I. C., et al. (1992) Journal of Biogeography, 19(2), 117. https://doi.org/10.2307/2845499.
+#' 
 #' @examples
 #' \donttest{
 #' 
@@ -32,7 +37,7 @@
 #'     
 #' @export
 GDD <- function(object, day.one = NULL, degree.days = NULL,
-                base = NULL, span = NULL, ...)
+                base = 10, span = 150, ...)
 {
   
   # validate parameters
@@ -40,14 +45,7 @@ GDD <- function(object, day.one = NULL, degree.days = NULL,
     stop("degree.days is missing with no default \n")
   }
   
-  if (is.null(base)) {
-    base <- 10
-  }
-  
-  if (is.null(span)) {
-    span <- 150
-  }
-  
+
   # get timespan for the day temperature
   if (dim(object)[2] == 2) {
     day <- .get_timespan(object, day.one, span, pars = "T2M_MAX", ...)

### 2019-11-28 R/favourite.R

@@ -1,24 +1,23 @@
 #' Favourability scores
 #'
-#' Compute favourability scores
+#' Summarise favourability scores from tricot data
 #' 
 #' @inheritParams rank_tricot
 #' @param reorder logical, if items should be reordered from higher favourability score to least favourability score
-#' @param x an object of class 'cmb_fvrt' for the plotting method. Generates a 'ggplot' object that can be passed to any ggplot2 method
+#' @param x an object of class 'gosset_fvrt' for the plotting method. Generates a 'ggplot' object that can be passed to any ggplot2 method
 #' @aliases favorite
-#' @return a data.frame with descriptive statistics
+#' @return A data.frame with the descriptive statistics:
 #' \item{N}{number of times the given item was evaluated}
 #' \item{best}{relative number of times (in percentage) the given item was ranked as first}
 #' \item{worst}{relative number of times (in percentage) the given item was ranked as worst}
 #' \item{wins}{relative number of times (in percentage) the given item wins against the others}
 #' \item{fav_score}{the favourability score, which is the difference between best and worst performance}
 #' @examples
-#' 
 #' data("breadwheat", package = "gosset")
 #' 
-#' favourite(data = breadwheat,
-#'           items = c(1:3),
-#'           input = c(18:19))
+#' fv <- favourite(data = breadwheat,
+#'                 items = c(1:3),
+#'                 input = c(18:19))
 #' 
 #'  
 #' @import ggplot2
@@ -105,7 +104,7 @@ favourite <- function(data = NULL, items = NULL,
     sumstats <- sumstats[rev(order(sumstats$fav_score)), ]
   }
   
-  class(sumstats) <- c("cmb_fvrt", class(sumstats))
+  class(sumstats) <- c("gosset_fvrt", class(sumstats))
   
   return(sumstats)
 
@@ -120,9 +119,9 @@ favorite <- function(...){
 }
 
 #' @rdname favourite
-#' @method plot cmb_fvrt
+#' @method plot gosset_fvrt
 #' @export
-plot.cmb_fvrt <- function(x, ...) {
+plot.gosset_fvrt <- function(x, ...) {
   
   p <- ggplot2::ggplot(data = x, 
                        ggplot2::aes(y = x$fav_score, 

### 2019-11-28 R/preference.R

@@ -1,174 +1,0 @@
-#' Pairwise preference
-#' 
-#' Compute pairwise preference matrix between comparisons
-#' 
-#' @inheritParams rank_tricot
-#' @param x an object of class 'cmb_pref' for the plotting method. Generates a 'ggplot' object that can be passed to any ggplot2 method
-#' @return a data.frame with pairwise comparisons 
-#' \item{player1}{the first player in the comparison}
-#' \item{player2}{the second player in the comparison}
-#' \item{win}{times player1 wins against player2}
-#' \item{ncontest}{number of contests between player1 and player2}
-#' \item{preference}{number of times player1 is preferred against player2}
-#' @examples 
-#' 
-#' # breadwheat data
-#' 
-#' data("breadwheat", package = "gosset")
-#' 
-#' preference(data = breadwheat,
-#'            items = c(1:3),
-#'            input = c(18:19))
-#' 
-#' # beans data where each observer compares 3 varieties
-#' # randomly distributed
-#' 
-#' library("PlackettLuce")
-#' data("beans", package = "PlackettLuce")
-#' 
-#' preference(data = beans,
-#'            items = c(1:3),
-#'            input = c(4:5))
-#' 
-#' # also include the local item
-#' preference(data = beans,
-#'            items = c(1:3),
-#'            input = c(4:5),
-#'            additional.rank = beans[c(6:8)])
-#'    
-#' @importFrom scales percent
-#' @importFrom stats aggregate
-#' @export
-preference <- function(data = NULL, items = NULL, input = NULL, ...){
-
-  # check data
-  data <- .check_data(data = data,
-                      items = items, 
-                      input = input)
-  # take rankings
-  # this allows the function to accept other arguments passed
-  # to build_rankings, such as including the comparison with the
-  # local item
-  dots <- list(...)
-  args <- c(data, dots)
-
-  R <- do.call("rank_tricot", args)  
-
-  # THERE IS AN EASIER SOLUTION USING gosset, SEE BELLOW
-  # BUT WE HAVE TO WAIT UNTIL THE PACKAGE IS READY TO PUBLISH IN CRAN
-  # get adjacency matrix to see how many times each item wins
-  bin <- .rank_binomial(R)
-  
-  # get number of contests
-  bin$ncontest <- bin$win1 + bin$win2
-  
[…]

### 2019-11-28 R/rainfall.R

@@ -17,6 +17,13 @@
 #' \item{Rx1day}{maximum 1-day rainfall (mm)}
 #' \item{Rx5day}{maximum 5-day rainfall (mm) }
 #' \item{Rtotal}{total rainfall (mm) in wet days (R >= 1)}
+#' @references 
+#' Aguilar E., et al. (2005). Journal of Geophysical Research, 110(D23), D23107. https://doi.org/10.1029/2005JD006119.
+#' 
+#' Kehel Z., et al. (2016). Identifying Climate Patterns during the Crop-Growing Cycle from 30 Years of CIMMYT Elite Spring Wheat International Yield Trials. In: Applied Mathematics and Omics to Assess Crop Genetic Resources for Climate Change Adaptive Traits (eds Bari A., Damania A. B., Mackay M., Dayanandan S.), pp. 151–174. CRC Press.
+#' 
+#' Sparks A. H. (2018). Journal of Open Source Software, 3(30), 1035. https://doi.org/10.21105/joss.01035.
+#' 
 #' @examples
 #' \donttest{
 #' 
@@ -46,13 +53,10 @@
 #' }       
 #'          
 #' @export
-rainfall <- function(object, day.one = NULL, span = NULL,
-                     index = NULL, ...)
+rainfall <- function(object, day.one = NULL, span = 150,
+                     index = c("MLDS","MLWS","R10mm","R20mm","SDII","Rx1day","Rx5day","Rtotal"), 
+                     ...)
 {
-  
-  if (is.null(index)) {
-    index <- c("MLDS","MLWS","R10mm","R20mm","SDII","Rx1day","Rx5day","Rtotal")
-  }
   
   # get timespan
   if (dim(object)[2] == 2) {

### 2019-11-28 README.Rmd
tests

@@
 
-The **gosset** package provides an environment for modelling metadata and experimental citizen science data.
+The **gosset** package provides the toolkit for a workflow on modelling metadata and experimental citizen science data, from collating data of different sources, model sellection and visualisation.
+The package is named after W.S.
+Gosset aka 'Student', a pioneer of modern statistics in small sample experimental design and analysis.
 
@@
 
-**gosset** is under development. Use with caution.
-
 ## Going further
 
-The full functionality of **gosset** is illustrated in the package vignette.
-The vignette can be found on the [package website](https://agrobioinfoservices.github.io/gosset/) or from within R once the package has been installed.
+The full functionality of **gosset** is illustrated in the package vignette. The vignette can be found on the [package website](https://agrobioinfoservices.github.io/gosset/).
 

### 2019-11-28 tests/testthat/test-akaike_weights.R

@@ -1,0 +1,40 @@
+context("test-akaike_weights")
+
+
+data("airquality")
+
+# try three model approaches
+mod1 <- glm(Temp ~ 1,
+            data = airquality,
+            family = poisson())
+
+mod2 <- glm(Temp ~ Ozone,
+            data = airquality,
+            family = poisson())
+
+mod3 <- glm(Temp ~ Ozone + Solar.R,
+            data = airquality,
+            family = poisson())
+
+# models AICs together in a single vector
+models <- c(mod1 = AIC(mod1),
+            mod2 = AIC(mod2),
+            mod3 = AIC(mod3))
+
+
+
+test_that("mod3 best", {
+  
+  # calculate akaike weights
+  aw <- akaike_weights(models)
+  
+  # the higher the better
+  aw <- names(models[which.max(aw$akaike_weights)])
+  
+  aw <- aw == "mod3"
+  
+  expect_equal(aw, TRUE)
+  
+})
+
+

### 2019-11-28 tests/testthat/test-kendallTau.R

@@ -1,0 +1,30 @@
+context("test-kendallTau")
+
+
+library("PlackettLuce")
+
+R <- matrix(c(1, 2, 3, 4,
+              1, 2, 3, 4,
+              1, 2, 3, 4,
+              1, 2, 3, 4,
+              1, 2, 3, 4,
+              1, 2, 3, 4), nrow = 6, byrow = TRUE)
+colnames(R) <- LETTERS[1:4]
+
+G <- group(as.rankings(R), 1:6)
+
+mod <- pltree(G ~ 1, data = G)
+
+preds <- predict(mod)
+
+
+
+test_that("kendall works", {
+  
+  k <- kendallTau(R, preds)
+  
+  k <- !any(as.vector(k) == c(1,9))
+  
+  expect_equal(k, FALSE)
+  
+})

### 2019-11-28 tests/testthat/test-pseudoR2.R

@@ -1,0 +1,57 @@
+context("test-pseudoR2")
+
+
+library("PlackettLuce")
+
+R <- matrix(c(1, 2, 3, 4,
+              1, 2, 3, 4,
+              1, 2, 3, 4,
+              1, 2, 3, 4,
+              1, 2, 3, 4,
+              1, 2, 3, 4), nrow = 6, byrow = TRUE)
+colnames(R) <- LETTERS[1:4]
+
+G <- group(as.rankings(R), 1:6)
+
+mod <- pltree(G ~ 1, data = G)
+
+test_that("pltree works", {
+  
+  p <- pseudoR2(mod)
+  
+  expect_equal(is.data.frame(p), TRUE)
+  
+})
+
+
+mod <- glm(Temp ~ Wind + Solar.R,
+           data = airquality,
+           family = poisson())
+
+pseudoR2(mod)
+
+test_that("default works", {
+  
+  p <- pseudoR2(mod)
+  
+  expect_equal(is.data.frame(p), TRUE)
+  
+})
+
+library("psychotree")
+
+test_that("bttree works", {
+  
+  pc <- rank_paircomp(G)
+  
+  df <- data.frame(x = rep(1, 6), y = rep(1, 6))
+  
+  df$pc <- pc
+  
+  mod <- bttree(pc ~ ., data = df)
+  
+  p <- pseudoR2(mod)
+  
+  expect_equal(is.data.frame(p), TRUE)
+  
+})

### 2020-01-06 README.Rmd
update-documentation

@@
 [![codecov](https://codecov.io/gh/agrobioinfoservices/gosset/master.svg)](https://codecov.io/github/agrobioinfoservices/gosset?branch=master)
+[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
 [![GPLv3 license](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.r-project.org/Licenses/GPL-3)
@@
 
+```r
+
 library("devtools")
-    devtools::install_github("agrobioinfoservices/gosset", upgrade = "never")
+devtools::install_github("agrobioinfoservices/gosset", build_vignettes = TRUE)
+
+```
+
+## Example
 
@@
 
-The full functionality of **gosset** is illustrated in the package vignette. The vignette can be found on the [package website](https://agrobioinfoservices.github.io/gosset/).
+The full functionality of **gosset** is illustrated in the package vignette.
+The vignette can be found on the [package website](https://agrobioinfoservices.github.io/gosset/) or from within `R` once the package has been installed, e.g.
+via
 
-## Contribution
+```r
 
-You are welcome to contribute to this project. Please read our [contribution guideline](CONTRIBUTING.md).
+vignette("Overview", package = "gosset")
 
-## Code of Conduct
+```
 
-Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.
+## Meta
+
+* Please [report any issues or bugs](https://github.com/agrobioinfoservices/gosset/issues).
+
+* License: GPL-3
+
+* Get citation information for *gosset* in R by typing
+`citation(package = "gosset")`.
+
+* You are welcome to contribute to the *gosset* project. Please read our [contribution guide lines](CONTRIBUTING.md).
+
+* Please note that the *gosset* project is released with a
+[Contributor Code of Conduct](CODE_OF_CONDUCT.md).
+By participating in the *gosset* project you agree to abide by its terms.

### 2020-01-06 R/crossvalidation.R

@@ -8,7 +8,7 @@
 #' of the form \eqn{y ~ x1 + ... + xn}
 #' @param data a data frame (or object coercible by as.data.frame to a data frame)
 #' containing the variables in the model
-#' @param k an integer for the number of splits in the cross-validation
+#' @param k an integer for the number of bins in the cross-validation
 #' @param folds an optional vector specifying the folds in the cross-validation
 #' @param mean.method a character for the method to calculate the mean of cross-validation estimators. 
 #' Options are: 'equal', arithmetic mean; 

### 2020-01-06 R/internal_functions.R

@@ -1,63 +1,73 @@
-# Find integer values in a vector with decimals
-# @param x a vector with numbers
-# @return a logical vector where TRUE indicates integer numbers
-# @examples
-# .is_wholenumber(c(1,3.1,4,0.9))
+#' Find integer values in a vector with decimals
+#' @param x a vector with numbers
+#' @return a logical vector where TRUE indicates integer numbers
+#' @examples
+#' .is_wholenumber(c(1,3.1,4,0.9))
+#' @noRd
 .is_wholenumber <- function(x) {
   x <- as.vector(t(x))
   x <- x %% 1 == 0
   return(x)
 }
 
-# Find a decimal number in a vector with integers
-# @param x a vector with numbers
-# @return a logical vector where TRUE indicates floating point numbers
-# @examples
-# .is_decimal(c(1,3.1,4,0.9))
+#' Find a decimal number in a vector with integers
+#' @param x a vector with numbers
+#' @return a logical vector where TRUE indicates floating point numbers
+#' @examples
+#' .is_decimal(c(1,3.1,4,0.9))
+#' @noRd
 .is_decimal <- function(x) {
   x <- as.vector(t(x))
   x <- !x %% 1 == 0
   return(x)
 }
 
-# Test a grouped_rankings object
-# @param object an object to have its class tested
-# @return a logical value where TRUE indicates that the object is of class "grouped_rankings"
-# @examples
-# .is_grouped_rankings(airquality)
+#' Is a grouped_rankings object
+#' @param object an object to have its class tested
+#' @return a logical value where TRUE indicates that the 
+#' object is of class "grouped_rankings"
+#' @examples
+#' .is_grouped_rankings(airquality)
+#' @noRd
 .is_grouped_rankings <- function(object) {
   
   c("grouped_rankings") %in% class(object)
   
 }
 
-# Test a rankings object
-# @param object an object to have its class tested
-# @return a logical value where TRUE indicates that the object is of class "rankings"
-# @examples
-# .is_rankings(airquality)
+#' Is a rankings object
+#' @param object an object to have its class tested
+#' @return a logical value where TRUE indicates that the 
+#' object is of class "rankings"
+#' @examples
+#' .is_rankings(airquality)
+#' @noRd
 .is_rankings <- function(object) {
   
   c("rankings") %in% class(object)
   
 }
 
[…]

### 2020-01-06 R/kendallTau.R

@@ -180,23 +180,24 @@ kendallTau.paircomp <- function(x, y, ...) {
 }
 
 
-# Kendall tau for a vector
-# 
-# Applies the "kendall" method from stats::cor with some 
-# previous treatment in the data, such as converting floating number into 
-# ranks (from the higher being the first and negative being the last) 
-# and removing zeros from incomplete ranks
-#
-# @param x a object of class numeric 
-# @param y a object of class numeric 
-# @param null.rm logical, to remove zeros from \code{x} and \code{y} 
-# @return The Kendall correlation coefficient and the Effective N
-# @examples
-# p1 <- c(1,2,3,4,5,6,7)
-# p2 <- c(1,2,0,3,5,7,6)
-# 
-# .get_kendall(p1, p2,  null.rm = TRUE)
-# .get_kendall(p1, p2, null.rm = FALSE)
+#' Kendall tau for a vector
+#' 
+#' Applies the "kendall" method from stats::cor with some 
+#' previous treatment in the data, such as converting floating number into 
+#' ranks (from the higher being the first and negative being the last) 
+#' and removing zeros from incomplete ranks
+#'
+#' @param x a object of class numeric 
+#' @param y a object of class numeric 
+#' @param null.rm logical, to remove zeros from \code{x} and \code{y} 
+#' @return The Kendall correlation coefficient and the Effective N
+#' @examples
+#' p1 <- c(1,2,3,4,5,6,7)
+#' p2 <- c(1,2,0,3,5,7,6)
+#' 
+#' .get_kendall(p1, p2,  null.rm = TRUE)
+#' .get_kendall(p1, p2, null.rm = FALSE)
+#' @noRd
 .get_kendall <- function(x, y, null.rm = TRUE) {
   
   keep <- !is.na(x) & !is.na(y)

### 2020-01-08 R/get_timeseries.R
fix-error-in-unlist-in-temperature

@@ -162,12 +162,19 @@
   
   # make a tibble
   if (n != 1) {
+    
+    dimnames(Y)[[2]] <- paste0("day", 1:ncol(Y))
+    
     Y <- tibble::as_tibble(Y[, 1:maxspan])
+    
   } else {
+    
     Y <- tibble::as_tibble(t(Y[, 1:maxspan]))
+    
+    names(Y) <- paste0("day", 1:ncol(Y))
   }
   
-  names(Y) <- paste0("day", 1:ncol(Y))
+  
   
   return(Y)
   

### 2020-01-08 R/rainfall.R

@@ -171,9 +171,9 @@ rainfall <- function(object, day.one = NULL,
     
     ind <- ind[order(ind$id), ]
     
-    ind$dates <- dates
+    ind$date <- dates
     
-    ind <- ind[, c("id", "dates", "index", "value")]
+    ind <- ind[, c("id", "date", "index", "value")]
     
     
   } 

### 2020-01-08 tests/testthat/test-temperature.R

@@ -5,11 +5,12 @@ library("gosset")
 load("../test_data.rda")
 
 
-maxDT <- c(33.5,34.9,32,27.5,32.8,27.4,32.6,31.6,32.3,28.5)
-minNT <- c(8.2,9.1,8.2,1.5,5.7,6,5.7,4.6,4.3,2.8)
+maxDT <- c(33.5, 34.9, 32, 27.5, 32.8, 27.4, 32.6, 31.6, 32.3, 28.5)
+minNT <- c(8.2, 9.1, 8.2, 1.5, 5.7, 6, 5.7, 4.6, 4.3, 2.8)
 
+# Test if the function is computing the right values
+# for maximum day temperature
 test_that("day equal", {
-  skip_on_cran()
   x <- temperature(object = temp, 
                    day.one = d,
                    span = 8)
@@ -19,8 +20,8 @@ test_that("day equal", {
   expect_equal(dt, TRUE)
 })
 
+# and for minimum day temperature
 test_that("night equal", {
-  skip_on_cran()
   x <- temperature(object = temp, 
                    day.one = d,
                    span = 8)
@@ -30,13 +31,32 @@ test_that("night equal", {
   expect_equal(nt, TRUE)
 })
 
+# also test if timeseries is working
+# and for minimum day temperature
+test_that("timeseries", {
+  
+  x <- temperature(object = temp, 
+                   day.one = d[1,],
+                   span = 10,
+                   timeseries = TRUE,
+                   intervals = 5)
+  
+  maxDT <- c(33.5, 33.5, 34.9, 34.9, 32, 32, 27.5, 
+             27.1, 32.8, 32.3, 27.6, 27.1, 32.8, 
+             32.3, 31.9, 31.1, 32.6, 31.7, 28.7, 28.2)
+  
+  x <- as.vector(t(x[x$index=="maxDT", "value"]))
+  
+  dt <- all.equal(maxDT, x)
+  
+  expect_equal(nt, TRUE)
+})
+
 test_that("nasapower works", {
   skip_on_cran()
-  r <- suppressWarnings(
-    temperature(object = lonlat, 
+  r <- temperature(object = lonlat, 
                    day.one = d,
                    span = 25)
-  )
   
   r <- as.vector(apply(r, 1, is.na))
   

### 2020-01-08 README.Rmd
fix-error-in-unlist-in-temperature
lifecycle-badge

@@
 [![GPLv3 license](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.r-project.org/Licenses/GPL-3)
+[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
 <!-- badges: end -->
@@
 ```r
-
 library("devtools")
 devtools::install_github("agrobioinfoservices/gosset", build_vignettes = TRUE)
-
 ```
@@
 ```r
-
 vignette("Overview", package = "gosset")
-
 ```

### 2022-02-26 R/crossvalidation.R
worth_map

@@ -61,7 +61,7 @@
 #'                       k = 3,
 #'                       seed = 999,
 #'                       family = poisson())
-#'                       
+#' }                    
 #' # Plackett-Luce Model
 #' if(require("PlackettLuce")) {
 #' # beans data from PlackettLuce

### 2022-02-26 R/gosset.R

@@ -1,4 +1,4 @@
-#' Tools for Analysis and Visualiation in Experimental Agriculture
+#' Tools for Data Analysis in Experimental Agriculture
 #'
 #' @name gosset
 #' @author Kauê de Sousa and Jacob van Etten 

### 2022-02-26 R/internal_functions.R

@@ -442,3 +442,21 @@
   return(x)
 }
 
+
+
+#' Color pallete
+#' @param x a integer
+#' @return the color pallete
+#' @noRd
+.col_pallet <- function(x, ...) {
+  
+  p <- c('#d73027','#4575b4', '#f46d43','#74add1', 
+         '#fdae61','#abd9e9', '#fee090', '#762a83',
+         '#a6dba0','#9970ab','#5aae61', '#c2a5cf', 
+         '#1b7837')
+  
+  v <- p[1:x]
+  
+  return(v)
+  
+}

### 2022-02-26 R/nodeinfo.R

@@ -256,7 +256,7 @@ plot.pltree <- function(x,
   # Extract ids from terminal nodes
   node_id <- partykit::nodeids(x, terminal = TRUE)
   
-  # get node information
+  # get models from each node
   nodes <- list()
   for (i in seq_along(node_id)) {
     nodes[[i]] <- x[[ node_id[i] ]]$node$info$object

### 2022-02-27 R/crossvalidation.R
AIC-deviance-method-from-David

@@ -20,8 +20,6 @@
 #' Options are: 'equal', arithmetic mean; 
 #' 'foldsize', weighted mean by the size in each fold; 
 #' 'Ztest' weighted through Z-test. See references 
-#' @param logLik.method a character for the method to compute logLik,
-#'  options are "tree" (default) or "worth"
 #' @param seed integer, the seed for random number generation. If \code{NULL} (the default), 
 #' \pkg{gosset} will set the seed randomly
 #' @param ... additional arguments passed the methods of the chosen model
@@ -94,7 +92,6 @@ crossvalidation <- function(formula,
                             k = 10,
                             folds = NULL, 
                             mean.method = "Ztest",
-                            logLik.method = "tree",
                             seed = NULL,
                             ...)
 {
@@ -161,14 +158,14 @@ crossvalidation <- function(formula,
   # split data into lists with training and test set
   # folds as numeric vector - gosset style
   if (!is.list(folds)) {
-  train <- list()
-  for (i in 1:k) {
-    train[[i]] <- data[folds != i ,]
-  }
-  
-  test <- list()
-  for (i in 1:k) {
-    test[[i]] <- data[folds == i  ,]
+    train <- list()
+    for (i in 1:k) {
+      train[[i]] <- data[folds != i ,]
+    }
+    
+    test <- list()
+    for (i in 1:k) {
+      test[[i]] <- data[folds == i  ,]
     }
   }
   
@@ -204,8 +201,8 @@ crossvalidation <- function(formula,
   # and the predictions of the test part of the data
   estimators <- try(mapply(function(X, Y) {
     a <- AIC(X, newdata = Y)
-    d <- deviance(X, newdata = Y, method = logLik.method)
-    p <- pseudoR2(X, newdata = Y, method = logLik.method)
+    d <- deviance(X, newdata = Y)
+    p <- pseudoR2(X, newdata = Y)
     data.frame(AIC = a, 
                deviance = d, p)
   }, X = mod, Y = test[]), silent = FALSE)
@@ -225,7 +222,7 @@ crossvalidation <- function(formula,
     try(predict(X, newdata = Y), silent = TRUE)
   }, X = model, Y = test[])
   
-  # if model is pltree take the kendall cor 
+  # if model is pltree take the kendall correlation
   if (model == "pltree") {
     
     # get the original rankings from each bin in test
[…]

### 2022-03-04 R/nodeinfo.R
webpage-vignette

@@ -239,6 +239,11 @@ plot.pltree <- function(x,
                         ref = NULL, 
                         ci.level = 0.95, ...){
   
+  
+  if (length(x) == 1) {
+    return(NextMethod(x, ...))
+  }
+  
   dots <- list(...)
   
   font.size <- dots[["font.size"]]

### 2022-06-14 inst/paper/example_01.R
example1

@@ -1,0 +1,24 @@
+library("PlackettLuce")
+library("gosset")
+library("climatrends")
+library("nasapower")
+
+data("nicabean", package = "gosset")
+
+dat <- nicabean$bean_rank
+
+covar <- nicabean$bean_covar
+
+
+traits <- unique(dat$trait)
+
+R <- list()
+
+for (i in seq_along(traits)) {
+  dat_i <- subset(dat, dat$trait == traits[i])
+  R[[i]] <- rank_numeric(data = dat_i,
+                         items = "item",
+                         input = "rank", 
+                         id = "id", 
+                         ascending = TRUE)
+}

### 2022-06-14 inst/paper/example_01.R
kendall-tau

@@ -1,24 +1,55 @@
-library("PlackettLuce")
 library("gosset")
+library("PlackettLuce")
 library("climatrends")
 library("nasapower")
 
 data("nicabean", package = "gosset")
 
-dat <- nicabean$bean_rank
+dat <- nicabean$trial
 
-covar <- nicabean$bean_covar
-
+covar <- nicabean$covar
 
 traits <- unique(dat$trait)
 
 R <- list()
 
 for (i in seq_along(traits)) {
+  
   dat_i <- subset(dat, dat$trait == traits[i])
+  
   R[[i]] <- rank_numeric(data = dat_i,
                          items = "item",
                          input = "rank", 
                          id = "id", 
                          ascending = TRUE)
 }
+
+
+# kendall 
+baseline <- which(grepl("OverallAppreciation", traits))
+
+kendall <- lapply(R[-baseline], function(X){
+  kendallTau(x = X, y = R[[baseline]])
+})
+
+kendall <- do.call("rbind", kendall)
+
+kendall$trait <- traits[-baseline]
+
+kendall
+
+# worth map
+
+mod <- lapply(R, PlackettLuce)
+
+worth_map(mod[-baseline],
+          labels = traits[-baseline], 
+          ref = "INTA Matagalpa") +
+  lab(x = "Variety",
+       y = "Trait")
+
+
+
+
+
+

### 2022-06-14 vignettes/Overview.R

@@ -1,27 +1,37 @@
-## ----fetch, message=FALSE, eval=FALSE, echo=TRUE------------------------------
-#  library("PlackettLuce")
-#  library("gosset")
-#  library("climatrends")
-#  library("nasapower")
-#  
-#  data("breadwheat", package = "gosset")
-#  
-#  head(breadwheat)
-#  
-#  dat <- breadwheat
-#  
+## ----fetch, message=FALSE, eval=TRUE, echo=TRUE-------------------------------
+library("gosset")
+library("PlackettLuce")
+library("climatrends")
+library("nasapower")
 
-## ---- message=FALSE, eval=FALSE, echo=TRUE------------------------------------
-#  traits <- c("yield",
-#              "grainquality",
-#              "germination")
-#  
-#  # names of colunms with varieties
-#  items <- paste0("variety_", letters[1:3])
-#  
-#  # name of varieties
-#  itemnames <- sort(unique(unlist(dat[items])))
-#  
+data("nicabean", package = "gosset")
+
+dat <- nicabean$trial
+
+covar <- nicabean$covar
+
+head(dat)
+
+
+## ---- message=FALSE, eval=TRUE, echo=TRUE-------------------------------------
+
+traits <- unique(dat$trait)
+
+R <- list()
+
+for (i in seq_along(traits)) {
+  
+  dat_i <- subset(dat, dat$trait == traits[i])
+  
+  R[[i]] <- rank_numeric(data = dat_i,
+                         items = "item",
+                         input = "rank", 
+                         id = "id", 
+                         ascending = TRUE)
+}
+
+head(R[[1]])
+
 
 ## ---- message=FALSE, eval=FALSE, echo=TRUE------------------------------------
 #  # build the rankings and put into a list

### 2022-06-20 inst/paper/paper.Rmd
update manuscript

---
title: 'gosset: An R package for analysis and synthesis of ranking data in agricultural
  experimentation'
abstract: null
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
bibliography: paper.bib
header-includes: \usepackage{caption} \captionsetup[figure]{labelformat=empty}
---

>Kauê de Sousa^1,2[*]^, Jacob van Etten^2^, David Brown^3,4^, Jonathan Steinke^2,5^  
>-
>
^1^ Department of Agricultural Sciences, Inland Norway University of Applied Sciences, 2318 Hamar, Norway  
^2^ Digital Inclusion, Bioversity International, Parc Scientifique Agropolis II, 34397, Montpellier Cedex 5, France   
^3^ Laboratory of Geo-Information Science and Remote Sensing, Wageningen University & Research, Droevendaalsesteeg 3, 6708 PB, Wageningen, The Netherlands   
^4^ Digital Inclusion, Bioversity International, 30501, Turrialba, Costa Rica   
^5^ Humboldt University Berlin, Berlin, Germany

^[*]^ Correspondence should be addressed to: <kaue.desousa@inn.no>

# Abstract
Appropriate data management and analysis are necessary to produce practical information from agricultural experimentation data. There is also an ongoing trend advocating for programmatic tools that supports reproducible workflows in scientific research. We developed the R package gosset, providing functionality to support analysis workflows with rank-based models, such as Plackett-Luce and Bradley-Terry. The gosset package facilitates data preparation, modelling and results presentation stages. We demonstrate the functionality of the package with a case of on-farm evaluations of common bean (*Phaseolus vulgaris* L.) genotypes in Nicaragua.

::: frontmatter
::: keyword
data-driven agriculture, Plackett-Luce model
:::
:::

\newpage

# Required Metadata {#required-metadata .unnumbered}

# Current code version {#current-code-version .unnumbered}

| **Nr.** | **Code metadata description**                                   | **Please fill in this column**                             |
|:-----------------|:--------------------------|:-------------------------|
| C1      | Current code version                                            | 0.4.003                                                    |
| C2      | Permanent link to code/repository used for this code version    | <https://github.com/AgrDataSci/gosset>                     |
| C3      | Code Ocean compute capsule                                      |                                                            |
| C4      | Legal Code License                                              | MIT                                                        |
| C5      | Code versioning system used                                     | Git                                                        |
| C6      | Software code languages, tools, and services used               | R                                                          |
| C7      | Compilation requirements, operating environments & dependencies |                                                            |
| C8      | If available Link to developer documentation/manual             | <https://agrdatasci.github.io/gosset/>                     |
| C9      | Support email for questions                                     | [kaue.desousa@inn.no](mailto:kaue.desousa@inn.no){.email} |

: *Code metadata (mandatory)*

# Motivation and significance

*Introduce the scientific background and the motivation for developing
the software.*

*Explain why the software is important, and describe the exact
(scientific) problem(s) it solves.*

*Indicate in what way the software has contributed (or how it will
contribute in the future) to the process of scientific discovery; if
available, this is to be supported by citing a research paper using the
software.*

*Provide a description of the experimental setting (how does the user
use the software?).*

*Introduce related work in literature (cite or list algorithms used,
other software etc.).*

Participatory experimentation approaches have been increasingly applied in agricultural
research [@deRoo2019farm]. While collecting data in ranking format is uncommon in general
agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA].
Recently developed approaches for on-farm experimentation,
such as the tricot methodology are based on the collection of data in ranking format [@vanetten_beza_2019].
On the other hand, newly proposed approaches for synthesis of crop variety evaluation data 
largely depend on the analysis of ranking data [@brown2020data].
The analysis of ranking data requires the use of appropriate statistical models such as Plackett-Luce  [@luce_individual_1959;@Plackett] or Bradley-Terry [@bradley1952rank]. Functionality for fitting Bradley-Terry and 
Plackett-Luce models are available in R with the packages BradleyTerry2 and PlackettLuce
respectively [@BradleyTerry2; @Turner2020]. However, extended functionality was required for the entire 
data science workflow, which usually includes: (1) Data preparation and cleaning, 
(2) modelling and validation, and (3) results presentation. For (1) gosset provides functions for 
converting and preparing data into ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2
respectively. For (2), gosset provides functions for model selection and validation using cross-validation. In the 
case of (3), enhanced functionality for plotting model results is provided by the gosset package.

# Software description

*Describe the software in as much as is necessary to establish a
vocabulary needed to explain its impact.*

The R package gosset provides functionality supporting the analysis
workflows in agricultural experimentation, especially with rank-based
approaches. The package is available in The Comprehensive R Archive Network (CRAN)
and can be installed by executing ``install.packages("gosset")``.

## Software Architecture

*Give a short overview of the overall software architecture; provide a
pictorial component overview or similar (if possible). If necessary
provide implementation details.*

The R package gosset is structured following the guidelines described in the manual
for creating R add-on packages [@r_extensions_2022]. This structure basically consist of 
files DESCRTIPTION, LICENSE, NAMESPACE and NEWS, and directories data, dev, docs, inst,
man, R, and vignettes. The package functions were developed following the S3 methods
style [@r_extensions_2022] and are contained in the R sub-directory.

## Software Functionalities

*Present the major functionalities of the software.*

**Data management and preparation**

When data from agricultural experiments is not in ranking format, it
should be transformed to be used as inputs into R packages for the
analysis of ranking data. For instance, the Plackett-Luce model [@luce_individual_1959;@Plackett]
is implemented in the R package as PlackettLuce,
which requires the data to be formatted as ranking matrix. Another
example is the Bradley-Terry model [@bradley1952rank], implemented
in the package BradleyTerry2 [@BradleyTerry2] and which requires the
input data to be formatted as paired comparisons. For these cases,
gosset provides the functions ``rank_numeric`` and ``rank_binomial`` respectively. The
function ``rank_numeric`` transforms a set of numeric values into an ordinal
ranking, considering if items with higher numeric values should be ranked first or
last. The function ``rank_binomial`` transforms data in rankings format into
pairwise comparisons, as required by the package BradleyTerry2.
Additionally, gosset provides the function ``rank_tricot``, for the case
when the experimental data is collected from tricot trials. In those trials,
farmers rank the technology under evaluation (e.g., crop variety) stating which
is the best and which is the worst from a set of three [@vanetten_beza_2019].

**Modelling**

The gosset package provides the following functions to support model
selection and validation of Bradley-Terry and Plackett-Luce models. The function ``pseudoR2``
computes goodness-of-fit measure McFadden's pseudo-R2 [@mcfadden1973conditional]. The ``AIC``
function computes the Akaike Information Criterion[@akaike1974]. The ``kendallTau`` function
computes the Kendall-tau rank correlation coefficient [@kendall_1938].

**Visualization and results presentation**

``plot``

``worth_map``

``worth_map``

``regret``

``reliability``

# Illustrative Examples
We demonstrate the functionality of the gosset package using the 
trial dataset "nicabean", which consists of trial data collected in
Nicaragua following the tricot approach. We use the Plackett-Luce model
implemented in the R package *PlackettLuce*[@Turner2020]. We use climate data
as model covariates to investigate the effect fo climate factors on the performance 
of common bean (*Phaseolus vulgaris* L.) genotypes. For obtaining the climate data,
we use the *nasapower* package[@nasapower]. Climatic indices were computed with the
*climatrends* package[@climatrendspkg].

First, the required packages and data are loaded.
``` {r}
library("gosset")
library("PlackettLuce")
library("climatrends")
library("nasapower")

#data("nicabean", package = "gosset")

load("D:/GitHub/gosset/data/nicabean.rda")

dat <- nicabean$trial

covar <- nicabean$covar

traits <- unique(dat$trait)

```
Make a PlackettLuce rank using the function ``rank_numeric``
```{r}
R <- vector(mode = "list", length = length(traits))

for (i in seq_along(traits)) {

  dat_i <- subset(dat, dat$trait == traits[i])

  R[[i]] <- rank_numeric(data = dat_i,
                         items = "item",
                         input = "rank", 
                         id = "id", 
                         ascending = TRUE)
}

```

```{r}
head(R[[1]], 10)
```
Kendall correlation between traits using kendallTau()
```

```

Worth map 
```

```

Plackett-Luce tree using environmental data 
```

```

Model selection using crossvalidation
```

```

pseudoR2

```r

```

``regret``
```r

```

 ``reliability``

\newpage

# Impact

***This is the main section of the article and the reviewers weight the
description here appropriately***

*Indicate in what way new research questions can be pursued as a result
of the software (if any).*

*Indicate in what way, and to what extent, the pursuit of existing
research questions is improved (if so).*

*Indicate in what way the software has changed the daily practice of its
users (if so).*

*Indicate how widespread the use of the software is within and outside
the intended user group.*

*Indicate in what way the software is used in commercial settings and/or
how it led to the creation of spin-off companies (if so).*

Reproducible and efficient workflows are fundamental in scientific research (Lowndes et al. 2017).
The gosset package provides functionality that was not previously available from 
other R packages and which enabled scientific studies based on the analysis of
ranking data. This functionality enables making the entire workflow to be reproducible
and more efficient. The utility of the gosset package has been demonstrated by enabling 
studies based on the analysis of ranking data. For instance, both @vanEtten2019 and  @deSousa2021
applied the Plackett-Luce model in combination with recursive partitioning [@Turner2020;@Zeileis2008].
In both studies, the gosset package supported the data preparation, model validation and results presentation
tasks. The gosset package is freely available and can be downloaded from 
CRAN https://cran.r-project.org/package=gosset.

# Conclusions

*Set out the conclusion of this original software publication.*

We described the functionality of the R package gosset to support 
the synthesis and analysis of ranking data. The package provide functions not available in
existing R packages for analyzing ranking data. We provided an illustrative example 
covering the main functionality across the stages involved in the analysis workflow.

# Conflict of Interest

No conflict of interest exists: We wish to confirm that there are no
known conflicts of interest associated with this publication and there
has been no significant financial support for this work that could have
influenced its outcome.

# Acknowledgements {#acknowledgements .unnumbered}

We acknowledge Olga Spellman (Science Writing Service of the Alliance of
Bioversity International and CIAT) for English editing of this
manuscript
\newpage

# References

**Zenodo repository for the gosset package:** <https://doi.org/10.5281/zenodo.6339989>

::: thebibliography
:::

*Please add the reference to the software repository if DOI for software
is available.*

### 2022-06-20 inst/paper/paper.Rmd
update manuscript

@@
 output:
+  pdf_document: default
   word_document: default
-  pdf_document: default
   html_document:
@@
 
->Kauê de Sousa^1,2[*]^, Jacob van Etten^2^, David Brown^3,4^, Jonathan Steinke^2,5^  
+>Kauê de Sousa^1,2[*]^, David Brown^3,4^, Jonathan Steinke^2,5^, Jacob van Etten^2^  
 >-
@@
 ::: keyword
-data-driven agriculture, Plackett-Luce model
+**Keywords**: Bradley-Terry, Plackett-Luce, data science, on-farm trials, participatory-research, tricot
 :::
@@
 approaches. The package is available in The Comprehensive R Archive Network (CRAN)
-and can be installed by executing ``install.packages("gosset")``.
+and can be installed by executing ``install.packages("gosset")``. The package is named 
+in honor to William Sealy Gosset, known by the pen name ‘Student’, a pioneer of modern statistics
+in small sample experimental design and analysis [@Ziliak_2019].
 
@@
 
-When data from agricultural experiments is not in ranking format, it
-should be transformed to be used as inputs into R packages for the
-analysis of ranking data. For instance, the Plackett-Luce model [@luce_individual_1959;@Plackett]
-is implemented in the R package as PlackettLuce,
-which requires the data to be formatted as ranking matrix. Another
-example is the Bradley-Terry model [@bradley1952rank], implemented
-in the package BradleyTerry2 [@BradleyTerry2] and which requires the
-input data to be formatted as paired comparisons. For these cases,
-gosset provides the functions ``rank_numeric`` and ``rank_binomial`` respectively. The
-function ``rank_numeric`` transforms a set of numeric values into an ordinal
-ranking, considering if items with higher numeric values should be ranked first or
-last. The function ``rank_binomial`` transforms data in rankings format into
-pairwise comparisons, as required by the package BradleyTerry2.
-Additionally, gosset provides the function ``rank_tricot``, for the case
-when the experimental data is collected from tricot trials. In those trials,
-farmers rank the technology under evaluation (e.g., crop variety) stating which
-is the best and which is the worst from a set of three [@vanetten_beza_2019].
+*   **``rank_binomial``** transforms a ranking object into a binary comparisons, as required
+by package BradleyTerry2 [@BradleyTerry2].  
+*   **``rank_numeric``** converts numeric values into rankings. The parameter ``ascending``
+indicates if the rankings should be made considering the numeric values in ascending order. 
+The default is ``asceding = FALSE``. This function is useful when the data were collected 
+as numerical observations, for instance, in a experiment measuring crop yield.  
+*   **``rank_tricot``** transforms data in tricot format into rankings.
 
@@
 
-The gosset package provides the following functions to support model
-selection and validation of Bradley-Terry and Plackett-Luce models. The function ``pseudoR2``
-computes goodness-of-fit measure McFadden's pseudo-R2 [@mcfadden1973conditional]. The ``AIC``
-function computes the Akaike Information Criterion[@akaike1974]. The ``kendallTau`` function
-computes the Kendall-tau rank correlation coefficient [@kendall_1938].
+*   **``AIC``** computes the Akaike Information Criterion @akaike1974] for a Bradley-Terry model fitted with
+BradleyTerry2 [@BradleyTerry2] or a Plackett-Luce model fitted with [@Turner2020].  
+*   **``btpermute``** deviance-based forward variable selection [@lysen2009permuted] procedure for Bradley-Terry models.  
+*   **``compare``** compare the agreenment between methods.  
+*   **``crossvalidation``** performs k-fold cross-validation, where k could be specified by the user.The default is
+10-fold. Folds can be provided as a vector for a custom cross-validation, such as blocked cross-validation.
+*   **``forward``** executes forward variable selection with cross-validation.  
+*   **``kendallTau``** computes the Kendall-tau rank correlation coefficient between two rankings [@kendall_1938].  
+*   **``kendallW``** computes Kendall's W (coefficient of concordance) among observed rankings and those predicted by
+the Plackett-Luce model.  
+*   **``pseudoR2``** computes goodness-of-fit measure McFadden's pseudo-R2 [@mcfadden1973conditional].
 
@@
 
-``plot``
+*   **``plot``** provides an alternative to the default plot function of the PlackettLuce package.
 
-``worth_map``
+*   **``regret``**
 
-``worth_map``
+*   **``reliability``**
 
-``regret``
+*   **``worth_bar``**
 
-``reliability``
+*   **``worth_map``**
 
@@
 trial dataset "nicabean", which consists of trial data collected in
-Nicaragua following the tricot approach. We use the Plackett-Luce model
-implemented in the R package *PlackettLuce*[@Turner2020]. We use climate data
-as model covariates to investigate the effect fo climate factors on the performance 
-of common bean (*Phaseolus vulgaris* L.) genotypes. For obtaining the climate data,
-we use the *nasapower* package[@nasapower]. Climatic indices were computed with the
-*climatrends* package[@climatrendspkg].
+Nicaragua.
+In this example we present a suggested workflow to assess trait prioritization and crop performance using decentralized on-farm data generated with the tricot approach [@vanetten_beza_2019].
 
-First, the required packages and data are loaded.
-``` {r}
+We use the `nicabean` dataset, which was generated with decentralized on-farm trials of common bean (*Phaseolus vulgaris* L.) varieties in Nicaragua over five seasons (between 2015 and 2016).
+Following the tricot approach, farmers were asked to test in their farms three varieties of common bean.
+The varieties were randomly assigned as incomplete blocks of size three (out of 10 varieties).
+The farmers assessed which of the three varieties has the best and worst performance in nine traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, taste) and overall appreciation).
+The farmers also provided their overall appreciation about the varieties, i.e., which
+variety has the best and the worst performance based on the overall performance considering all the traits.
+
+We use the Plackett-Luce model implemented in the R package *PlackettLuce*[@Turner2020]. 
+To consider the effecto of climate on the variety performance, we use climate data
+For obtaining the climate data, we used the *nasapower* package[@nasapower]. 
+Climatic indices were computed with the *climatrends* package[@climatrendspkg].
+
+First, we load the required packages and data. The ``nicabean`` dataset contains the 
+tricot rankings (``trial``), the climate covariates (``covar``) and the traits observations (``traits``).
+``` {r fetch, message=FALSE, eval=TRUE, echo=TRUE}
 library("gosset")
@@
 
-#data("nicabean", package = "gosset")
-
-load("D:/GitHub/gosset/data/nicabean.rda")
+data("nicabean", package = "gosset")
 
@@
 ```
-Make a PlackettLuce rank using the function ``rank_numeric``
+Make a PlackettLuce rank using the function ``rank_numeric``.
 ```{r}
@@
 ```{r}
-head(R[[1]], 10)
+head(R[[1]], 5)
 ```
+
 Kendall correlation between traits using kendallTau()

### 2022-06-21 inst/paper/paper.Rmd
Update paper.Rmd

@@
 # Illustrative Examples
-We demonstrate the functionality of the gosset package using the 
-trial dataset "nicabean", which consists of trial data collected in
-Nicaragua.
-In this example we present a suggested workflow to assess trait prioritization and crop performance using decentralized on-farm data generated with the tricot approach [@vanetten_beza_2019].
+To demonstrate the functionality of the gosset package, we use the `nicabean` dataset,
+which was generated with decentralized on-farm trials of common bean (*Phaseolus vulgaris* L.)
+varieties in Nicaragua over five seasons (between 2015 and 2016). Following the tricot 
+approach [@vanetten_beza_2019], farmers were asked to test in their farms three varieties
+of common bean. The varieties were randomly assigned as incomplete blocks of size three 
+(out of 10 varieties). The farmers assessed which of the three varieties has the best and
+worst performance in nine traits (vigor, architecture, resistance to pests, resistance to 
+diseases, tolerance to drought, yield, marketability, taste) and overall appreciation). 
+The farmers also provided their overall appreciation about the varieties, i.e., which
+variety has the best and the worst performance based on the overall performance considering
+all the traits.
 
-We use the `nicabean` dataset, which was generated with decentralized on-farm trials of common bean (*Phaseolus vulgaris* L.) varieties in Nicaragua over five seasons (between 2015 and 2016).
-Following the tricot approach, farmers were asked to test in their farms three varieties of common bean.
-The varieties were randomly assigned as incomplete blocks of size three (out of 10 varieties).
-The farmers assessed which of the three varieties has the best and worst performance in nine traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, taste) and overall appreciation).
-The farmers also provided their overall appreciation about the varieties, i.e., which
-variety has the best and the worst performance based on the overall performance considering all the traits.
-
-We use the Plackett-Luce model implemented in the R package *PlackettLuce*[@Turner2020]. 
-To consider the effecto of climate on the variety performance, we use climate data
-For obtaining the climate data, we used the *nasapower* package[@nasapower]. 
-Climatic indices were computed with the *climatrends* package[@climatrendspkg].
+We used the Plackett-Luce model implemented in the R package *PlackettLuce* [@Turner2020]. 
+To consider the effect of climate on the variety performance, we used climate data
+For obtaining the climate data, we used the *nasapower* package [@nasapower]. 
+Climatic indices were computed with the *climatrends* package [@climatrendspkg].
 
@@
 tricot rankings (``trial``), the climate covariates (``covar``) and the traits observations (``traits``).
-``` {r fetch, message=FALSE, eval=TRUE, echo=TRUE}
+``` {r, message = FALSE, eval = TRUE, echo = TRUE}
 library("gosset")
@@
 library("nasapower")
+library("ggplot2")
 
@@
 Make a PlackettLuce rank using the function ``rank_numeric``.
-```{r}
+```{r, message = FALSE, eval = TRUE, echo = TRUE}
 R <- vector(mode = "list", length = length(traits))
@@
 ```
-
-```{r}
+The following lines show the first 10 rankings in the format 
+used by the PlackettLuce package.  
+```{r fetch, message = FALSE, eval = TRUE, echo = TRUE}
 head(R[[1]], 5)
 ```
+The correlation among rankings can be assessed using the function ``kendallTau``.  
+```{r}
+baseline <- which(grepl("OverallAppreciation", traits))
 
-Kendall correlation between traits using kendallTau()
+kendall <- lapply(R[-baseline], function(X){
+  kendallTau(x = X, y = R[[baseline]])
+})
+
+kendall <- do.call("rbind", kendall)
+
+kendall$trait <- traits[-baseline]
+
+print(kendall)
 ```
+
+We fit a series of Plackett-Luce models using the function ``PlackettLuce``
+from the package of the same name as input in the lapply function.
+
+```{r, message=FALSE, eval=TRUE, echo=TRUE}
+
+mod <- lapply(R, PlackettLuce)
 
@@
 
-Worth map 
-```
+The ``worth_map`` function can be used to visually assess variety performance
+based on different traits. The values represented in a worth_map are worth estimates
+in log scale.
 
+```{r, message=FALSE, eval=TRUE, echo=TRUE}
+worth_map(mod[-baseline],
+          labels = traits[-baseline], 
+          ref = "Amadeus 77") +
+  labs(x = "Variety",
+       y = "Trait")
 ```
@@
 and more efficient. The utility of the gosset package has been demonstrated by enabling 
-studies based on the analysis of ranking data. For instance, both @vanEtten2019 and  @deSousa2021
+studies based on the analysis of ranking data. For instance, @vanEtten2019, @Moyo_2021  and @deSousa2021
 applied the Plackett-Luce model in combination with recursive partitioning [@Turner2020;@Zeileis2008].
-In both studies, the gosset package supported the data preparation, model validation and results presentation
-tasks. The gosset package is freely available and can be downloaded from 
-CRAN https://cran.r-project.org/package=gosset.
+In these studies, the gosset package supported data preparation, model validation and results presentation
+tasks.
 

### 2022-06-21 inst/paper/paper.Rmd
update-rmd-paragraphs

@@
 ---
-title: 'gosset: An R package for analysis and synthesis of ranking data in agricultural
-  experimentation'
+title: '`gosset`: An R package for analysis and synthesis of ranking data in agricultural experimentation'
 abstract: null
@@
 
-*Introduce the scientific background and the motivation for developing
-the software.*
+*Introduce the scientific background and the motivation for developing the software.*
 
-*Explain why the software is important, and describe the exact
-(scientific) problem(s) it solves.*
+*Explain why the software is important, and describe the exact (scientific) problem(s) it solves.*
 
-*Indicate in what way the software has contributed (or how it will
-contribute in the future) to the process of scientific discovery; if
-available, this is to be supported by citing a research paper using the
-software.*
+*Indicate in what way the software has contributed (or how it will contribute in the future) to the process of scientific discovery; if available, this is to be supported by citing a research paper using the software.*
 
-*Provide a description of the experimental setting (how does the user
-use the software?).*
+*Provide a description of the experimental setting (how does the user use the software?).*
 
-*Introduce related work in literature (cite or list algorithms used,
-other software etc.).*
+*Introduce related work in literature (cite or list algorithms used, other software etc.).*
 
-Participatory experimentation approaches have been increasingly applied in agricultural
-research [@deRoo2019farm]. While collecting data in ranking format is uncommon in general
-agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA].
-Recently developed approaches for on-farm experimentation,
+Participatory experimentation approaches have been increasingly applied in agricultural research [@deRoo2019farm]. While collecting data in ranking format is uncommon in general
+agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA]. Recently developed approaches for on-farm experimentation,
 such as the tricot methodology are based on the collection of data in ranking format [@vanetten_beza_2019].
-On the other hand, newly proposed approaches for synthesis of crop variety evaluation data 
-largely depend on the analysis of ranking data [@brown2020data].
-The analysis of ranking data requires the use of appropriate statistical models such as Plackett-Luce  [@luce_individual_1959;@Plackett] or Bradley-Terry [@bradley1952rank].
-Functionality for fitting Bradley-Terry and
-Plackett-Luce models are available in R with the packages BradleyTerry2 and PlackettLuce
-respectively [@BradleyTerry2; @Turner2020]. However, extended functionality was required for the entire 
-data science workflow, which usually includes: (1) Data preparation and cleaning, 
-(2) modelling and validation, and (3) results presentation. For (1) gosset provides functions for 
-converting and preparing data into ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2
-respectively. For (2), gosset provides functions for model selection and validation using cross-validation. In the 
-case of (3), enhanced functionality for plotting model results is provided by the gosset package.
+On the other hand, newly proposed approaches for synthesis of crop variety evaluation data largely depend on the analysis of ranking data [@brown2020data].
+The analysis of ranking data requires the use of appropriate statistical models such as Plackett-Luce [@luce_individual_1959; @Plackett] or Bradley-Terry [@bradley1952rank].
+Functionality for fitting Bradley-Terry and Plackett-Luce models are available in R with the packages BradleyTerry2 and PlackettLuce respectively [@BradleyTerry2; @Turner2020].
+However, extended functionality was required for the entire data science workflow, which usually includes: (1) Data preparation and cleaning, (2) modelling and validation, and (3) results presentation.
+For (1) gosset provides functions for converting and preparing data into ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2
+respectively.
+For (2), gosset provides functions for model selection and validation using cross-validation.
+In the case of (3), enhanced functionality for plotting model results is provided by the gosset package.
 
@@
 
-*Describe the software in as much as is necessary to establish a
-vocabulary needed to explain its impact.*
+*Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.*
 
-The R package gosset provides functionality supporting the analysis
-workflows in agricultural experimentation, especially with rank-based
-approaches. The package is available in The Comprehensive R Archive Network (CRAN)
-and can be installed by executing ``install.packages("gosset")``. The package is named 
-in honor to William Sealy Gosset, known by the pen name ‘Student’, a pioneer of modern statistics
-in small sample experimental design and analysis [@Ziliak_2019].
+The R package gosset provides functionality supporting the analysis workflows in agricultural experimentation, especially with rank-based approaches.
+The package is available in The Comprehensive R Archive Network (CRAN) and can be installed by executing `install.packages("gosset")`.
+The package is named in honor to William Sealy Gosset, known by the pen name ‘Student’, a pioneer of modern statistics in small sample experimental design and analysis [@Ziliak_2019].
 
@@
 
-*Give a short overview of the overall software architecture; provide a
-pictorial component overview or similar (if possible). If necessary
-provide implementation details.*
+*Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.*
 
-The R package gosset is structured following the guidelines described in the manual
-for creating R add-on packages [@r_extensions_2022]. This structure basically consist of 
-files DESCRTIPTION, LICENSE, NAMESPACE and NEWS, and directories data, dev, docs, inst,
-man, R, and vignettes. The package functions were developed following the S3 methods
-style [@r_extensions_2022] and are contained in the R sub-directory.
+The R package gosset is structured following the guidelines described in the manual for creating R add-on packages [@r_extensions_2022].
+This structure basically consist of files DESCRTIPTION, LICENSE, NAMESPACE and NEWS, and directories data, dev, docs, inst, man, R, and vignettes.
+The package functions were developed following the S3 methods style [@r_extensions_2022] and are contained in the R sub-directory.
 
@@
 
-**Data management and preparation**
+### Data management and preparation
 
-*   **``rank_binomial``** transforms a ranking object into a binary comparisons, as required
-by package BradleyTerry2 [@BradleyTerry2].  
-*   **``rank_numeric``** converts numeric values into rankings. The parameter ``ascending``
-indicates if the rankings should be made considering the numeric values in ascending order. 
-The default is ``asceding = FALSE``. This function is useful when the data were collected 
-as numerical observations, for instance, in a experiment measuring crop yield.  
-*   **``rank_tricot``** transforms data in tricot format into rankings.
+* `rank_binomial` transforms a ranking object into a binary comparisons, as required by package BradleyTerry2 [@BradleyTerry2].  
+* `rank_numeric` converts numeric values into rankings.
+The parameter `ascending = ` indicates if the rankings should be made considering the numeric values in ascending order.
+The default is `asceding = FALSE`.
+This function is useful when the data were collected as numerical observations, for instance, in a experiment measuring crop yield.
+* `rank_tricot` transforms data in tricot format into rankings.
 
-**Modelling**
+### Modelling
 
-*   **``AIC``** computes the Akaike Information Criterion @akaike1974] for a Bradley-Terry model fitted with
-BradleyTerry2 [@BradleyTerry2] or a Plackett-Luce model fitted with [@Turner2020].  
-*   **``btpermute``** deviance-based forward variable selection [@lysen2009permuted] procedure for Bradley-Terry models.  
-*   **``compare``** compare the agreenment between methods.  
-*   **``crossvalidation``** performs k-fold cross-validation, where k could be specified by the user.The default is
-10-fold. Folds can be provided as a vector for a custom cross-validation, such as blocked cross-validation.
-*   **``forward``** executes forward variable selection with cross-validation.  
-*   **``kendallTau``** computes the Kendall-tau rank correlation coefficient between two rankings [@kendall_1938].  
-*   **``kendallW``** computes Kendall's W (coefficient of concordance) among observed rankings and those predicted by
-the Plackett-Luce model.  
-*   **``pseudoR2``** computes goodness-of-fit measure McFadden's pseudo-R2 [@mcfadden1973conditional].
+* `AIC` computes the Akaike Information Criterion @akaike1974] for a Bradley-Terry model fitted with BradleyTerry2 [@BradleyTerry2] or a Plackett-Luce model fitted with [@Turner2020].  
+* `btpermute` deviance-based forward variable selection [@lysen2009permuted] procedure for Bradley-Terry models.
+* `compare` compare the agreenment between methods.  
+* `crossvalidation` performs k-fold cross-validation, where k could be specified by the user.The default is 10-fold.
+Folds can be provided as a vector for a custom cross-validation, such as blocked cross-validation.
+* `forward` executes forward variable selection with cross-validation.  
+* `kendallTau` computes the Kendall-tau rank correlation coefficient between two rankings [@kendall_1938].  
+* `kendallW` computes Kendall's W (coefficient of concordance) among observed rankings and those predicted by the Plackett-Luce model.  
+*   `pseudoR2` computes goodness-of-fit measure McFadden's pseudo-R2 [@mcfadden1973conditional].
 
-**Visualization and results presentation**
+### Visualization and results presentation
 
-*   **``plot``** provides an alternative to the default plot function of the PlackettLuce package.
+* `plot` provides an alternative to the default plot function of the PlackettLuce package.
 
-*   **``regret``**
+* `regret`
 
-*   **``reliability``**
+* `reliability`
 
-*   **``worth_bar``**
+* `worth_bar`
 
-*   **``worth_map``**
+* `worth_map`
 
 # Illustrative Examples
-To demonstrate the functionality of the gosset package, we use the `nicabean` dataset,
-which was generated with decentralized on-farm trials of common bean (*Phaseolus vulgaris* L.)
-varieties in Nicaragua over five seasons (between 2015 and 2016). Following the tricot 
-approach [@vanetten_beza_2019], farmers were asked to test in their farms three varieties
-of common bean. The varieties were randomly assigned as incomplete blocks of size three 
-(out of 10 varieties). The farmers assessed which of the three varieties has the best and
-worst performance in nine traits (vigor, architecture, resistance to pests, resistance to 
-diseases, tolerance to drought, yield, marketability, taste) and overall appreciation). 
-The farmers also provided their overall appreciation about the varieties, i.e., which
-variety has the best and the worst performance based on the overall performance considering
-all the traits.
+
+To demonstrate the functionality of the gosset package, we use the `nicabean` dataset, which was generated with decentralized on-farm trials of common bean (*Phaseolus vulgaris* L.) varieties in Nicaragua over five seasons (between 2015 and 2016).
+Following the tricot  approach [@vanetten_beza_2019], farmers were asked to test in their farms three varieties of common bean.
+The varieties were randomly assigned as incomplete blocks of size three (out of 10 varieties).
+The farmers assessed which of the three varieties has the best and worst performance in nine traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, taste) and overall appreciation).
+The farmers also provided their overall appreciation about the varieties, i.e., which variety has the best and the worst performance based on the overall performance considering all the traits.
 
 We used the Plackett-Luce model implemented in the R package *PlackettLuce* [@Turner2020].
-To consider the effect of climate on the variety performance, we used climate data
+To consider the effect of climate on the variety performance, we used climate data.
 For obtaining the climate data, we used the *nasapower* package [@nasapower].
@@
 
-First, we load the required packages and data. The ``nicabean`` dataset contains the 
-tricot rankings (``trial``), the climate covariates (``covar``) and the traits observations (``traits``).
+First, we load the required packages and data. The `nicabean` dataset contains the tricot rankings (`trial`), the climate covariates (`covar`) and the traits observations (`traits`).
+
 ``` {r, message = FALSE, eval = TRUE, echo = TRUE}
@@
 ```
-Make a PlackettLuce rank using the function ``rank_numeric``.
+
+Make a PlackettLuce rank using the function `rank_numeric`.
+
 ```{r, message = FALSE, eval = TRUE, echo = TRUE}
@@
 ```
-The following lines show the first 10 rankings in the format 
-used by the PlackettLuce package.  
+
+The following lines show the first 10 rankings in the format used by the PlackettLuce package.
+
 ```{r fetch, message = FALSE, eval = TRUE, echo = TRUE}
@@
 ```
-The correlation among rankings can be assessed using the function ``kendallTau``.  
+
+The correlation among rankings can be assessed using the function `kendallTau`.
+
 ```{r}
@@
 
-We fit a series of Plackett-Luce models using the function ``PlackettLuce``
-from the package of the same name as input in the lapply function.
+We fit a series of Plackett-Luce models using the function `PlackettLuce` from the package of the same name as input in the lapply function.
 
@@
 
-The ``worth_map`` function can be used to visually assess variety performance
-based on different traits. The values represented in a worth_map are worth estimates
-in log scale.
+The `worth_map` function can be used to visually assess variety performance based on different traits. The values represented in a worth_map are worth estimates in log scale.
 
@@
 
-Plackett-Luce tree using environmental data 
-```
-
-```
-
-Model selection using crossvalidation
-```
-
-```
-
-pseudoR2
-
-```r
-
-```
-
-``regret``
-```r
-
-```
-
- ``reliability``
-
 \newpage
@@
 
-***This is the main section of the article and the reviewers weight the
-description here appropriately***
+***This is the main section of the article and the reviewers weight the description here appropriately***
 
-*Indicate in what way new research questions can be pursued as a result
-of the software (if any).*
+*Indicate in what way new research questions can be pursued as a result of the software (if any).*
 
-*Indicate in what way, and to what extent, the pursuit of existing
-research questions is improved (if so).*
+*Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).*
 
-*Indicate in what way the software has changed the daily practice of its
-users (if so).*
+*Indicate in what way the software has changed the daily practice of its users (if so).*
 
-*Indicate how widespread the use of the software is within and outside
-the intended user group.*
+*Indicate how widespread the use of the software is within and outside the intended user group.*
 
-*Indicate in what way the software is used in commercial settings and/or
-how it led to the creation of spin-off companies (if so).*
+*Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).*
 
-Reproducible and efficient workflows are fundamental in scientific research (Lowndes et al. 2017).
-The gosset package provides functionality that was not previously available from 
-other R packages and which enabled scientific studies based on the analysis of
-ranking data. This functionality enables making the entire workflow to be reproducible
-and more efficient. The utility of the gosset package has been demonstrated by enabling 
-studies based on the analysis of ranking data. For instance, @vanEtten2019, @Moyo_2021  and @deSousa2021
-applied the Plackett-Luce model in combination with recursive partitioning [@Turner2020;@Zeileis2008].
-In these studies, the gosset package supported data preparation, model validation and results presentation
-tasks.
+Reproducible and efficient workflows are fundamental in scientific research (Lowndes et al.
+2017).
+The gosset package provides functionality that was not previously available from other R packages and which enabled scientific studies based on the analysis of ranking data.
+This functionality enables making the entire workflow to be reproducible and more efficient.
+The utility of the gosset package has been demonstrated by enabling studies based on the analysis of ranking data.
+For instance, @vanEtten2019, @Moyo_2021  and @deSousa2021 applied the Plackett-Luce model in combination with recursive partitioning [@Turner2020;@Zeileis2008].
+In these studies, the gosset package supported data preparation, model validation and results presentation tasks.
 
@@
 
-We described the functionality of the R package gosset to support 
-the synthesis and analysis of ranking data. The package provide functions not available in
-existing R packages for analyzing ranking data. We provided an illustrative example 
-covering the main functionality across the stages involved in the analysis workflow.
+We described the functionality of the R package gosset to support the synthesis and analysis of ranking data.
+The package provide functions not available in existing R packages for analyzing ranking data.
+We provided an illustrative example covering the main functionality across the stages involved in the analysis workflow.
 
@@
 
-No conflict of interest exists: We wish to confirm that there are no
-known conflicts of interest associated with this publication and there
-has been no significant financial support for this work that could have
-influenced its outcome.
+No conflict of interest exists: We wish to confirm that there are no known conflicts of interest associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
 
@@
 
-We acknowledge Olga Spellman (Science Writing Service of the Alliance of
-Bioversity International and CIAT) for English editing of this
-manuscript
+We acknowledge Olga Spellman (Science Writing Service of the Alliance of Bioversity International and CIAT) for English editing of this manuscript
+
 \newpage
@@
 
-*Please add the reference to the software repository if DOI for software
-is available.*
+*Please add the reference to the software repository if DOI for software is available.*

### 2022-06-21 inst/paper/paper.Rmd
Update paper.Rmd

@@
 output:
+  word_document: default
   pdf_document: default
-  word_document: default
   html_document:
@@
 
-*Introduce the scientific background and the motivation for developing the software.*
-
-*Explain why the software is important, and describe the exact (scientific) problem(s) it solves.*
-
-*Indicate in what way the software has contributed (or how it will contribute in the future) to the process of scientific discovery; if available, this is to be supported by citing a research paper using the software.*
-
-*Provide a description of the experimental setting (how does the user use the software?).*
-
-*Introduce related work in literature (cite or list algorithms used, other software etc.).*
-
 Participatory experimentation approaches have been increasingly applied in agricultural research [@deRoo2019farm]. While collecting data in ranking format is uncommon in general
@@
 
-*Describe the software in as much as is necessary to establish a vocabulary needed to explain its impact.*
-
 The R package gosset provides functionality supporting the analysis workflows in agricultural experimentation, especially with rank-based approaches.
@@
 
-*Give a short overview of the overall software architecture; provide a pictorial component overview or similar (if possible). If necessary provide implementation details.*
-
 The R package gosset is structured following the guidelines described in the manual for creating R add-on packages [@r_extensions_2022].
@@
 ## Software Functionalities
-
-*Present the major functionalities of the software.*
 
@@
 
-* `regret`
+* `regret` computes the regret [@Loomes_1982;  ].
 
-* `reliability`
+* `reliability` computes the reliability of crop varieties, the probability to outperform the reference variety [@eskridge_1992].
 
-* `worth_bar`
+* `worth_bar`  creates a bar plot of the estimated worth values for each evaluated variety.
 
-* `worth_map`
+* `worth_map` creates a heatmap plot of the estimated worth values for all varieties considering each of the
+evaluated traits.
 
@@
 
-The following lines show the first 10 rankings in the format used by the PlackettLuce package.
+The following lines show the first 10 rankings for the trait vigor, 
+in the format used by the PlackettLuce package.
 
@@
 
-We fit a series of Plackett-Luce models using the function `PlackettLuce` from the package of the same name as input in the lapply function.
+For each trait, we fit a Plackett-Luce model using the function `PlackettLuce` from the package of the same name.
+We use ``lapply`` with the list of rankings and the `PlackettLuce` function as input parameters.
 
@@
 
-The `worth_map` function can be used to visually assess variety performance based on different traits. The values represented in a worth_map are worth estimates in log scale.
+The `worth_map` function can be used to visually assess variety performance based on different traits. The values represented in a worth_map are worth estimates in the log scale.
 
@@
 
-***This is the main section of the article and the reviewers weight the description here appropriately***
-
-*Indicate in what way new research questions can be pursued as a result of the software (if any).*
-
-*Indicate in what way, and to what extent, the pursuit of existing research questions is improved (if so).*
-
-*Indicate in what way the software has changed the daily practice of its users (if so).*
-
-*Indicate how widespread the use of the software is within and outside the intended user group.*
-
-*Indicate in what way the software is used in commercial settings and/or how it led to the creation of spin-off companies (if so).*
-
 Reproducible and efficient workflows are fundamental in scientific research (Lowndes et al.
@@
 # Conclusions
-
-*Set out the conclusion of this original software publication.*
 

### 2022-06-22 inst/paper/paper.Rmd
Update paper.Rmd

@@
 This function is useful when the data were collected as numerical observations, for instance, in a experiment measuring crop yield.
-* `rank_tricot` transforms data in tricot format into rankings.
+* `rank_tricot` transforms data in tricot format into ``PlackettLuce`` rankings.
 
@@
 * `btpermute` deviance-based forward variable selection [@lysen2009permuted] procedure for Bradley-Terry models.  
-* `compare` compare the agreenment between methods.  
+* `compare` compare the agreement between methods.  
 * `crossvalidation` performs k-fold cross-validation, where k could be specified by the user.The default is 10-fold.
@@
 * `kendallTau` computes the Kendall-tau rank correlation coefficient between two rankings [@kendall_1938].  
-* `kendallW` computes Kendall's W (coefficient of concordance) among observed rankings and those predicted by the Plackett-Luce model.  
-*   `pseudoR2` computes goodness-of-fit measure McFadden's pseudo-R2 [@mcfadden1973conditional].
+* `kendallW` computes Kendall's W (coefficient of concordance) among observed rankings and those predicted by the Plackett-Luce model [@Kendall_1939].  
+*   `pseudoR2` computes goodness-of-fit metrics, such as McFadden's pseudo-R2 [@mcfadden1973conditional].
 
@@
 The farmers also provided their overall appreciation about the varieties, i.e., which variety has the best and the worst performance based on the overall performance considering all the traits.
-
 We used the Plackett-Luce model implemented in the R package *PlackettLuce* [@Turner2020].
-To consider the effect of climate on the variety performance, we used climate data.
-For obtaining the climate data, we used the *nasapower* package [@nasapower].
-Climatic indices were computed with the *climatrends* package [@climatrendspkg].
 
@@
 
+To consider the effect of climate factors on variety performance, we used the climatic covariates to fit a 
+Plackett-Luce tree.
+For brevity, we used the variables maximum nighttime temperature (maxNT) and  maximum daytime temperature, but more variables can be used in a Plackett-Luce tree.
+Temperature data were obtained from the NASA POWER dataset, using the R package *nasapower* [@nasapower].
+The temperature data downloaded from NASA POWER were used for
+computing the climatic variables with the package *climatrends* [@climatrendspkg]. In this case, we only use the rankings
+corresponding to the overall performance of varieties.
+
+```{r , message=FALSE, eval=FALSE, echo=TRUE}
+# temp <- temperature(covar[, c("longitude","latitude")], 
+#                     day.one = covar[, "planting_date"],
+#                     span = 80)
+# 
+# head(temp[, 1:2])
+```
+
+To be coupled with climatic covariates, the rankings should be made formatted as grouped rankings. For this
+we used the function ``rank_numeric`` with the parameter ``group = TRUE``.
+
+```{r , message=FALSE, eval=FALSE, echo=TRUE}
+R <- vector(mode = "list", length = length(traits))
+
+for (i in seq_along(traits)) {
+
+  dat_i <- subset(dat, dat$trait == traits[i])
+
+  R[[i]] <- rank_numeric(data = dat_i,
+                         items = "item",
+                         input = "rank", 
+                         id = "id", 
+                         ascending = TRUE,
+                         group = TRUE)
+}
+```
+
+Now we can fit the Plackett-Luce tree with climate covariates. Notice that we used the 9^th^ element in the list
+of rankings as we are only using the overall appreciation ranking.
+
+```{r , message=FALSE, eval=FALSE, echo=TRUE}
+
+R <- R[[9]]
+
+pld <- cbind(R, temp)
+
+pl <- pltree(R ~ maxNT + maxDT, 
+             alpha = .05,
+             data = pld)
+```
+
+The following is an example of the plot made with the function ``plot`` in the gosset package.
+```{r , message=FALSE, eval=FALSE, echo=TRUE}
+
+plot.pltree(pl)
+
+```
+
 \newpage

### 2022-06-22 inst/paper/paper.Rmd
updating

@@
 output:
-  pdf_document: default
-  word_document: default
-  html_document:
-    df_print: paged
+   pdf_document:
+     fig_caption: true
+     number_sections: false
+     keep_tex: false
+latex_engine: pdflatex
+#output:
+# word_document: default
 editor_options:
@@
 bibliography: paper.bib
-header-includes: \usepackage{caption} \captionsetup[figure]{labelformat=empty}
+csl: citationstyle.csl
+header-includes:
+  - \usepackage{pdfpages}
+  - \usepackage{caption}
+  - \usepackage{xcolor}
+  - \captionsetup[figure]{font=small}
+  - \captionsetup[table]{font=small}
 ---
+
+```{r setup_opts, include=FALSE}
+library("knitr")
+knitr::opts_chunk$set(echo = FALSE,
+                      error = FALSE,
+                      message=FALSE,
+                      warning = FALSE)
+```
 
@@
 
-: *Code metadata (mandatory)*
+: *Code metadata*
 
@@
 
-* `plot` provides an alternative to the default plot function of the PlackettLuce package.
-
+* `plot` provides an alternative method to the default plot function of the PlackettLuce package.  
 * `regret` computes the regret [@Loomes_1982;  ].  
-
 * `reliability` computes the reliability of crop varieties, the probability to outperform the reference variety [@eskridge_1992].  
-
 * `worth_bar`  creates a bar plot of the estimated worth values for each evaluated variety.  
-
 * `worth_map` creates a heatmap plot of the estimated worth values for all varieties considering each of the
@@
 
-# Illustrative Examples
+# Illustrative example
 
@@
 The varieties were randomly assigned as incomplete blocks of size three (out of 10 varieties).
-The farmers assessed which of the three varieties has the best and worst performance in nine traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, taste) and overall appreciation).
-The farmers also provided their overall appreciation about the varieties, i.e., which variety has the best and the worst performance based on the overall performance considering all the traits.
-We used the Plackett-Luce model implemented in the R package *PlackettLuce* [@Turner2020].
+The farmers assessed which of the three varieties had the best and worst performance in eight traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, taste).
+The farmers also provided their overall appreciation about the varieties, i.e., which variety had the best and the worst performance based on the overall performance considering all the traits.
+To analyze the data, we use the Plackett-Luce model implemented in the R package *PlackettLuce* [@Turner2020].
 
-First, we load the required packages and data. The `nicabean` dataset contains the tricot rankings (`trial`), the climate covariates (`covar`) and the traits observations (`traits`).
+The `nicabean` is a list with two data frames: i) `trial`, contains the trial data with farmers' evaluations, ranked from 1 to 3, with 1 being the higher ranked variety and 3 the lowest ranked variety for the given trait and incomplete block; ii) `covar`, contains the covariates associated to the on-farm trial plots and farmers.
+This example will require the packages PlackettLuce [@Turner2020], climatrends [@climatrends], nasapower [@nasapower] and ggplot2 [@ggplot2].
 
@@
 
-Make a PlackettLuce rank using the function `rank_numeric`.
+To start the analysis of the data, we transform the rankings into Plackett-Luce rankings using the function `rank_numeric()`.
+We run iteratively over the traits adding the rankings to a list called `R`.
+Since the varieties are ranked in an ascending order, with 1 being the higher ranked and 3 the lower ranked, we use the argument `asceding = TRUE` to indicate which order should be used.
 
@@
 
-The following lines show the first 10 rankings for the trait vigor, 
-in the format used by the PlackettLuce package.
+Then, using the function `kendallTau()` we can assess the correlation between rankings. This approach can be support at least two decision making processes by, for example i) indicating the traits
 
-```{r fetch, message = FALSE, eval = TRUE, echo = TRUE}
-head(R[[1]], 5)
-```
-
-The correlation among rankings can be assessed using the function `kendallTau`.
-
-```{r}
+```{r kendall1, message=FALSE, eval=TRUE, echo=TRUE}
 baseline <- which(grepl("OverallAppreciation", traits))
@@
 kendall$trait <- traits[-baseline]
+```
 
-print(kendall)
+```{r kendall2, message=FALSE, eval=TRUE, echo=FALSE}
+
+kendall <- kendall[,c(3, 1)]
+kendall[,2] <- round(kendall[,2], 3)
+
+kable(kendall, 
+          caption = "Kendall correlation between 'overall performance' and the other traits assessed in the Nicaragua bean on-farm trials.",
+          align = "l",
+          row.names = FALSE)
+
 ```
@@
 
-```{r, message=FALSE, eval=TRUE, echo=TRUE}
+```{r, message=FALSE, eval=TRUE, echo=TRUE, fig.cap="Trait performance (log-worth) of bean varieties in Niragua.
+Variety 'Amadeus' is set as reference (log-worth = 0).
+Blue values indicate a superior performance of varieties for a given trait, compared to the reference.
+Red values indicate a variety with weak performance for the given trait, compared to the reference."}
 worth_map(mod[-baseline],
@@
 Temperature data were obtained from the NASA POWER dataset, using the R package *nasapower* [@nasapower].
-The temperature data downloaded from NASA POWER were used for
-computing the climatic variables with the package *climatrends* [@climatrendspkg]. In this case, we only use the rankings
-corresponding to the overall performance of varieties.
+The temperature data downloaded from NASA POWER were used for computing the climatic variables with the package *climatrends* [@climatrends].
+In this case, we only use the rankings corresponding to the overall performance of varieties.
 
@@
 
-**Zenodo repository for the gosset package:** <https://doi.org/10.5281/zenodo.6339989>
-
-::: thebibliography
+::: {#refs}
 :::
-
-*Please add the reference to the software repository if DOI for software is available.*

### 2022-06-22 inst/paper/paper.Rmd
Update paper.Rmd

@@
 bibliography: paper.bib
-csl: citationstyle.csl
+csl: apa.csl
 header-includes:

### 2022-06-23 inst/paper/paper.Rmd
paper updates

@@
 ::: keyword
-**Keywords**: Bradley-Terry, Plackett-Luce, data science, on-farm trials, participatory-research, tricot
+**Keywords**: Bradley-Terry, Plackett-Luce, data science, on-farm trials, participatory-research, tricot approach
 :::
@@
 |:-----------------|:--------------------------|:-------------------------|
-| C1      | Current code version                                            | 0.4.003                                                    |
+| C1      | Current code version                                            | 0.4                                                    |
 | C2      | Permanent link to code/repository used for this code version    | <https://github.com/AgrDataSci/gosset>                     |
@@
 
-Participatory experimentation approaches have been increasingly applied in agricultural research [@deRoo2019farm]. While collecting data in ranking format is uncommon in general
-agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA]. Recently developed approaches for on-farm experimentation,
-such as the tricot methodology are based on the collection of data in ranking format [@vanetten_beza_2019].
+Participatory experimentation approaches have been increasingly applied in agricultural research [@deRoo2019farm].
+While collecting data in ranking format is uncommon in general agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA].
+Recently developed approaches for on-farm experimentation, such as the tricot methodology are based on the collection of data in ranking format [@vanetten_beza_2019].
 On the other hand, newly proposed approaches for synthesis of crop variety evaluation data largely depend on the analysis of ranking data [@brown2020data].
+
 The analysis of ranking data requires the use of appropriate statistical models such as Plackett-Luce [@luce_individual_1959; @Plackett] or Bradley-Terry [@bradley1952rank].
@@
 However, extended functionality was required for the entire data science workflow, which usually includes: (1) Data preparation and cleaning, (2) modelling and validation, and (3) results presentation.
-For (1) gosset provides functions for converting and preparing data into ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2
-respectively.
+For (1) gosset provides functions for converting and preparing data into ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2 respectively.
 For (2), gosset provides functions for model selection and validation using cross-validation.
@@
 The package is available in The Comprehensive R Archive Network (CRAN) and can be installed by executing `install.packages("gosset")`.
-The package is named in honor to William Sealy Gosset, known by the pen name ‘Student’, a pioneer of modern statistics in small sample experimental design and analysis [@Ziliak_2019].
+The package is named in honour to William Sealy Gosset, known by the pen name ‘Student’, a pioneer of modern statistics in small sample experimental design and analysis [@Ziliak_2019].
 
@@
 
+* `compare` measure the agreement between two methods [@MartinBland1986].
 * `plot` provides an alternative method to the default plot function of the PlackettLuce package.  
-* `regret` computes the regret [@Loomes_1982;  ].  
-* `reliability` computes the reliability of crop varieties, the probability to outperform the reference variety [@eskridge_1992].  
-* `worth_bar`  creates a bar plot of the estimated worth values for each evaluated variety.  
-* `worth_map` creates a heatmap plot of the estimated worth values for all varieties considering each of the
+* `regret` computes the regret coefficients, the loss under the worst possible outcome. A common heuristic in risk assessment strategy [@Loomes_1982].  
+* `reliability` computes the probability of a set of items to outperform a reference item. A common heuristic in plant breeding [@eskridge_1992]. 
+* `worth_bar`  creates a bar plot of the estimated worth values for each evaluated item.  
+* `worth_map` creates a heatmap plot of the estimated worth values for all items considering each of the
 evaluated traits.
@@
 The varieties were randomly assigned as incomplete blocks of size three (out of 10 varieties).
-The farmers assessed which of the three varieties had the best and worst performance in eight traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, taste).
+The farmers assessed which of the three varieties had the best and worst performance in eight traits (vigour, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, taste).
 The farmers also provided their overall appreciation about the varieties, i.e., which variety had the best and the worst performance based on the overall performance considering all the traits.
-To analyze the data, we use the Plackett-Luce model implemented in the R package *PlackettLuce* [@Turner2020].
+To analyse the data, we use the Plackett-Luce model implemented in the R package *PlackettLuce* [@Turner2020].
 
-The `nicabean` is a list with two data frames: i) `trial`, contains the trial data with farmers' evaluations, ranked from 1 to 3, with 1 being the higher ranked variety and 3 the lowest ranked variety for the given trait and incomplete block; ii) `covar`, contains the covariates associated to the on-farm trial plots and farmers.
+The `nicabean` is a list with two data frames.
+The first, `trial`, contains the trial data with farmers' evaluations, ranked from 1 to 3, with 1 being the higher ranked variety and 3 the lowest ranked variety for the given trait and incomplete block.
+The rankings in this dataset were previously transformed from tricot rankings (where participants indicate best and worst) to ordinal rankings using the function `rank_tricot()`.
+The second data frame, `covar`, contains the covariates associated to the on-farm trial plots and farmers.
 This example will require the packages PlackettLuce [@Turner2020], climatrends [@climatrends], nasapower [@nasapower] and ggplot2 [@ggplot2].
@@
 
-To start the analysis of the data, we transform the rankings into Plackett-Luce rankings using the function `rank_numeric()`.
+To start the analysis of the data, we transform the ordinal rankings into Plackett-Luce rankings using the function `rank_numeric()`.
 We run iteratively over the traits adding the rankings to a list called `R`.
@@
 
-Then, using the function `kendallTau()` we can assess the correlation between rankings. This approach can be support at least two decision making processes by, for example i) indicating the traits
+Then, using the function `kendallTau()` we can assess the Kendall tau ($\tau$) coefficient [@Kendall1938].
+This approach can be used, for example, to assess the drivers of farmers choices or to prioritize traits to be tested in a next stage of tricot trials (e.g.
+a lite version of tricot with less questions).
+We use the overall appreciation as the reference trait, and compare the Kendall tau with the other 8 traits.
 
@@
 
+The kendall correlation shows that farmers prioritized the traits yield ($\tau$ = 0.749), taste ($\tau$ = 0.653) and marketability ($\tau$ = 0.639) when assessing overall appreciation.
+
 ```{r kendall2, message=FALSE, eval=TRUE, echo=FALSE}
@@
 
-For each trait, we fit a Plackett-Luce model using the function `PlackettLuce` from the package of the same name.
-We use ``lapply`` with the list of rankings and the `PlackettLuce` function as input parameters.
+Then, for each trait, we fit a Plackett-Luce model using the function `PlackettLuce()` from the package of the same name.
+This will allow us to continue the analysis of the results using the other functions in the package gosset.
 
@@
 
-The `worth_map` function can be used to visually assess variety performance based on different traits. The values represented in a worth_map are worth estimates in the log scale.
+The `worth_map()` function can be used to visually assess item performance based on different characteristics.
+The values represented in a worth_map are log-worth estimates.
+From the breeder or product developer perspective the function `worth_map()` offers a visualization tool to help in identifying item performance based on different characteristics.
 

### 2022-06-24 inst/paper/paper.Rmd
update manuscript

--- 2022-06-22 inst/paper/paper.Rmd
@@
 ---
-title: '`gosset`: An R package for analysis and synthesis of ranking data in agricultural experimentation'
+title: '`gosset`: An R package for analysis and synthesis of ranking data in agricultural
+  experimentation'
 abstract: null
 output:
+  word_document: default
   pdf_document:
-     fig_caption: true
-     number_sections: false
-     keep_tex: false
+    fig_caption: yes
+    number_sections: no
+    keep_tex: no
 latex_engine: pdflatex
-#output:
-# word_document: default
 editor_options:
@@
 
-To consider the effect of climate factors on variety performance, we used the climatic covariates to fit a 
+To consider the effect of climate factors on variety performance, we used climatic covariates to fit a 
 Plackett-Luce tree.
@@
 
-pld <- cbind(R, temp)
+pld <- cbind(R, nicabean_temp_ind)
 
-pl <- pltree(R ~ maxNT + maxDT, 
-             alpha = .05,
+pl <- pltree(R ~ ., 
+             alpha = .9,
              data = pld)
@@
 ```{r , message=FALSE, eval=FALSE, echo=TRUE}
-
-plot.pltree(pl)
+plot(pl)
 
 ```
+!["Plackett-Luce tree plot with gosset function *plot*"](pltree_01.png)
 

### 2022-06-24 inst/paper/paper.Rmd
update manuscript

@@
 output:
-  word_document: default
   pdf_document:
@@
     keep_tex: no
+  word_document: default
 latex_engine: pdflatex
@@
 For brevity, we used the variables maximum nighttime temperature (maxNT) and  maximum daytime temperature, but more variables can be used in a Plackett-Luce tree.
-Temperature data were obtained from the NASA POWER dataset, using the R package *nasapower* [@nasapower].
+Temperature data were obtained from the AgERA5 dataset, using the R package *ag5Tools* [@ag5tools].
 The temperature data downloaded from NASA POWER were used for computing the climatic variables with the package *climatrends* [@climatrends].
@@
 pl <- pltree(R ~ ., 
-             alpha = .9,
+             alpha = .6,
              data = pld)

### 2022-06-24 inst/paper/paper.Rmd
Update paper.Rmd

@@
 The `nicabean` is a list with two data frames: i) `trial`, contains the trial data with farmers' evaluations, ranked from 1 to 3, with 1 being the higher ranked variety and 3 the lowest ranked variety for the given trait and incomplete block; ii) `covar`, contains the covariates associated to the on-farm trial plots and farmers.
-This example will require the packages PlackettLuce [@Turner2020], climatrends [@climatrends], nasapower [@nasapower] and ggplot2 [@ggplot2].
+This example will require the packages PlackettLuce [@Turner2020], climatrends [@climatrends], and ggplot2 [@ggplot2].
 
@@
 library("climatrends")
-library("nasapower")
 library("ggplot2")
@@
 Temperature data were obtained from the AgERA5 dataset, using the R package *ag5Tools* [@ag5tools].
-The temperature data downloaded from NASA POWER were used for computing the climatic variables with the package *climatrends* [@climatrends].
+The temperature data downloaded from AgERA5 were used for computing the climatic variables with the package *climatrends* [@climatrends].
 In this case, we only use the rankings corresponding to the overall performance of varieties.

### 2022-06-24 inst/paper/paper.Rmd
Update paper.Rmd

@@
 output:
+  word_document: default
   pdf_document:
@@
     keep_tex: no
-  word_document: default
 latex_engine: pdflatex
@@
 * `plot` provides an alternative method to the default plot function of the PlackettLuce package.  
-* `regret` computes the regret [@Loomes_1982;  ].  
+* `regret` computes the regret [@Loomes_1982].  
 * `reliability` computes the reliability of crop varieties, the probability to outperform the reference variety [@eskridge_1992].  
@@
 data("nicabean", package = "gosset")
+load("nicabean_temp_ind.rda")
 
@@
 
-```{r , message=FALSE, eval=FALSE, echo=TRUE}
+```{r , message=FALSE, eval=TRUE, echo=TRUE}
 R <- vector(mode = "list", length = length(traits))
@@
 
-```{r , message=FALSE, eval=FALSE, echo=TRUE}
+```{r , message=FALSE, eval=TRUE, echo=TRUE}
 
@@
 \newpage
+
+Table 3 shows the results of reliability for the evaluated common bean varieties in each of 
+the resulting nodes of the Plackett-Luce tree.
+
+```{r rel1, message=FALSE, eval=TRUE, echo=FALSE}
+nicabean_rel <- reliability(pl, ref = "Amadeus 77")
+
+kable(t(nicabean_rel), col.names = c("node 2", "node 3"), caption = "Reliability of evaluated varieties
+      in each of the resulting nodes of the Plackett-Luce tree")
+
+```
 

### 2022-06-25 inst/paper/paper.Rmd
paper.rmd

@@
 bibliography: paper.bib
-csl: apa.csl
+csl: citationstyle.csl
 header-includes:
@@
 This function is useful when the data were collected as numerical observations, for instance, in a experiment measuring crop yield.
-* `rank_tricot` transforms data in tricot format into ``PlackettLuce`` rankings.
+* `rank_tricot` transforms data in tricot format into PlackettLuce rankings.
 
@@
 * `btpermute` deviance-based forward variable selection [@lysen2009permuted] procedure for Bradley-Terry models.
-* `compare` compare the agreement between methods.  
 * `crossvalidation` performs k-fold cross-validation, where k could be specified by the user.The default is 10-fold.
@@
 
-* `plot` provides an alternative method to the default plot function of the PlackettLuce package.  
-* `regret` computes the regret [@Loomes_1982].  
-* `reliability` computes the reliability of crop varieties, the probability to outperform the reference variety [@eskridge_1992].  
-* `worth_bar`  creates a bar plot of the estimated worth values for each evaluated variety.  
-* `worth_map` creates a heatmap plot of the estimated worth values for all varieties considering each of the
+* `compare` is a visualization approach to compare measures from two different methods [@MartinBland1986].
+* `plot` provides an alternative S3 method to `plot.pltree()` method implemented by the PlackettLuce package.  
+* `regret` computes the regret coefficients, the loss under the worst possible outcome. A common heuristic
+in risk assessment strategy [@Loomes_1982].  
+* `reliability` computes the probability of a set of items to outperform a reference item. A common
+heuristic in plant breeding [@eskridge_1992].  
+* `worth_bar` creates a bar plot of the estimated *worth* for each evaluated item.  
+* `worth_map` creates a heatmap plot of the estimated log-worth for all items considering each of the
 evaluated traits.
@@
 
-The `nicabean` is a list with two data frames: i) `trial`, contains the trial data with farmers' evaluations, ranked from 1 to 3, with 1 being the higher ranked variety and 3 the lowest ranked variety for the given trait and incomplete block; ii) `covar`, contains the covariates associated to the on-farm trial plots and farmers.
+The `nicabean` is a list with two data frames.
+The first, `trial`, contains the trial data with farmers’ evaluations, ranked from 1 to 3, with 1 being the higher ranked variety and 3 the lowest ranked variety for the given trait and incomplete block.
+The rankings in this dataset were previously transformed from tricot rankings (where participants indicate best and worst) to ordinal rankings using the function `rank_tricot()`.
+The second data frame, `covar`, contains the covariates associated to the on-farm trial plots and farmers.
 This example will require the packages PlackettLuce [@Turner2020], climatrends [@climatrends], and ggplot2 [@ggplot2].
@@
 
-Then, using the function `kendallTau()` we can assess the correlation between rankings. This approach can be support at least two decision making processes by, for example i) indicating the traits
+Then, using the function `kendallTau()` we can assess the Kendall tau ($\tau$) coefficient [@kendall_1938].
+This approach can be used, for example, to assess the drivers of farmers choices or to prioritize traits to be tested in a next stage of tricot trials (e.g.
+a lite version of tricot with no more than 4 traits to assess).
+We use the overall appreciation as the reference trait, and compare the Kendall tau with the other 8 traits.
 
@@
 kable(kendall, 
-          caption = "Kendall correlation between 'overall performance' and the other traits assessed in the Nicaragua bean on-farm trials.",
+          caption = "Kendall tau correlation between 'overall performance' and the other traits assessed in the Nicaragua bean on-farm trials.",
           align = "l",
@@
 
-For each trait, we fit a Plackett-Luce model using the function `PlackettLuce` from the package of the same name.
-We use ``lapply`` with the list of rankings and the `PlackettLuce` function as input parameters.
+Then, for each trait, we fit a Plackett-Luce model using the function `PlackettLuce()` from the package of the same name.
+This will allow us to continue the analysis of the trial data using the other functions in the package gosset.
 
@@
 Plackett-Luce tree.
-For brevity, we used the variables maximum nighttime temperature (maxNT) and  maximum daytime temperature, but more variables can be used in a Plackett-Luce tree.
-Temperature data were obtained from the AgERA5 dataset, using the R package *ag5Tools* [@ag5tools].
+For brevity, we used the variables maximum night time temperature (maxNT) and  maximum day time temperature, but more variables can be used in a Plackett-Luce tree.
+Temperature data were obtained from the AgERA5 dataset, using the R package ag5Tools [@ag5tools].
 The temperature data downloaded from AgERA5 were used for computing the climatic variables with the package *climatrends* [@climatrends].

### 2022-06-25 inst/paper/paper.Rmd

@@
 output:
-  word_document: default
+  word_document:
   pdf_document:
@@
     keep_tex: no
+  reference_docx: paper_template_format.docx
 latex_engine: pdflatex
@@
 ```
+=======
+To consider the effect of climatic factors on variety performance, we used temperature-derived covariates to fit a
+Plackett-Luce tree.
+For brevity, we used the variables maximum nighttime temperature (maxNT) and  maximum daytime temperature, but more variables can be used in a Plackett-Luce tree.
+Temperature data were obtained from the AgERA5 dataset [@boogaard_2020], using the R package *ag5Tools* [@ag5tools].
+The temperature data downloaded from AgERA5 were used for computing the climatic variables with the package *climatrends* [@climatrends].
+In this case, we only use the rankings corresponding to the overall performance of varieties.
+>>>>>>> Stashed changes
 
@@
 
-Table 3 shows the results of reliability for the evaluated common bean varieties in each of 
-the resulting nodes of the Plackett-Luce tree.
+We can use the function ``reliability`` to compute the reliability of the evaluated common bean
+varieties in each of the resulting nodes of the Plackett-Luce tree (Table 3).
 
@@
       in each of the resulting nodes of the Plackett-Luce tree")
+
+```
+
+With the function ``crossvalidation`` we can validate the model.
+
+```{r, message=FALSE, eval=TRUE, echo=FALSE}
+nb_cv <- crossvalidation(R ~ .,
+                         alpha = .6,
+                         data = pld)
+
+```
+```{r cv, message=FALSE, eval=TRUE, echo=FALSE}
+
+kable(nb_cv$coeffs[,c(1:4, 7, 8)], caption = "Results of 10-fold random cross-validation")
 

### 2022-06-25 inst/paper/paper.Rmd
update manuscript

@@
     keep_tex: no
-  reference_docx: paper_template_format.docx
 latex_engine: pdflatex

### 2022-06-25 inst/paper/paper.Rmd
example1

--- 2022-06-25 inst/paper/paper.Rmd
@@
 output:
-  word_document: default
+#  word_document: default
   pdf_document:
@@
 
-# Required Metadata {#required-metadata .unnumbered}
+# Code metadata {#required-metadata .unnumbered}
 
-# Current code version {#current-code-version .unnumbered}
-
-| **Nr.** | **Code metadata description**                                   | **Please fill in this column**                             |
-|:-----------------|:--------------------------|:-------------------------|
-| C1      | Current code version                                            | 0.4                                                    |
-| C2      | Permanent link to code/repository used for this code version    | <https://github.com/AgrDataSci/gosset>                     |
-| C3      | Code Ocean compute capsule                                      |                                                            |
-| C4      | Legal Code License                                              | MIT                                                        |
-| C5      | Code versioning system used                                     | Git                                                        |
-| C6      | Software code languages, tools, and services used               | R                                                          |
-| C7      | Compilation requirements, operating environments & dependencies |                                                            |
-| C8      | If available Link to developer documentation/manual             | <https://agrdatasci.github.io/gosset/>                     |
-| C9      | Support email for questions                                     | [kaue.desousa@inn.no](mailto:kaue.desousa@inn.no){.email} |
+| **Code metadata description**                 | **Please fill in this column**                             |
+|:--------------------------|:--------------------------|
+| Current code version                                            | 0.4                                                       |
+| Code repository    | <https://github.com/AgrDataSci/gosset>                    |
+| Legal code license                                              | MIT                                                       |
+| Code versioning system used                                     | git                                                       |
+| Software code languages, tools, and services used               | R                                                         |
+| Compilation requirements, operating environments & dependencies | R                                                         |
+| Link to developer documentation             | <https://agrdatasci.github.io/gosset/>                    |
+| Support email for questions                                     | [kaue.desousa@inn.no](mailto:kaue.desousa@inn.no){.email} |
 
@@
 
-Participatory experimentation approaches have been increasingly applied in agricultural research [@deRoo2019farm]. While collecting data in ranking format is uncommon in general
-agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA]. Recently developed approaches for on-farm experimentation,
-such as the tricot methodology are based on the collection of data in ranking format [@vanetten_beza_2019].
+Participatory experimentation approaches have been increasingly applied in agricultural research [@deRoo2019farm].
+While collecting data in ranking format is uncommon in general agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA].
+Recently developed approaches for on-farm experimentation, such as the tricot methodology are based on the collection of data in ranking format [@vanetten_beza_2019].
 On the other hand, newly proposed approaches for synthesis of crop variety evaluation data largely depend on the analysis of ranking data [@brown2020data].
+
 The analysis of ranking data requires the use of appropriate statistical models such as Plackett-Luce [@luce_individual_1959; @Plackett] or Bradley-Terry [@bradley1952rank].
-Functionality for fitting Bradley-Terry and Plackett-Luce models are available in R with the packages BradleyTerry2 and PlackettLuce respectively [@BradleyTerry2; @Turner2020].
+Functionality for fitting Bradley-Terry and Plackett-Luce models are available in R with the packages BradleyTerry2 [@BradleyTerry2] and PlackettLuce [@Turner2020] respectively.
 However, extended functionality was required for the entire data science workflow, which usually includes: (1) Data preparation and cleaning, (2) modelling and validation, and (3) results presentation.
-For (1) gosset provides functions for converting and preparing data into ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2
-respectively.
+For (1) gosset provides functions for converting and preparing data into ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2 respectively.
 For (2), gosset provides functions for model selection and validation using cross-validation.
@@
 
-* `AIC` computes the Akaike Information Criterion @akaike1974] for a Bradley-Terry model fitted with BradleyTerry2 [@BradleyTerry2] or a Plackett-Luce model fitted with [@Turner2020].  
+* `AIC` computes the Akaike Information Criterion [@akaike1974] for a Bradley-Terry model [@BradleyTerry2] or a Plackett-Luce model [@Turner2020].  
 * `btpermute` deviance-based forward variable selection [@lysen2009permuted] procedure for Bradley-Terry models.
@@
 The varieties were randomly assigned as incomplete blocks of size three (out of 10 varieties).
-The farmers assessed which of the three varieties had the best and worst performance in eight traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, taste).
+The farmers assessed which of the three varieties had the best and worst performance in eight traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, and taste).
 The farmers also provided their overall appreciation about the varieties, i.e., which variety had the best and the worst performance based on the overall performance considering all the traits.
@@
 The second data frame, `covar`, contains the covariates associated to the on-farm trial plots and farmers.
-This example will require the packages PlackettLuce [@Turner2020], climatrends [@climatrends], and ggplot2 [@ggplot2].
+This example will require the packages PlackettLuce [@Turner2020], climatrends [@climatrends], chirps [@chirps] and ggplot2 [@ggplot2].
 
-``` {r, message = FALSE, eval = TRUE, echo = TRUE}
+``` {r starting, message = FALSE, eval = TRUE, echo = TRUE}
 library("gosset")
@@
 library("climatrends")
+library("chirps")
 library("ggplot2")
@@
 
-```{r, message = FALSE, eval = TRUE, echo = TRUE}
+```{r rankings, message = FALSE, eval = TRUE, echo = TRUE}
 R <- vector(mode = "list", length = length(traits))
@@
 
-```{r, message=FALSE, eval=TRUE, echo=TRUE}
+```{r PLmodel, message=FALSE, eval=TRUE, echo=TRUE}
 
@@
 
-```{r, message=FALSE, eval=TRUE, echo=TRUE, fig.cap="Trait performance (log-worth) of bean varieties in Niragua.
+```{r worthmap, message=FALSE, eval=TRUE, echo=TRUE, fig.cap="Trait performance (log-worth) of bean varieties in Niragua.
 Variety 'Amadeus' is set as reference (log-worth = 0).
@@
 
-To consider the effect of climate factors on variety performance, we used climatic covariates to fit a 
-Plackett-Luce tree.
-For brevity, we used the variables maximum night time temperature (maxNT) and  maximum day time temperature, but more variables can be used in a Plackett-Luce tree.
-Temperature data were obtained from the AgERA5 dataset, using the R package ag5Tools [@ag5tools].
-The temperature data downloaded from AgERA5 were used for computing the climatic variables with the package *climatrends* [@climatrends].
-In this case, we only use the rankings corresponding to the overall performance of varieties.
+To consider the effect of climate factors on yield, we used climatic covariates to fit a Plackett-Luce tree.
+For simplicity, we used the total rainfall (Rtotal) derived from CHIRPS data [@Funk2015], and computed using the R packages chirps [@chirps] and [@climatrends].
+However, additional covariates can be used in a Plackett-Luce tree, for example using temperature data from R package ag5Tools [@ag5tools] or nasapower [@nasapower].
+
+We request the CHIRPS data using the package chirps. Data should be returned as a matrix. This process can take some minutes to be implemented.
 
 ```{r , message=FALSE, eval=FALSE, echo=TRUE}
-# temp <- temperature(covar[, c("longitude","latitude")], 
-#                     day.one = covar[, "planting_date"],
-#                     span = 80)
-# 
-# head(temp[, 1:2])
+dates <- c(min(covar[, "planting_date"]),
+           max(covar[, "planting_date"]) + 70)
+
+chirps <- get_chirps(covar[, c("longitude","latitude")], 
+                     dates = as.character(dates),
+                     as.matrix = TRUE,
+                     server = "ClimateSERV")
 ```
 
-To be coupled with climatic covariates, the rankings should be made formatted as grouped rankings. For this
-we used the function ``rank_numeric`` with the parameter ``group = TRUE``.
+Than we prepare the matrix to compute the rainfall indices using the function `rainfall()` from package climatrends.
+
+To be linket to covariates, the rankings should be coerced to 'grouped rankings'. For this we used the function `group()` from PlackettLuce. We retain the ranking corresponding to yield.
 
@@
 ```{r rel1, message=FALSE, eval=TRUE, echo=FALSE}
-nicabean_rel <- reliability(pl, ref = "Amadeus 77")
-
-kable(t(nicabean_rel), col.names = c("node 2", "node 3"), caption = "Reliability of evaluated varieties
-      in each of the resulting nodes of the Plackett-Luce tree")
+# nicabean_rel <- reliability(pl, ref = "Amadeus 77")
+# 
+# kable(t(nicabean_rel), col.names = c("node 2", "node 3"), caption = "Reliability of evaluated varieties
+#       in each of the resulting nodes of the Plackett-Luce tree")
 

### 2022-06-25 inst/paper/example_01.R

@@ -2,6 +2,8 @@ library("gosset")
 library("PlackettLuce")
 library("climatrends")
 library("nasapower")
+library("ggplot2")
+library("chirps")
 
 data("nicabean", package = "gosset")
 
@@ -11,7 +13,7 @@ covar <- nicabean$covar
 
 traits <- unique(dat$trait)
 
-R <- list()
+R <- vector(mode = "list", length = length(traits))
 
 for (i in seq_along(traits)) {
   
@@ -24,23 +26,6 @@ for (i in seq_along(traits)) {
                          ascending = TRUE)
 }
 
-head(R[[1]])
-
-r <- R[[1]]
-
-class(r)
-
-r2 <- unclass(r)
-
-class(r2)
-
-mod <- PlackettLuce(r2)
-
-mod
-
-
-
-# kendall 
 baseline <- which(grepl("OverallAppreciation", traits))
 
 kendall <- lapply(R[-baseline], function(X){
@@ -51,10 +36,6 @@ kendall <- do.call("rbind", kendall)
 
 kendall$trait <- traits[-baseline]
 
-kendall
-
-# worth map
-
 mod <- lapply(R, PlackettLuce)
 
 worth_map(mod[-baseline],
@@ -62,6 +43,36 @@ worth_map(mod[-baseline],
           ref = "Amadeus 77") +
   labs(x = "Variety",
        y = "Trait")
+
+
+dates <- c(min(covar[, "planting_date"]),
+           max(covar[, "planting_date"]) + 70)
+
+chirps <- get_chirps(covar[, c("longitude","latitude")], 
+                     dates = as.character(dates),
+                     as.matrix = TRUE,
+                     server = "ClimateSERV")
+
+# rename the matrix
+newnames <- dimnames(chirps)[[2]]
+newnames <- gsub("chirps-v2.0.", "", newnames)
+newnames <- gsub("[.]", "-", newnames)
+
+dimnames(chirps)[[2]] <- newnames
+
+rain <- rainfall(chirps, day.one = covar$planting_date, span = 45)
+
+Y <- which(grepl("Yield", traits))
+
+Y <- group(R[[Y]], index = 1:length(R[[Y]]))
+
+pldY <- cbind(Y, rain)
+
+treeY <- pltree(Y ~ Rtotal, data = pldY, alpha = 0.1)
+
+plot(treeY)
+
+reliability(treeY, ref = "Amadeus 77")
 
 
 

### 2022-06-25 inst/paper/paper.Rmd
paper

@@
 output:
-#  word_document: default
   pdf_document:
@@
     keep_tex: no
+#output: word_document: default
 latex_engine: pdflatex
@@
 
-To consider the effect of climate factors on yield, we used climatic covariates to fit a Plackett-Luce tree.
-For simplicity, we used the total rainfall (Rtotal) derived from CHIRPS data [@Funk2015], and computed using the R packages chirps [@chirps] and [@climatrends].
+To consider the effect of climate factors on yield, we use agro-climatic covariates to fit a Plackett-Luce tree.
+For simplicity, we use the total rainfall (Rtotal) derived from CHIRPS data [@Funk2015], and computed using the R packages chirps [@chirps] and [@climatrends].
 However, additional covariates can be used in a Plackett-Luce tree, for example using temperature data from R package ag5Tools [@ag5tools] or nasapower [@nasapower].
@@
 
-```{r , message=FALSE, eval=FALSE, echo=TRUE}
+```{r chirps, message=FALSE, eval=FALSE, echo=TRUE}
 dates <- c(min(covar[, "planting_date"]),
@@
 
+```{r chirps2, message=FALSE, eval=TRUE, echo=FALSE}
+
+load("nicabean_chirps.rda")
+
+```
+
 Than we prepare the matrix to compute the rainfall indices using the function `rainfall()` from package climatrends.
+
+```{r chirps3, message=FALSE, eval=FALSE, echo=TRUE}
+dates <- c(min(covar[, "planting_date"]),
+           max(covar[, "planting_date"]) + 70)
+
+chirps <- get_chirps(covar[, c("longitude","latitude")], 
+                     dates = as.character(dates),
+                     as.matrix = TRUE,
+                     server = "ClimateSERV")
+```
 
@@
 #       in each of the resulting nodes of the Plackett-Luce tree")
+
+```
+
+With the function ``crossvalidation`` we can validate the model.
+
+```{r, message=FALSE, eval=TRUE, echo=FALSE}
+nb_cv <- crossvalidation(R ~ .,
+                         alpha = .6,
+                         data = pld)
+
+```
+```{r cv, message=FALSE, eval=TRUE, echo=FALSE}
+
+kable(nb_cv$coeffs[,c(1:4, 7, 8)], caption = "Results of 10-fold random cross-validation")
 

### 2022-06-25 inst/paper/paper.Rmd
example1

@@
 
-Than we prepare the matrix to compute the rainfall indices using the function `rainfall()` from package climatrends.
+Than we prepare the matrix to compute the rainfall indices using the function `rainfall()` from package climatrends. We compute rainfall indices for the first 45 days after planting.
 
-```{r chirps3, message=FALSE, eval=FALSE, echo=TRUE}
-dates <- c(min(covar[, "planting_date"]),
-           max(covar[, "planting_date"]) + 70)
+```{r chirps3, message=FALSE, eval=TRUE, echo=TRUE}
+newnames <- dimnames(chirps)[[2]]
+newnames <- gsub("chirps-v2.0.", "", newnames)
+newnames <- gsub("[.]", "-", newnames)
 
-chirps <- get_chirps(covar[, c("longitude","latitude")], 
-                     dates = as.character(dates),
-                     as.matrix = TRUE,
-                     server = "ClimateSERV")
+dimnames(chirps)[[2]] <- newnames
+
+rain <- rainfall(chirps, day.one = covar$planting_date, span = 45)
 ```
 
-To be linket to covariates, the rankings should be coerced to 'grouped rankings'. For this we used the function `group()` from PlackettLuce. We retain the ranking corresponding to yield.
+To be linked to covariates, the rankings should be coerced to a 'grouped_rankings' object. For this we use the function `group()` from PlackettLuce. We retain the ranking corresponding to yield.
 
-```{r , message=FALSE, eval=TRUE, echo=TRUE}
-R <- vector(mode = "list", length = length(traits))
+```{r grouped_ranking, message=FALSE, eval=TRUE, echo=TRUE}
+G <- which(grepl("Yield", traits))
 
-for (i in seq_along(traits)) {
-
-  dat_i <- subset(dat, dat$trait == traits[i])
-
-  R[[i]] <- rank_numeric(data = dat_i,
-                         items = "item",
-                         input = "rank",
-                         id = "id",
-                         ascending = TRUE,
-                         group = TRUE)
-}
+G <- group(R[[G]], index = 1:length(R[[G]]))
 ```
@@
 
-```{r , message=FALSE, eval=TRUE, echo=TRUE}
+```{r pltree, message=FALSE, eval=TRUE, echo=TRUE}
+pldG <- cbind(G, rain)
 
-R <- R[[9]]
-
-pld <- cbind(R, nicabean_temp_ind)
-
-pl <- pltree(R ~ .,
-             alpha = .6,
-             data = pld)
+tree <- pltree(G ~ Rtotal, data = pldG, alpha = 0.1)
 ```
 
-The following is an example of the plot made with the function ``plot`` in the gosset package.
-```{r , message=FALSE, eval=FALSE, echo=TRUE}
-plot(pl)
+The following is an example of the plot made with the function `plot()` in the gosset package.
 
+```{r , message=FALSE, eval=TRUE, echo=TRUE, fig.width=4, fig.height=4, fig.cap="Plackett-Luce tree of bean yield in on-farm trials"}
+plot(tree)
 ```
-!["Plackett-Luce tree plot with gosset function *plot*"](pltree_01.png)
 
-\newpage
-
-We can use the function ``reliability`` to compute the reliability of the evaluated common bean
+We can use the function `reliability()` to compute the reliability of the evaluated common bean
 varieties in each of the resulting nodes of the Plackett-Luce tree (Table 3).
@@
 ```{r rel1, message=FALSE, eval=TRUE, echo=FALSE}
-# nicabean_rel <- reliability(pl, ref = "Amadeus 77")
-# 
-# kable(t(nicabean_rel), col.names = c("node 2", "node 3"), caption = "Reliability of evaluated varieties
-#       in each of the resulting nodes of the Plackett-Luce tree")
-
-```
-
-With the function ``crossvalidation`` we can validate the model.
-
-```{r, message=FALSE, eval=TRUE, echo=FALSE}
-nb_cv <- crossvalidation(R ~ .,
-                         alpha = .6,
-                         data = pld)
-
-```
-```{r cv, message=FALSE, eval=TRUE, echo=FALSE}
-
-kable(nb_cv$coeffs[,c(1:4, 7, 8)], caption = "Results of 10-fold random cross-validation")
-
+rel <- reliability(tree, ref = "Amadeus 77")
 ```

### 2022-06-25 R/reliabilty.R

@@ -107,7 +107,7 @@ reliability.PlackettLuce <- function(x, ref, ...) {
   
   rownames(x) <- 1:nrow(x)
   
-  names(x) <- c("item", "reliabity", "reliabilitySE", "worth", "Zvalue", "Pr(>|z|)", "")
+  names(x) <- c("item", "reliability", "reliabilitySE", "worth", "Zvalue", "Pr(>|z|)", "")
   
   return(x)
   

### 2022-06-25 inst/paper/example_01.R

@@ -48,10 +48,12 @@ worth_map(mod[-baseline],
 dates <- c(min(covar[, "planting_date"]),
            max(covar[, "planting_date"]) + 70)
 
-chirps <- get_chirps(covar[, c("longitude","latitude")], 
-                     dates = as.character(dates),
-                     as.matrix = TRUE,
-                     server = "ClimateSERV")
+# chirps <- get_chirps(covar[, c("longitude","latitude")], 
+#                      dates = as.character(dates),
+#                      as.matrix = TRUE,
+#                      server = "ClimateSERV")
+
+load("inst/paper/nicabean_chirps.rda")
 
 # rename the matrix
 newnames <- dimnames(chirps)[[2]]
@@ -62,18 +64,17 @@ dimnames(chirps)[[2]] <- newnames
 
 rain <- rainfall(chirps, day.one = covar$planting_date, span = 45)
 
-Y <- which(grepl("Yield", traits))
+G <- which(grepl("Yield", traits))
 
-Y <- group(R[[Y]], index = 1:length(R[[Y]]))
+G <- group(R[[G]], index = 1:length(R[[G]]))
 
-pldY <- cbind(Y, rain)
+pldG <- cbind(G, rain)
 
-treeY <- pltree(Y ~ Rtotal, data = pldY, alpha = 0.1)
+tree <- pltree(G ~ SDII, data = pldG, alpha = 0.1)
 
-plot(treeY)
+plot(tree)
 
-reliability(treeY, ref = "Amadeus 77")
-
+reliability(tree, ref = "Amadeus 77")
 
 
 

### 2022-06-26 inst/paper/paper.Rmd
paper

@@
 ^5^ Humboldt University Berlin, Berlin, Germany  
-
 ^[*]^ Correspondence should be addressed to: <kaue.desousa@inn.no>
@@
 # Abstract
+
 Appropriate data management and analysis are necessary to produce practical information from agricultural experimentation data.
@@
 data("nicabean", package = "gosset")
-load("nicabean_temp_ind.rda")
 
@@
 The following is an example of the plot made with the function `plot()` in the gosset package.
+The functions `node_labels()`, `node_rules()` and `top_items()` can be used to identify the splitting variables in the tree, the rules used to split the tree and the best items in each node, respectively.
 
-```{r , message=FALSE, eval=TRUE, echo=TRUE, fig.width=4, fig.height=4, fig.cap="Plackett-Luce tree of bean yield in on-farm trials"}
-plot(tree)
+```{r node_info, message=FALSE, eval=FALSE,echo=TRUE}
+plot(tree, ref = "Amadeus 77", ci.level = 0.9)
+
+node_labels(tree)
+
+node_rules(tree)
+
+top_items(tree, top = 3)
+```
+
+```{r pltree3, message=FALSE, eval=TRUE, echo=FALSE, out.width="50%", fig.align='center', fig.cap="Effect of total rainfall (Rtotal) on yield of common beans in on-farm trials.
+Agroclimate variables are obtained from planting date to the first 45 days of plant growth.
+Axis X presents log-worth, the probability of outperforming the other varieties in the set."}
+knitr::include_graphics("pltree_01.png")
 ```
@@
 varieties in each of the resulting nodes of the Plackett-Luce tree (Table 3).
+This helps in identifying the varieties with higher probability to outperform a variety check (Amadeus 77).
+For simplicity, we present only the varieties with reliability $\geq$ 0.5.
 
-```{r rel1, message=FALSE, eval=TRUE, echo=FALSE}
+```{r rel1, message=FALSE, eval=FALSE, echo=TRUE}
+reliability(tree, ref = "Amadeus 77")
+```
+
+```{r rel2, message=FALSE, eval=TRUE, echo=FALSE}
+
 rel <- reliability(tree, ref = "Amadeus 77")
+
+rel <- rel[rel$reliability >= 0.5, ]
+
+rel <- rel[c(1:5)]
+
+rel[c(3:5)] <- lapply(rel[c(3:5)], function(x){round(x, 3)})
+
+kable(rel,
+          caption = "Reliability of common bean varieties based on yield performance under different rainfall conditions from planting date to the first 45 days of plant growth.
+Variety Amadeus 77 is set as reference.",
+          align = "l",
+          row.names = FALSE)
+
 ```
@@
 Reproducible and efficient workflows are fundamental in scientific research (Lowndes et al.
-2017).
+2017) [@Lowndes2017].
 The gosset package provides functionality that was not previously available from other R packages and which enabled scientific studies based on the analysis of ranking data.
@@
 The utility of the gosset package has been demonstrated by enabling studies based on the analysis of ranking data.
-For instance, @vanEtten2019, @Moyo_2021  and @deSousa2021 applied the Plackett-Luce model in combination with recursive partitioning [@Turner2020;@Zeileis2008].
+For instance, @vanEtten2019, @Moyo_2021 and @deSousa2021 applied the Plackett-Luce model in combination with recursive partitioning [@Turner2020; @Zeileis2008].
 In these studies, the gosset package supported data preparation, model validation and results presentation tasks.

### 2022-06-26 R/nodeinfo.R

@@ -326,7 +326,7 @@ build_tree_nodes <- function(x,
                              node.ids = NULL,
                              n.obs = NULL){
   
-  if (isTRUE(log)) ref <- NULL
+  if (isFALSE(log)) ref <- NULL
   
   if (is.null(node.ids)) {
     node.ids <- 1:length(x)

### 2022-06-26 inst/paper/example_01.R

@@ -70,13 +70,30 @@ G <- group(R[[G]], index = 1:length(R[[G]]))
 
 pldG <- cbind(G, rain)
 
-tree <- pltree(G ~ SDII, data = pldG, alpha = 0.1)
+pldG$Rtotal <- round(pldG$Rtotal, 2)
 
-plot(tree)
+tree <- pltree(G ~ Rtotal, data = pldG, alpha = 0.1)
+
+plot(tree, ref = "Amadeus 77", ci.level = 0.9)
+node_labels(tree)
+node_rules(tree)
+top_items(tree, top = 3)
+
+ggsave("inst/paper/pltree_01.png", 
+       last_plot(),
+       width = 10,
+       height = 10)
+
 
 reliability(tree, ref = "Amadeus 77")
 
 
+rel <- reliability(tree, ref = "Amadeus 77")
 
+rel <- rel[rel$reliability >= 0.5, ]
 
+rel <- rel[c(1:5)]
 
+rel[c(3:5)] <- lapply(rel[c(3:5)], function(x){round(x, 3)})
+
+rel

### 2022-06-26 inst/paper/paper.Rmd
share-paper

@@
 abstract: null
+# output:
+#   pdf_document:
+#     fig_caption: yes
+#     number_sections: no
+#     keep_tex: no
 output: 
-  pdf_document:
-    fig_caption: yes
-    number_sections: no
-    keep_tex: no
-#output: word_document: default
+ word_document: default
 latex_engine: pdflatex
@@
 There is also an ongoing trend advocating for programmatic tools that supports reproducible workflows in scientific research.
-We developed the R package gosset, providing functionality to support analysis workflows with rank-based models, such as Plackett-Luce and Bradley-Terry.
+We developed the R package `gosset`, providing functionality to support analysis workflows with rank-based models.
 The gosset package facilitates data preparation, modelling and results presentation stages.
-We demonstrate the functionality of the package with a case of on-farm evaluations of common bean (*Phaseolus vulgaris* L.) genotypes in Nicaragua.
+We demonstrate the functionality of the package with a case of a decentralized on-farm trial of common bean (*Phaseolus vulgaris* L.) varieties in Nicaragua.
 
@@
 ::: keyword
-**Keywords**: Bradley-Terry, Plackett-Luce, data science, on-farm trials, participatory-research, tricot
+**Keywords**: Bradley-Terry, data science, Plackett-Luce, on-farm trials, tricot approach
 :::
@@
 While collecting data in ranking format is uncommon in general agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA].
-Recently developed approaches for on-farm experimentation, such as the tricot methodology are based on the collection of data in ranking format [@vanetten_beza_2019].
+Recently developed approaches for on-farm experimentation, such as the tricot approach are based on the collection of data in ranking format [@vanetten_beza_2019].
 On the other hand, newly proposed approaches for synthesis of crop variety evaluation data largely depend on the analysis of ranking data [@brown2020data].
@@
 
-To start the analysis of the data, we transform the rankings into Plackett-Luce rankings using the function `rank_numeric()`.
+To start the analysis of the data, we transform the ordinal rankings into Plackett-Luce rankings (a sparse matrix) using the function `rank_numeric()`.
 We run iteratively over the traits adding the rankings to a list called `R`.
@@
 
-The `worth_map()` function can be used to visually assess item performance based on different characteristics.
+The `worth_map()` function can be used to visually assess and compare item performance based on different characteristics.
 The values represented in a worth_map are log-worth estimates.
-From the breeder or product developer perspective the function `worth_map()` offers a visualization tool to help in identifying item performance based on different characteristics.
+From the breeder or product developer perspective the function `worth_map()` offers a visualization tool to help in identifying item performance based on different characteristics and select crossing materials.
 
@@
 To consider the effect of climate factors on yield, we use agro-climatic covariates to fit a Plackett-Luce tree.
-For simplicity, we use the total rainfall (Rtotal) derived from CHIRPS data [@Funk2015], and computed using the R packages chirps [@chirps] and [@climatrends].
-However, additional covariates can be used in a Plackett-Luce tree, for example using temperature data from R package ag5Tools [@ag5tools] or nasapower [@nasapower].
+For simplicity, we use the total rainfall (Rtotal) derived from CHIRPS data [@Funk2015], obtained in R using the R package chirps [@chirps].
+Additional covariates can be used in a Plackett-Luce tree, for example using temperature data from R package ag5Tools [@ag5tools] or nasapower [@nasapower].
 
@@
 
-Than we prepare the matrix to compute the rainfall indices using the function `rainfall()` from package climatrends. We compute rainfall indices for the first 45 days after planting.
+Than we prepare the matrix to compute the rainfall indices using the function `rainfall()` from package climatrends [@climatrends]. We compute rainfall indices for the first 45 days after planting.
 
@@
 ```{r grouped_ranking, message=FALSE, eval=TRUE, echo=TRUE}
-G <- which(grepl("Yield", traits))
+yield <- which(grepl("Yield", traits))
 
-G <- group(R[[G]], index = 1:length(R[[G]]))
+G <- group(R[[yield]], index = 1:length(R[[yield]]))
 ```
 
-Now we can fit the Plackett-Luce tree with climate covariates. Notice that we used the 9^th^ element in the list
-of rankings as we are only using the overall appreciation ranking.
+Now we can fit the Plackett-Luce tree with climate covariates.
 
@@
 
-```{r pltree3, message=FALSE, eval=TRUE, echo=FALSE, out.width="50%", fig.align='center', fig.cap="Effect of total rainfall (Rtotal) on yield of common beans in on-farm trials.
+```{r pltree3, message=FALSE, eval=TRUE, echo=FALSE, out.width="60%", fig.align='center', fig.cap="Effect of total rainfall (Rtotal) on yield of common beans in on-farm trials.
 Agroclimate variables are obtained from planting date to the first 45 days of plant growth.
@@
 
+The result shows that three varieties can marginally outperform Amadeus 77 under drier growing conditions (Rtotal $\leq$ 193.82 mm) whereas two varieties have a superior yield performance when under higher rainfall conditions (Rtotal $>$ 193.82 mm) compared to the reference.
+This approach helps in identifying superior varieties for different target population environments.
+For example, the variety ALS 0532-6 shows weak performance in the whole yield ranking, however for the sub-group of higher rainfall, the variety outperforms all the others.
+Combining rankings with socio-economic covariates could also support the identification of superior materials for different market segments.
+
+A better approach to assess the performance of varieties can be using the 'Overall Appreciation", since we expect this trait to capture the performance of the variety not only for yield, but for all the other traits prioritized by farmers (Table 2).
+To support this hypotheses, we use the function `compare()` which applies the approach proposed by Bland and Altman (1986) [@MartinBland1986] to assess the agreement between two different measures.
+We compare overall vs yield.
+If both measures completely agree, all the varieties should be centered to 0 in the axis Y.
+
+```{r compare, message=FALSE, echo=TRUE, eval=FALSE, out.width="50%"}
+Overall <- PlackettLuce(R[[baseline]])
+Yield <- PlackettLuce(R[[yield]])
+
+compare(Overall, Yield) +
+  labs(x = "Average log(worth)",
+       y = "Difference (Overall Appreciation - Yield)")
+```
+
+The chart shows no complete agreement between overall appreciation and yield.
+For example, variety SX 14825-7-1 shows superior performance for overall appreciation when compared with yield.
+Looking at the log-worth in the heat map of Figure 1, we can argue that the superior performance of the given variety is also related to taste, marketability and diseases resistance.
+This performance, however, was not captured when assessing only yield.
+
+```{r compare2, message=FALSE, eval=TRUE, echo=FALSE, out.width="50%", fig.align='center', fig.cap="Agreement between overall appreciation and yield for crop variety performance in on-farm trials."}
+knitr::include_graphics("compare_02.png")
+```
+
+Here we presented a simple workflow to assess crop variety performance and trait prioritization in decentralized on-farm trials with the tricot approach.
+A more complex workflow will also utilize other functions available in gosset, for example, a forward selection combined with `crossvalidation()` to improve model robustness, or model selection with `btpermute()` to consider all possible permutations in Bradley-Terry models, or a risk analysis using `regret()` to support the selection of varieties, or using `rank_numeric()` to combine legacy data and deal with heterogeneous data from different trials.
+All of these were previously implemented and validated elsewhere [@vanEtten2019; @Moyo_2021; @deSousa2021; @Steinkemindata2019; @brownlegacydata].
+
 # Impact
@@
 This functionality enables making the entire workflow to be reproducible and more efficient.
-The utility of the gosset package has been demonstrated by enabling studies based on the analysis of ranking data.
-For instance, @vanEtten2019, @Moyo_2021 and @deSousa2021 applied the Plackett-Luce model in combination with recursive partitioning [@Turner2020; @Zeileis2008].
+The utility of the gosset package has been demonstrated by enabling studies based on the analysis of decentralized on-farm trial data and/or  heterogeneous data from different sources.
+For instance, van Etten et al.
+(2019) [@vanEtten2019], Moyo et al.
+(2021) [@Moyo_2021] and de Sousa et al.
+(2021) [@deSousa2021] applied the Plackett-Luce model in combination with recursive partitioning [@Turner2020; @Zeileis2008].
 In these studies, the gosset package supported data preparation, model validation and results presentation tasks.
@@
 
-No conflict of interest exists: We wish to confirm that there are no known conflicts of interest associated with this publication and there has been no significant financial support for this work that could have influenced its outcome.
+The authors declare no conflict of interests.
 
@@
 
-We acknowledge Olga Spellman (Science Writing Service of the Alliance of Bioversity International and CIAT) for English editing of this manuscript
+We acknowledge Olga Spellman (Science Writing Service of the Alliance of Bioversity International and CIAT) for English editing of this manuscript.
 

### 2022-06-26 R/compare.R

@@ -80,9 +80,9 @@ compare.default <- function(x, y, labels = NULL, ...){
 compare.PlackettLuce <- function(x, y, ...){
   
   # get the probability of winning and than set back as log
-  X <- coefficients(x, ref = 1, log = TRUE)
+  X <- log(coefficients(x, log = FALSE))
   
-  Y <- coefficients(y, ref = 1, log = TRUE)
+  Y <- log(coefficients(y, log = FALSE))
   
   lab <- names(X)
   

### 2022-06-26 inst/paper/paper.Rmd
paper.rmd

@@
 abstract: null
+output:
+  pdf_document:
+    fig_caption: yes
+    #number_sections: no
+    keep_tex: no
 # output: 
-#   pdf_document:
-#     fig_caption: yes
-#     number_sections: no
-#     keep_tex: no
-output: 
- word_document: default
+#  word_document: default
 latex_engine: pdflatex
@@
 ^5^ Humboldt University Berlin, Berlin, Germany  
-^[*]^ Correspondence should be addressed to: <kaue.desousa@inn.no>
+^[*]^ Correspondence should be addressed to: <k.desousa@cgiar.org>
 
@@
 | Link to developer documentation             | <https://agrdatasci.github.io/gosset/>                    |
-| Support email for questions                                     | [kaue.desousa@inn.no](mailto:kaue.desousa@inn.no){.email} |
+| Support email for questions                                     | [desousa.kaue@gmail.com](mailto:desousa.kaue@gmail.com){.email} |
 

### 2022-06-26 inst/paper/paper.Rmd
Update paper.Rmd

@@
     fig_caption: yes
-    #number_sections: no
+    number_sections: no
     keep_tex: no
@@
 
-Than we prepare the matrix to compute the rainfall indices using the function `rainfall()` from package climatrends [@climatrends]. We compute rainfall indices for the first 45 days after planting.
+We compute the rainfall indices from planting date to the first 45 days of plant growth using the function `rainfall()` from package climatrends [@climatrends].
 
@@
 We acknowledge Olga Spellman (Science Writing Service of the Alliance of Bioversity International and CIAT) for English editing of this manuscript.
+The nicabean data was generated through the Cooperative Agreement AID-OAA-F-14-00035, which was made possible by the generous support of the American people through the US Agency for International Development (USAID).
+The package was developed as part of the CGIAR Research Programs on Climate Change, Agriculture and Food Security (CCAFS) and the CGIAR Research Program on Roots, Tubers and Bananas (RTB), which were carried out with support from the CGIAR Trust Fund and through bilateral funding agreements (details are at https://www.cgiar.org/funders).
+New analytical approaches were developed during the projects Accelerated Varietal Improvement and Seed Systems in Africa (AVISA) and 10000FARMS supported by the Bill and Melinda Gates Foundation.
+The views expressed in this document cannot be taken to reflect the official opinions of these organizations.
 

### 2022-06-27 R/reliabilty.R
vignette

@@ -107,6 +107,8 @@ reliability.PlackettLuce <- function(x, ref, ...) {
   
   names(x) <- c("item", "reliability", "reliabilitySE", "worth", "Zvalue", "Pr(>|z|)")
   
+  class(x) <- union("gosset_df", class(x))
+  
   return(x)
   
 }
@@ -139,7 +141,7 @@ reliability.pltree <- function(x, ref, ...) {
   
   rel <- cbind(node = rep(node_id, each = nitems), rel)
   
-  names(rel)[8] <- ""
+  class(rel) <- union("gosset_df", class(rel))
   
   return(rel)
   

### 2022-06-27 vignettes/Overview.R

@@ -1,8 +1,15 @@
-## ----fetch, message=FALSE, eval=TRUE, echo=TRUE-------------------------------
+## ----setup_opts, include=FALSE------------------------------------------------
+library("knitr")
+knitr::opts_chunk$set(echo = FALSE,
+                      error = FALSE,
+                      message=FALSE,
+                      warning = FALSE)
+
+## ----starting, message = FALSE, eval = TRUE, echo = TRUE----------------------
 library("gosset")
 library("PlackettLuce")
 library("climatrends")
-library("nasapower")
+library("chirps")
 library("ggplot2")
 
 data("nicabean", package = "gosset")
@@ -11,30 +18,25 @@ dat <- nicabean$trial
 
 covar <- nicabean$covar
 
-lapply(nicabean, head)
-
-
-## ----rank, message=FALSE, eval=TRUE, echo=TRUE--------------------------------
-
 traits <- unique(dat$trait)
 
-R <- list()
+
+## ----rankings, message = FALSE, eval = TRUE, echo = TRUE----------------------
+R <- vector(mode = "list", length = length(traits))
 
 for (i in seq_along(traits)) {
-  
+
   dat_i <- subset(dat, dat$trait == traits[i])
-  
+
   R[[i]] <- rank_numeric(data = dat_i,
                          items = "item",
-                         input = "rank", 
-                         id = "id", 
+                         input = "rank",
+                         id = "id",
                          ascending = TRUE)
 }
 
-head(R[[1]])
 
-
-## ----kendall, message=FALSE, eval=TRUE, echo=TRUE-----------------------------
+## ----kendall1, message=FALSE, eval=TRUE, echo=TRUE----------------------------
 baseline <- which(grepl("OverallAppreciation", traits))
 
 kendall <- lapply(R[-baseline], function(X){
@@ -45,45 +47,101 @@ kendall <- do.call("rbind", kendall)
 
 kendall$trait <- traits[-baseline]
 
-print(kendall)
+## ----kendall2, message=FALSE, eval=TRUE, echo=FALSE---------------------------
 
-## ----worth, message=FALSE, eval=TRUE, echo=TRUE-------------------------------
+kendall <- kendall[,c(3, 1)]
+kendall[,2] <- round(kendall[,2], 3)
+
+kable(kendall,
[…]

### 2022-06-27 vignettes/example_01.R

@@ -1,0 +1,123 @@
+library("gosset")
+library("PlackettLuce")
+library("climatrends")
+library("nasapower")
+library("ggplot2")
+library("chirps")
+
+data("nicabean", package = "gosset")
+
+dat <- nicabean$trial
+
+covar <- nicabean$covar
+
+traits <- unique(dat$trait)
+
+R <- vector(mode = "list", length = length(traits))
+
+for (i in seq_along(traits)) {
+  
+  dat_i <- subset(dat, dat$trait == traits[i])
+  
+  R[[i]] <- rank_numeric(data = dat_i,
+                         items = "item",
+                         input = "rank", 
+                         id = "id", 
+                         ascending = TRUE)
+}
+
+baseline <- which(grepl("OverallAppreciation", traits))
+
+kendall <- lapply(R[-baseline], function(X){
+  kendallTau(x = X, y = R[[baseline]])
+})
+
+kendall <- do.call("rbind", kendall)
+
+kendall$trait <- traits[-baseline]
+
+mod <- lapply(R, PlackettLuce)
+
+worth_map(mod[-baseline],
+          labels = traits[-baseline], 
+          ref = "Amadeus 77") +
+  labs(x = "Variety",
+       y = "Trait")
+
+
+dates <- c(min(covar[, "planting_date"]),
+           max(covar[, "planting_date"]) + 70)
+
+# chirps <- get_chirps(covar[, c("longitude","latitude")], 
+#                      dates = as.character(dates),
+#                      as.matrix = TRUE,
+#                      server = "ClimateSERV")
+
+load("inst/paper/nicabean_chirps.rda")
+
+# rename the matrix
+newnames <- dimnames(chirps)[[2]]
+newnames <- gsub("chirps-v2.0.", "", newnames)
+newnames <- gsub("[.]", "-", newnames)
+
+dimnames(chirps)[[2]] <- newnames
+
+rain <- rainfall(chirps, day.one = covar$planting_date, span = 45)
+
+yield <- which(grepl("Yield", traits))
+
+G <- group(R[[yield]], index = 1:length(R[[yield]]))
+
+pldG <- cbind(G, rain)
+
+pldG$Rtotal <- round(pldG$Rtotal, 2)
+
+tree <- pltree(G ~ Rtotal, data = pldG, alpha = 0.1)
+
[…]

### 2022-06-27 vignettes/Overview.R
docs

@@ -1,10 +1,3 @@
-## ----setup_opts, include=FALSE------------------------------------------------
-library("knitr")
-knitr::opts_chunk$set(echo = FALSE,
-                      error = FALSE,
-                      message=FALSE,
-                      warning = FALSE)
-
 ## ----starting, message = FALSE, eval = TRUE, echo = TRUE----------------------
 library("gosset")
 library("PlackettLuce")
@@ -19,6 +12,8 @@ dat <- nicabean$trial
 covar <- nicabean$covar
 
 traits <- unique(dat$trait)
+
+dat
 
 
 ## ----rankings, message = FALSE, eval = TRUE, echo = TRUE----------------------
@@ -52,10 +47,7 @@ kendall$trait <- traits[-baseline]
 kendall <- kendall[,c(3, 1)]
 kendall[,2] <- round(kendall[,2], 3)
 
-kable(kendall,
-          caption = "Kendall tau correlation between 'overall performance' and the other traits assessed in the Nicaragua bean on-farm trials.",
-          align = "l",
-          row.names = FALSE)
+kendall
 
 
 ## ----PLmodel, message=FALSE, eval=TRUE, echo=TRUE-----------------------------
@@ -98,22 +90,26 @@ yield <- which(grepl("Yield", traits))
 
 G <- group(R[[yield]], index = 1:length(R[[yield]]))
 
+head(G)
+
+
 ## ----pltree, message=FALSE, eval=TRUE, echo=TRUE------------------------------
 pldG <- cbind(G, rain)
 
 tree <- pltree(G ~ Rtotal, data = pldG, alpha = 0.1)
 
-## ----node_info, message=FALSE, eval=FALSE,echo=TRUE---------------------------
-#  plot(tree, ref = "Amadeus 77", ci.level = 0.9)
-#  
-#  node_labels(tree)
-#  
-#  node_rules(tree)
-#  
-#  top_items(tree, top = 3)
+print(tree)
 
-## ----pltree3, message=FALSE, eval=TRUE, echo=FALSE, out.width="60%", fig.align='center', fig.cap="Effect of total rainfall (Rtotal) on yield of common beans in on-farm trials. Agroclimate variables are obtained from planting date to the first 45 days of plant growth. Axis X presents log-worth, the probability of outperforming the other varieties in the set."----
-knitr::include_graphics("pltree_01.png")
+
[…]

### 2022-06-27 R/reliability.R
docs

@@ -12,6 +12,8 @@
 #' @param ... additional arguments passed to methods
 #' @return the reliability based on the worth parameters
 #' @author Kauê de Sousa, David Brown, Jacob van Etten
+#' @references 
+#'  Eskridge and Mumm (1992). Theoret. Appl. Genetics 84, 494–500 \doi{10.1007/BF00229512}.
 #' @examples
 #' # vector example
 #' x <- c(9.5, 12, 12.3, 17)

### 2022-06-29 inst/paper/paper.Rmd
add-multcompPL

@@
     keep_tex: no
-# output: 
-#  word_document: default
+  word_document: default
 latex_engine: pdflatex

### 2022-07-07 inst/paper/paper.Rmd
Update paper.Rmd

@@
 
-Appropriate data management and analysis are necessary to produce practical information from agricultural experimentation data.
-There is also an ongoing trend advocating for programmatic tools that supports reproducible workflows in scientific research.
-We developed the R package `gosset`, providing functionality to support analysis workflows with rank-based models.
+Appropriate data management and analysis are necessary for generating practical information from agricultural experimentation data.
+Modern approaches advocate using programmatic tools that support reproducible workflows in scientific research.
+We developed the R package gosset, which provides useful functionality for supporting analysis workflows with rank-based models.
 The gosset package facilitates data preparation, modelling and results presentation stages.
-We demonstrate the functionality of the package with a case of a decentralized on-farm trial of common bean (*Phaseolus vulgaris* L.) varieties in Nicaragua.
+It uses novel functions not available in existing R packages for analyzing ranking data.
+This paper demonstrates the package functionality using the case study of a decentralized on-farm trial of common bean (*Phaseolus vulgaris* L.) varieties in Nicaragua.
 
@@
 While collecting data in ranking format is uncommon in general agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA].
-Recently developed approaches for on-farm experimentation, such as the tricot approach are based on the collection of data in ranking format [@vanetten_beza_2019].
-On the other hand, newly proposed approaches for synthesis of crop variety evaluation data largely depend on the analysis of ranking data [@brown2020data].
+Recently developed approaches for on-farm experimentation, such as the triadic comparison of technologies (tricot) approach are based on the collection of data in ranking format  [@vanetten_beza_2019].
+On the other hand, newly proposed approaches for synthesizing crop variety evaluation data largely depend on the analysis of ranking data [@brown2020data].
 
@@
 However, extended functionality was required for the entire data science workflow, which usually includes: (1) Data preparation and cleaning, (2) modelling and validation, and (3) results presentation.
-For (1) gosset provides functions for converting and preparing data into ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2 respectively.
+For (1) gosset provides functions for converting and preparing data into a ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2 respectively.
 For (2), gosset provides functions for model selection and validation using cross-validation.
@@
 The R package gosset provides functionality supporting the analysis workflows in agricultural experimentation, especially with rank-based approaches.
-The package is available in The Comprehensive R Archive Network (CRAN) and can be installed by executing `install.packages("gosset")`.
-The package is named in honor to William Sealy Gosset, known by the pen name ‘Student’, a pioneer of modern statistics in small sample experimental design and analysis [@Ziliak_2019].
+The package is available in the Comprehensive R Archive Network (CRAN) and can be installed by executing `install.packages("gosset")`.
+The package is named in honor of William Sealy Gosset, known by the pen name ‘Student’, a pioneer of modern statistics in small sample experimental design and analysis [@Ziliak_2019].
 
@@
 The R package gosset is structured following the guidelines described in the manual for creating R add-on packages [@r_extensions_2022].
-This structure basically consist of files DESCRTIPTION, LICENSE, NAMESPACE and NEWS, and directories data, dev, docs, inst, man, R, and vignettes.
+This structure consists of files DESCRIPTION, LICENSE, NAMESPACE and NEWS, and directories data, dev, docs, inst, man, R, and vignettes.
 The package functions were developed following the S3 methods style [@r_extensions_2022] and are contained in the R sub-directory.
@@
 
-* `rank_binomial` transforms a ranking object into a binary comparisons, as required by package BradleyTerry2 [@BradleyTerry2].  
+* `rank_binomial` transforms a ranking object into a binary comparison, as required by package BradleyTerry2 [@BradleyTerry2].  
 * `rank_numeric` converts numeric values into rankings.
@@
 The default is `asceding = FALSE`.
-This function is useful when the data were collected as numerical observations, for instance, in a experiment measuring crop yield.
+This function is useful when the data have been  collected as numerical observations, for instance, in an experiment measuring crop yield.
 * `rank_tricot` transforms data in tricot format into PlackettLuce rankings.
@@
 * `btpermute` deviance-based forward variable selection [@lysen2009permuted] procedure for Bradley-Terry models.
-* `crossvalidation` performs k-fold cross-validation, where k could be specified by the user.The default is 10-fold.
+* `crossvalidation` performs k-fold cross-validation, where k could be specified by the user.
+The default is 10-fold.
 Folds can be provided as a vector for a custom cross-validation, such as blocked cross-validation.
@@
 * `plot` provides an alternative S3 method to `plot.pltree()` method implemented by the PlackettLuce package.  
-* `regret` computes the regret coefficients, the loss under the worst possible outcome. A common heuristic
+* `regret` computes the regret coefficients, the loss under the worst possible outcome; a common heuristic
 in risk assessment strategy [@Loomes_1982].  
-* `reliability` computes the probability of a set of items to outperform a reference item. A common
+* `reliability` computes the probability of a set of items outperforming a reference item; a common
 heuristic in plant breeding [@eskridge_1992].  
@@
 The farmers assessed which of the three varieties had the best and worst performance in eight traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, and taste).
-The farmers also provided their overall appreciation about the varieties, i.e., which variety had the best and the worst performance based on the overall performance considering all the traits.
+The farmers also provided their overall appreciation of the varieties, i.e., which variety had the best and the worst performance based on the overall performance considering all the traits.
 To analyze the data, we use the Plackett-Luce model implemented in the R package *PlackettLuce* [@Turner2020].
@@
 
-To start the analysis of the data, we transform the ordinal rankings into Plackett-Luce rankings (a sparse matrix) using the function `rank_numeric()`.
+To start the data analysis, we transform the ordinal rankings into Plackett-Luce rankings (a sparse matrix) using the function `rank_numeric()`.
 We run iteratively over the traits adding the rankings to a list called `R`.
-Since the varieties are ranked in an ascending order, with 1 being the higher ranked and 3 the lower ranked, we use the argument `asceding = TRUE` to indicate which order should be used.
+Since the varieties are ranked in an ascending order, with 1 being the higher ranked and 3 the lower ranked, we use the argument `ascending = TRUE` to indicate which order should be used.
 
@@
 This approach can be used, for example, to assess the drivers of farmers choices or to prioritize traits to be tested in a next stage of tricot trials (e.g.
-a lite version of tricot with no more than 4 traits to assess).
+a light version of tricot with no more than 4 traits to assess).
 We use the overall appreciation as the reference trait, and compare the Kendall tau with the other 8 traits.
@@
 Then, for each trait, we fit a Plackett-Luce model using the function `PlackettLuce()` from the package of the same name.
-This will allow us to continue the analysis of the trial data using the other functions in the package gosset.
+This will allow us to continue the trial data analysis using the other functions in the package gosset.
 
@@
 
-```{r worthmap, message=FALSE, eval=TRUE, echo=TRUE, fig.cap="Trait performance (log-worth) of bean varieties in Niragua.
+```{r worthmap, message=FALSE, eval=TRUE, echo=TRUE, fig.cap="Trait performance (log-worth) of bean varieties in Nicaragua.
 Variety 'Amadeus' is set as reference (log-worth = 0).
@@
 ```{r pltree3, message=FALSE, eval=TRUE, echo=FALSE, out.width="60%", fig.align='center', fig.cap="Effect of total rainfall (Rtotal) on yield of common beans in on-farm trials.
-Agroclimate variables are obtained from planting date to the first 45 days of plant growth.
+Agroclimate variables are obtained from planting date over the first 45 days of plant growth.
 Axis X presents log-worth, the probability of outperforming the other varieties in the set."}
@@
 varieties in each of the resulting nodes of the Plackett-Luce tree (Table 3).
-This helps in identifying the varieties with higher probability to outperform a variety check (Amadeus 77).
+This helps in identifying the varieties with higher probability of outperforming a check variety (Amadeus 77).
 For simplicity, we present only the varieties with reliability $\geq$ 0.5.
@@
 
-The result shows that three varieties can marginally outperform Amadeus 77 under drier growing conditions (Rtotal $\leq$ 193.82 mm) whereas two varieties have a superior yield performance when under higher rainfall conditions (Rtotal $>$ 193.82 mm) compared to the reference.
+The results show that three varieties can marginally outperform Amadeus 77 under drier growing conditions (Rtotal $\leq$ 193.82 mm) whereas two varieties have a superior yield performance when under higher rainfall conditions (Rtotal $>$ 193.82 mm) compared to the reference.
 This approach helps in identifying superior varieties for different target population environments.
@@
 
-A better approach to assess the performance of varieties can be using the 'Overall Appreciation", since we expect this trait to capture the performance of the variety not only for yield, but for all the other traits prioritized by farmers (Table 2).
+A better approach for assessing the performance of varieties can be using the "Overall Appreciation", since we expect this trait to capture the performance of the variety not only for yield, but for all the other traits prioritized by farmers (Table 2).
 To support this hypotheses, we use the function `compare()` which applies the approach proposed by Bland and Altman (1986) [@MartinBland1986] to assess the agreement between two different measures.
@@
 
-Here we presented a simple workflow to assess crop variety performance and trait prioritization in decentralized on-farm trials with the tricot approach.
-A more complex workflow will also utilize other functions available in gosset, for example, a forward selection combined with `crossvalidation()` to improve model robustness, or model selection with `btpermute()` to consider all possible permutations in Bradley-Terry models, or a risk analysis using `regret()` to support the selection of varieties, or using `rank_numeric()` to combine legacy data and deal with heterogeneous data from different trials.
+Here we present a simple workflow to assess crop variety performance and trait prioritization in decentralized on-farm trials with the tricot approach.
+A more complex workflow would also utilize other functions available in gosset, Examples include: (1) a forward selection combined with `crossvalidation()` to improve model robustness, or (2) model selection with `btpermute()` to consider all possible permutations in Bradley-Terry models, or (3) a risk analysis using `regret()` to support the selection of varieties, or also (4) using `rank_numeric()` to combine legacy data and deal with heterogeneous data from different trials.
 All of these were previously implemented and validated elsewhere [@vanEtten2019; @Moyo_2021; @deSousa2021; @Steinkemindata2019; @brownlegacydata].
@@
 The gosset package provides functionality that was not previously available from other R packages and which enabled scientific studies based on the analysis of ranking data.
-This functionality enables making the entire workflow to be reproducible and more efficient.
+This functionality allows reproducibility and greater efficiency of the entire workflow.
 The utility of the gosset package has been demonstrated by enabling studies based on the analysis of decentralized on-farm trial data and/or  heterogeneous data from different sources.
@@
 We described the functionality of the R package gosset to support the synthesis and analysis of ranking data.
-The package provide functions not available in existing R packages for analyzing ranking data.
+The package provides functions that are not available in existing R packages for analyzing ranking data.
 We provided an illustrative example covering the main functionality across the stages involved in the analysis workflow.
@@
 
-We acknowledge Olga Spellman (Science Writing Service of the Alliance of Bioversity International and CIAT) for English editing of this manuscript.
+We acknowledge Vincent Johnson (Science Writing Service of the Alliance of Bioversity International and CIAT) for English editing of this manuscript.
 The nicabean data was generated through the Cooperative Agreement AID-OAA-F-14-00035, which was made possible by the generous support of the American people through the US Agency for International Development (USAID).
-The package was developed as part of the CGIAR Research Programs on Climate Change, Agriculture and Food Security (CCAFS) and the CGIAR Research Program on Roots, Tubers and Bananas (RTB), which were carried out with support from the CGIAR Trust Fund and through bilateral funding agreements (details are at https://www.cgiar.org/funders).
+The package was developed as part of the CGIAR Research Program (CRP) on Climate Change, Agriculture and Food Security (CCAFS) and the (CRP) on Roots, Tubers and Bananas (RTB), which were carried out with support from the CGIAR Trust Fund and through bilateral funding agreements (details are at https://www.cgiar.org/funders).
 New analytical approaches were developed during the projects Accelerated Varietal Improvement and Seed Systems in Africa (AVISA) and 10000FARMS supported by the Bill and Melinda Gates Foundation.

### 2022-07-07 inst/paper/paper.Rmd
Update paper.Rmd

@@
 ^4^ Digital Inclusion, Bioversity International, 30501, Turrialba, Costa Rica   
-^5^ Humboldt University Berlin, Berlin, Germany  
+^5^ Thaer Institute of Agricultural and Horticultural Sciences, Humboldt University Berlin, Unter den Linden 6, 10099 Berlin, Germany  
 ^[*]^ Correspondence should be addressed to: <k.desousa@cgiar.org>

### 2022-07-08 inst/paper/paper.Rmd
Update paper.Rmd
Edited version

@@
 
-Appropriate data management and analysis are necessary for generating practical information from agricultural experimentation data.
-Modern approaches advocate using programmatic tools that support reproducible workflows in scientific research.
-We developed the R package gosset, which provides useful functionality for supporting analysis workflows with rank-based models.
+To derive insights from data, researchers working on agricultural experiments need appropriate data management and analysis tools.
+To ensure that workflows are reproducible and can be applied on a routine basis, programmatic tools are needed.
+Such tools are increasingly necessary for rank-based data, a type of data that is generated in on-farm experimentation and data synthesis exercises, among others.
+To address this need, we developed the R package gosset, which provides functionality for rank-based data and models.
 The gosset package facilitates data preparation, modelling and results presentation stages.
-It uses novel functions not available in existing R packages for analyzing ranking data.
+It introduces novel functions not available in existing R packages for analyzing ranking data.
 This paper demonstrates the package functionality using the case study of a decentralized on-farm trial of common bean (*Phaseolus vulgaris* L.) varieties in Nicaragua.
@@
 
-Participatory experimentation approaches have been increasingly applied in agricultural research [@deRoo2019farm].
-While collecting data in ranking format is uncommon in general agricultural research settings, it is often collected in participatory experiments [@Coe2002AnalyzingRA].
-Recently developed approaches for on-farm experimentation, such as the triadic comparison of technologies (tricot) approach are based on the collection of data in ranking format  [@vanetten_beza_2019].
-On the other hand, newly proposed approaches for synthesizing crop variety evaluation data largely depend on the analysis of ranking data [@brown2020data].
+Participatory on-farm experimentation approaches are reaching scale in agricultural research [@deRoo2019farm].
+Participatory experiments often collect data as rankings, a format that is less common in other agricultural research settings [@Coe2002AnalyzingRA].
+A recently developed approach for on-farm experimentation, triadic comparison of technologies (tricot), makes intensive use of data in ranking format [@vanetten_beza_2019].
+Also, a newly proposed approach for synthesizing crop variety evaluation data largely depends on the analysis of ranking data [@brown2020data].
 
-The analysis of ranking data requires the use of appropriate statistical models such as Plackett-Luce [@luce_individual_1959; @Plackett] or Bradley-Terry [@bradley1952rank].
-Functionality for fitting Bradley-Terry and Plackett-Luce models are available in R with the packages BradleyTerry2 [@BradleyTerry2] and PlackettLuce [@Turner2020] respectively.
+The analysis of ranking data requires the use of appropriate statistical models such as the Plackett-Luce model [@luce_individual_1959; @Plackett] or the Bradley-Terry model [@bradley1952rank].
+Functionality for fitting these models are available in R with the packages BradleyTerry2 [@BradleyTerry2] and PlackettLuce [@Turner2020] respectively.
 However, extended functionality was required for the entire data science workflow, which usually includes: (1) Data preparation and cleaning, (2) modelling and validation, and (3) results presentation.
@@
 
-The R package gosset provides functionality supporting the analysis workflows in agricultural experimentation, especially with rank-based approaches.
+The R package gosset provides functionality supporting the analysis workflows in agricultural experimentation, especially for rank-based approaches.
 The package is available in the Comprehensive R Archive Network (CRAN) and can be installed by executing `install.packages("gosset")`.
-The package is named in honor of William Sealy Gosset, known by the pen name ‘Student’, a pioneer of modern statistics in small sample experimental design and analysis [@Ziliak_2019].
+The package is named in honor of William Sealy Gosset, known by the pen name ‘Student’.
+Gosset was a pioneer of modern statistics in small sample experimental design and analysis.
+As a beer brewer at Guinness, he developed practical approaches to experimentation to compare barley varieties and beer brewing practices [@Ziliak_2019].
 
@@
 Following the tricot  approach [@vanetten_beza_2019], farmers were asked to test in their farms three varieties of common bean.
-The varieties were randomly assigned as incomplete blocks of size three (out of 10 varieties).
-The farmers assessed which of the three varieties had the best and worst performance in eight traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, and taste).
-The farmers also provided their overall appreciation of the varieties, i.e., which variety had the best and the worst performance based on the overall performance considering all the traits.
+The varieties were randomly assigned as incomplete blocks, each representing 3 varieties out of a total set of 10 varieties.
+Each farmer assessed which of the three varieties in one incomplete block had the best and worst performance in eight traits (vigor, architecture, resistance to pests, resistance to diseases, tolerance to drought, yield, marketability, and taste).
+The farmers also provided their overall appreciation of the varieties, by indicating which variety had the best and the worst performance based on the overall performance considering all the traits.
 To analyze the data, we use the Plackett-Luce model implemented in the R package *PlackettLuce* [@Turner2020].
 
-The `nicabean` is a list with two data frames.
+The `nicabean` dataset is a list with two data frames.
 The first, `trial`, contains the trial data with farmers’ evaluations, ranked from 1 to 3, with 1 being the higher ranked variety and 3 the lowest ranked variety for the given trait and incomplete block.
@@
 
-To start the data analysis, we transform the ordinal rankings into Plackett-Luce rankings (a sparse matrix) using the function `rank_numeric()`.
+To start the data analysis, we transform the ordinal rankings into the Plackett-Luce rankings format (a sparse matrix) using the function `rank_numeric()`.
 We run iteratively over the traits adding the rankings to a list called `R`.
@@
 
-Then, using the function `kendallTau()` we can assess the Kendall tau ($\tau$) coefficient [@kendall_1938].
-This approach can be used, for example, to assess the drivers of farmers choices or to prioritize traits to be tested in a next stage of tricot trials (e.g.
-a light version of tricot with no more than 4 traits to assess).
+Then, using the function `kendallTau()` we assess the Kendall tau ($\tau$) coefficient [@kendall_1938].
+This approach can be used, for example, to assess what traits influence farmers' choices or to prioritize traits to be tested in a next stage of tricot trials (e.g.
+a lighter version of tricot with no more than 4 traits to assess).
 We use the overall appreciation as the reference trait, and compare the Kendall tau with the other 8 traits.
@@
 
-The kendall correlation shows that farmers prioritized the traits yield ($\tau$ = 0.749), taste ($\tau$ = 0.653) and marketability ($\tau$ = 0.639) when assessing overall appreciation.
+The Kendall correlation shows that farmers prioritized the traits yield ($\tau$ = 0.749), taste ($\tau$ = 0.653) and marketability ($\tau$ = 0.639) when assessing overall appreciation.
 
@@
 Agroclimate variables are obtained from planting date over the first 45 days of plant growth.
-Axis X presents log-worth, the probability of outperforming the other varieties in the set."}
+The x axis presents log-worth, the log-probability of outperforming the other varieties in the set."}
 knitr::include_graphics("pltree_01.png")
@@
 This helps in identifying the varieties with higher probability of outperforming a check variety (Amadeus 77).
-For simplicity, we present only the varieties with reliability $\geq$ 0.5.
+For the sake of simplicity, we present only the varieties with reliability $\geq$ 0.5.
 
@@
 A better approach for assessing the performance of varieties can be using the "Overall Appreciation", since we expect this trait to capture the performance of the variety not only for yield, but for all the other traits prioritized by farmers (Table 2).
-To support this hypotheses, we use the function `compare()` which applies the approach proposed by Bland and Altman (1986) [@MartinBland1986] to assess the agreement between two different measures.
+To assess this, we use the function `compare()` which applies the approach proposed by Bland and Altman (1986) [@MartinBland1986] to assess the agreement between two different measures.
 We compare overall vs yield.

### 2022-07-08 inst/paper/paper.Rmd
Update paper.Rmd
minor edits

@@
 The analysis of ranking data requires the use of appropriate statistical models such as the Plackett-Luce model [@luce_individual_1959; @Plackett] or the Bradley-Terry model [@bradley1952rank].
-Functionality for fitting these models are available in R with the packages BradleyTerry2 [@BradleyTerry2] and PlackettLuce [@Turner2020] respectively.
+Functionality for fitting these models is available in R with the packages BradleyTerry2 [@BradleyTerry2] and PlackettLuce [@Turner2020] respectively.
 However, extended functionality was required for the entire data science workflow, which usually includes: (1) Data preparation and cleaning, (2) modelling and validation, and (3) results presentation.
@@
 To consider the effect of climate factors on yield, we use agro-climatic covariates to fit a Plackett-Luce tree.
-For simplicity, we use the total rainfall (Rtotal) derived from CHIRPS data [@Funk2015], obtained in R using the R package chirps [@chirps].
+For simplicity, we use the total rainfall (Rtotal) derived from CHIRPS data [@Funk2015], obtained using the R package chirps [@chirps].
 Additional covariates can be used in a Plackett-Luce tree, for example using temperature data from R package ag5Tools [@ag5tools] or nasapower [@nasapower].

### 2022-09-23 R/kendallW.R
v0.5

@@ -16,22 +16,6 @@
 #' @param ... additional arguments passed to methods
 #' @return Kendall's W (coefficient of concordance) 
 #' @details 
-#' @examples
-#' 
-#' library("PlackettLuce")
-#' G <- rank_tricot(data = beans,
-#'                  items = c(1:3),
-#'                  input = c(4:5),
-#'                  additional.rank = beans[c(6:8)],
-#'                  group = TRUE)
-#'                  
-#' beans <- cbind(G, beans)
-#' 
-#' beans_plt <- pltree(G ~ maxTN,
-#'                     data = beans)
-#'                    
-#' 
-#' kendallW(beans_plt)
 #' @importFrom DescTools KendallW
 #' @export
 kendallW <- function(x, ...) {
@@ -92,5 +76,4 @@ kendallW.pltree <- function(x, newdata = NULL, ...) {
    kendallW(pred_ranks, obs_ranks_mtx)
   
 }
-
 

### 2022-09-23 R/reliability.R

@@ -39,6 +39,7 @@
 #' 
 #' reliability(mod, ref = "orange")
 #' 
+#' \donttest{
 #' # Case 3. PlackettLuce tree
 #' 
 #' data("beans", package = "PlackettLuce")
@@ -53,8 +54,9 @@
 #' 
 #' tree <- pltree(G ~ maxTN + season + lon, data = pld)
 #' 
-#' reliability(tree, ref = "INTA Matagalpa")
+#' reliability(tree, ref = "Local")
 #' 
+#' }
 #' @importFrom methods addNextMethod asMethodDefinition assignClassDef
 #' @importFrom qvcalc qvcalc
 #' @importFrom stats update pnorm

### 2023-04-10 R/kendallTau.R
v1.0
rename functions, add p-val to kendallTau

@@ -143,7 +143,17 @@ kendallTau.default = function(x, y, null.rm = TRUE, ...){
   
   kt[2] = N_effective
   
-  names(kt) = c("kendallTau", "N_effective")
+  n = N_effective
+  
+  # calculate z-value using n effective
+  z = (3 * tau_cor * sqrt(n * (n - 1))) / sqrt(2 * (2 * n + 5))
+  
+  # then p-value
+  p = stats::pnorm(z, lower.tail = FALSE)
+  
+  kt = c(kt, z, p)
+  
+  names(kt) = c("kendallTau", "N_effective", "Zvalue", "Pr(>|z|)")
   
   kt = t(as.data.frame(kt))
   
@@ -195,9 +205,17 @@ kendallTau.matrix = function(x, y, null.rm = TRUE, average = TRUE, na.omit = FAL
   # This is used for significance testing later
   N_effective = 0.5 + sqrt(0.25 + 2 * sum(N)) 
   
-  kt = c(tau_average, N_effective)
+  n = N_effective
   
-  names(kt) = c("kendallTau", "N_effective")
+  # calculate z-value using n effective
+  z = (3 * tau_average * sqrt(n * (n - 1))) / sqrt(2 * (2 * n + 5))
+  
+  # then p-value
+  p = stats::pnorm(z, lower.tail = FALSE)
+  
+  kt = c(tau_average, N_effective, z, p)
+  
+  names(kt) = c("kendallTau", "N_effective", "Zvalue", "Pr(>|z|)")
   
   kt = t(as.data.frame(kt))
   

### 2023-04-10 R/rank_paircomp.R

@@ -1,7 +1,7 @@
 #' Pairwise comparison rankings
 #'
-#' Pairwise comparisons from a ranking object. Ties are not 
-#' taken into account, then they are added as NA's.
+#' Pairwise comparisons from a PlackettLuce ranking object. 
+#'  Ties are not taken into account, then coerced to NA's.
 #'
 #' @author Kauê de Sousa and Jacob van Etten
 #' @family rank functions
@@ -43,12 +43,12 @@
 #'                  input = c("overall_best","overall_worst"))
 #' 
 #' 
-#' PC = rank_paircomp(R)
+#' PC = set_paircomp(R)
 #' 
 #' @importFrom psychotools paircomp
 #' @importFrom utils combn
 #' @export
-rank_paircomp = function(object){
+set_paircomp = function(object){
   
   R = object
   

### 2023-04-10 dev/distil-rankings.R

@@ -1,0 +1,66 @@
+library("ClimMobTools")
+library("gosset")
+library("PlackettLuce")
+
+data("breadwheat")
+
+head(breadwheat)
+
+dat = breadwheat
+
+dat$id = 1:nrow(dat)
+
+# get the list of traits in breadwheat. The pattern for the data is c("_best", "_worst")
+traits = getTraitList(dat, pattern = c("_best", "_worst"))
+
+pack_index = paste0("variety_", letters[1:3])
+
+# now transform into PlackettLuce rankings
+# if you use group = TRUE the ranking can be used for PL trees
+R = lapply(traits, function(x){
+  rank_tricot(dat, 
+              items = pack_index,
+              input = x$string,
+              group = FALSE)
+})
+
+# Now you have the PL rankings that can be used to fit PL model 
+mod = lapply(R, PlackettLuce)
+
+lapply(mod, summary)
+
+# Now you can convert to ranking 
+# (not really needed, only for data storage or publication)
+rank_data = data.frame()
+
+for (i in seq_along(traits)) {
+  
+  r = unclass(R[[i]])
+  
+  for (j in seq_along(dat$id)) {
+    
+    id = dat$id[j]
+    
+    plots = as.vector(unlist(dat[dat$id == id, pack_index]))
+    
+    x = r[j, plots]
+    
+    d = data.frame(id = id, 
+                   plot = LETTERS[1:3],
+                   tech_name = plots,
+                   trait = as.vector(traits[[i]]$trait_label),
+                   rank = x)
+    
+    rank_data = rbind(rank_data, d)
+    
+  }
+  
+}
+
+rank_data
+
+rownames(rank_data) = 1:nrow(rank_data)
+
+head(rank_data)
+
+

### 2023-04-10 dev/rankings-on-climmob.R

@@ -1,0 +1,35 @@
+library("gosset")
+library("ClimMobTools")
+library("PlackettLuce")
+
+tech = c("apple", "banana", "pear", "orange", "grape")
+
+set.seed(1951)
+R = randomise(6, tech)
+
+R$best = c("C", "B", "A", "B", "Not observed", "A")
+R$worst = c("A", "Tie", "C", "Not observed", "Not observed", "C")
+
+R
+
+r1 = rank_tricot(R, items = 1:3, input = 4:5)
+
+r1 = unclass(r1)
+
+d = data.frame()
+for (i in seq_len(nrow(r1))) {
+  x = r1[i, ]
+  x = x[x!=0]
+  x = data.frame(id = i,
+                 tech = names(x),
+                 rank = x)
+  d = rbind(d, x)
+}
+
+rownames(d) = 1:nrow(d)
+
+d
+
+rank_tricot(R, items = 1:3, input = 4:5, validate.rankings = T)
+
+r2

### 2023-04-10 vignettes/Overview.R

@@ -44,8 +44,8 @@ kendall$trait <- traits[-baseline]
 
 ## ----kendall2, message=FALSE, eval=TRUE, echo=FALSE---------------------------
 
-kendall <- kendall[,c(3, 1)]
-kendall[,2] <- round(kendall[,2], 3)
+kendall <- kendall[,c(5, 1, 2, 3)]
+kendall[,2:4] <- lapply(kendall[,2:4], function(x) round(x, 3))
 
 kendall
 

### 2024-10-08 R/cassava.R
v1.4

@@ -1,0 +1,68 @@
+#' Gari-Eba Consumer Acceptability in Cameroon and Nigeria
+#'
+#' Data from decentralized consumer trials of gari-eba, a sub-product of 
+#' cassava (\emph{Manihot esculenta} Crantz). Consumer testing was carried 
+#' out in 2022 in Cameroon and Nigeria using the tricot approach. Diverse consumers in villages,
+#' towns and cities evaluated the overall acceptability of 
+#' gari-eba made from 13 cassava genotypes. Apart from overall preference of the 
+#' samples, the following traits were evaluated for eba based on the 
+#' triangulated insights obtained by earlier survey and participatory work in 
+#' the three areas. Nigeria (Osun and Benue States): colour, smoothness,
+#' mouldability, stretchability, and taste. Cameroon (Littoral zone): colour,
+#' odour, taste, firmness and stretchability. Traits in common are: colour,
+#' taste and stretchability.
+#' 
+#' @keywords dataset
+#' @format A data frame with 1000 records and 27 variables
+#' \describe{
+#'     \item{\code{id}}{The tricot package id.}
+#'     \item{\code{option_a}}{The name of genotype A in the comparison.}
+#'     \item{\code{option_b}}{The name of genotype B in the comparison.}
+#'     \item{\code{option_c}}{The name of genotype C in the comparison.}
+#'     \item{\code{country}}{The country where the experiment was conducted.}
+#'     \item{\code{gender}}{The participant gender.}
+#'     \item{\code{age}}{The participant age.}
+#'     \item{\code{consumption}}{Indicates how often the participant consumes eba.}
+#'     \item{\code{consumptionform}}{Indicates in which meal the participant consumes eba.}
+#'     \item{\code{colour_pos}}{The sample ranked as best for colour ("A",
+#'     "B" or "C").}
+#'     \item{\code{colour_neg}}{The sample ranked as worst for colour ("A",
+#'     "B" or "C").}
+#'     \item{\code{smoothness_pos}}{The sample ranked as best for smoothness ("A",
+#'     "B" or "C").}
+#'     \item{\code{smoothness_neg}}{The sample ranked as worst for smoothness ("A",
+#'     "B" or "C").}
+#'     \item{\code{mouldability_pos}}{The sample ranked as best for mouldability ("A",
+#'     "B" or "C").}
[…]

### 2024-10-08 vignettes/vignette-1-trait-prioritization-and-crop-performance.R

@@ -57,7 +57,7 @@ kendall
 mod = lapply(R, PlackettLuce)
 
 
-## ----worthmap, message=FALSE, eval=TRUE, echo=TRUE, fig.cap="Trait performance (log-worth) of bean varieties in Niragua. Variety 'Amadeus' is set as reference (log-worth = 0). Blue values indicate a superior performance of varieties for a given trait, compared to the reference.  Red values indicate a variety with weak performance for the given trait, compared to the reference."----
+## ----worthmap, message=FALSE, eval=TRUE, echo=TRUE, fig.cap="Trait performance (log-worth) of bean varieties in Niragua. Variety 'Amadeus' is set as reference (log-worth = 0). Green values indicate a superior performance of varieties for a given trait.  Brown values indicate a variety with weak performance for the given trait"----
 worth_map(mod[-baseline],
           labels = traits[-baseline],
           ref = "Amadeus 77") +

### 2024-11-21 dev/analysis-cassava.R
cassava-vignette

@@ -3,11 +3,9 @@ library("ClimMobTools")
 library("PlackettLuce")
 library("ggplot2")
 library("patchwork")
-library("multcompView")
+
 
 data("cassava", package = "gosset")
-
-#dat = cassava[cassava$country == "Nigeria", ]
 
 dat = cassava
 
@@ -21,25 +19,29 @@ dat = dat[, keep]
 
 names(dat)
 
-pack = c("option_a", "option_b", "option_c")
-
-items = sort(unique(unlist(dat[pack])))
-
-items
-
 trait_list = getTraitList(dat, c("_pos", "_neg"))
 
 traits = unlist(lapply(trait_list, function(x) x$trait_label))
 
 traits = gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", traits, perl = TRUE)
 
+traits
+
+
+
+pack = c("option_a", "option_b", "option_c")
+
+items = sort(unique(unlist(dat[pack])))
+
+ref = "Akpu"
+
 ov = which(traits %in% "Overall")
+
+
 
 table(unlist(dat[pack]), rep(dat$gender, 3))
 
 table(unlist(dat[pack]), rep(dat$country, 3))
-
-country = sort(unique(dat$country))
 
 R = lapply(trait_list, function(x) {
   rank_tricot(dat, 
@@ -48,55 +50,13 @@ R = lapply(trait_list, function(x) {
               validate.rankings = TRUE)
 })
 
-lapply(R, function(x){
-  likelihood_ratio(x, split = dat$gender)
-})
+# Assess the full data
+mod = lapply(R, PlackettLuce)
 
-lapply(R, function(x){
-  likelihood_ratio(x, split = dat$country)
-})
-
-
-G = lapply(trait_list, function(x) {
-  rank_tricot(dat, 
-              items = pack, 
-              input = x$string,
-              validate.rankings = TRUE,
-              group = TRUE)
-})
-
-
-trait_plot = list()
-
-for (i in seq_along(country)) {
-  
-  mod_i = lapply(R, function(x) {
-    PlackettLuce(x[dat$country == country[i], ])
-  })
-  
-  trait_plot[[i]] = worth_map(mod_i, 
-            labels = traits) +
-    labs(x = "", 
-         y = "",
-         title = country[i]) +
-    scale_fill_distiller(palette = "BrBG", 
-                         direction = 1, 
-                         na.value = "white", 
[…]

### 2024-11-21 dev/analysis-cassava.R
update-commaon-bean-vignette

@@ -72,24 +72,23 @@ ggplot(data = rel,
            width = 0.7,
            position = "dodge", 
            show.legend = FALSE) +
-  scale_fill_manual(values = "#b2df8a") +
+  scale_fill_manual(values = "#5aae61") +
   geom_vline(xintercept = 0.5,
-             colour = "#1f78b4",
+             colour = "grey40",
              linewidth = 1) +
-  theme_bw() +
-  theme(panel.grid.major = element_blank(),
-        strip.background =element_rect(fill="white"),
-        text = element_text(color = "grey20"),
-        strip.background.x = element_blank(),
-        strip.placement = "outside",
-        legend.position = "bottom",
-        strip.text = element_text(size = 12, color = "grey20"),
-        legend.text = element_text(size = 12, color = "grey20"),
-        axis.text = element_text(size = 12, color = "grey20"),
-        axis.title = element_text(size = 12, color = "grey20"),
-        legend.title = element_blank()) +
+  theme_classic() +
   labs(x = "Probability of outperforming",
        y = "")
+
+
+worth_bar(mod[[ov]])
+
+worth_map(mod, labels = traits) +
+  labs(x = "", y = "") +
+  scale_fill_distiller(palette = "BrBG", 
+                       direction = 1, 
+                       na.value = "white", 
+                       name = "")
 
 
 # Slice the data
@@ -143,7 +142,7 @@ for (i in seq_along(slice_lvs)) {
 }
 
 
-trait_plot[[1]] + trait_plot[[2]]
+trait_plot[[1]] + trait_plot[[2]] + plot_layout(ncol = 1)
 
 
 
@@ -186,18 +185,9 @@ ggplot(data = select,
   geom_vline(xintercept = mean(select$score),
              colour = "grey40",
              linewidth = 1) +
-  theme_bw() +
-  theme(panel.grid.major = element_blank(),
-        strip.background = element_rect(fill="white"),
-        text = element_text(color = "grey20"),
-        strip.background.x = element_blank(),
-        strip.placement = "outside",
-        legend.position = "bottom",
-        strip.text = element_text(size = 12, color = "grey20"),
-        legend.text = element_text(size = 12, color = "grey20"),
-        axis.text = element_text(size = 12, color = "grey20"),
-        axis.title = element_text(size = 12, color = "grey20"),
-        legend.title = element_blank()) +
+  theme_classic() +
+  theme(legend.title = element_blank(),
+        legend.position = "bottom") +
   labs(x = "Selection score",
        y = "")
 

### 2024-11-21 vignettes/vignette-1-trait-prioritization-and-crop-performance.R
update-site

@@ -57,7 +57,7 @@ kendall
 mod = lapply(R, PlackettLuce)
 
 
-## ----worthmap, message=FALSE, eval=TRUE, echo=TRUE, fig.cap="Trait performance (log-worth) of bean varieties in Niragua. Variety 'Amadeus' is set as reference (log-worth = 0). Green values indicate a superior performance of varieties for a given trait.  Brown values indicate a variety with weak performance for the given trait"----
+## ----worthmap, message=FALSE, eval=TRUE, echo=TRUE----------------------------
 worth_map(mod[-baseline],
           labels = traits[-baseline],
           ref = "Amadeus 77") +
@@ -133,5 +133,5 @@ Yield = PlackettLuce(R[[yield]])
 
 compare(Overall, Yield) +
   labs(x = "Average log(worth)",
-       y = "Difference (Overall Appreciation - Yield)")
+       y = "Difference (Overall appreciation - Yield)")
 

### 2024-11-21 vignettes/vignette-2-gari-eba-consumer-testing.R

@@ -1,0 +1,205 @@
+## ----setup, include=FALSE-----------------------------------------------------
+knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
+
+## ----starting, message = FALSE, eval = TRUE, echo = TRUE----------------------
+library("gosset")
+library("ClimMobTools")
+library("PlackettLuce")
+library("ggplot2")
+library("patchwork")
+
+data("cassava", package = "gosset")
+
+dat = cassava
+
+head(dat[, 1:11])
+
+
+## ----select, message = FALSE, eval = TRUE, echo = TRUE------------------------
+keep = unlist(lapply(dat[1:ncol(dat)], function(x) sum(is.na(x))))
+
+keep = keep == 0
+
+dat = dat[, keep]
+
+names(dat)
+
+
+## ----traitlist, message = FALSE, eval = TRUE, echo = TRUE---------------------
+# extract list of traits from the data
+trait_list = getTraitList(dat, pattern = c("_pos", "_neg"))
+
+# trait names extracted from the function 
+traits = unlist(lapply(trait_list, function(x) x$trait_label))
+
+# clean trait names and put them title case
+traits = gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", traits, perl = TRUE)
+
+traits
+
+## ----prep, message = FALSE, eval = TRUE, echo = TRUE--------------------------
+pack = c("option_a", "option_b", "option_c")
+
+items = sort(unique(unlist(dat[pack])))
+
+ref = "Akpu"
+
+ov = which(traits %in% "Overall")
+
+## ----rank, message = FALSE, eval = TRUE, echo = TRUE--------------------------
+R = lapply(trait_list, function(x) {
+  rank_tricot(dat, 
+              items = pack, 
+              input = x$string,
+              validate.rankings = TRUE)
+})
+
+## ----plfit, message = FALSE, eval = TRUE, echo = TRUE-------------------------
+mod = lapply(R, PlackettLuce)
+
+plot(mod[[ov]],
+     ref = ref,
+     log = TRUE,
+     levels = rev(items))
+
+
+## ----reliability, message = FALSE, eval = TRUE, echo = TRUE-------------------
+rel = reliability(mod[[ov]], ref = "Akpu")
+
+ggplot(data = rel,
+       aes(x = reliability, 
+           y = item,
+           fill = "white")) +
+  geom_bar(stat = "identity",
+           width = 0.7,
[…]

### 2024-11-22 vignettes/vignette-2-gari-eba-consumer-testing.R
update-cassava-vignette

@@ -46,6 +46,23 @@ ref = "Akpu"
 
 ov = which(traits %in% "Overall")
 
+table(unlist(dat[pack]), rep(dat$country, 3))
+
+table(unlist(dat[pack]), rep(dat$gender, 3))
+
+# rmv = data.frame(table(unlist(dat[pack]), rep(dat$country, 3)))
+# 
+# rmv = as.vector(rmv[rmv$Freq == 0, "Var1"])
+# 
+# dat[pack] = lapply(dat[pack], function(x){
+#   x[x %in% rmv] = NA
+#   x
+# })
+# 
+# keep = colSums(apply(dat[pack], 1, is.na)) < 2
+# 
+# dat = dat[keep, ]
+
 ## ----rank, message = FALSE, eval = TRUE, echo = TRUE--------------------------
 R = lapply(trait_list, function(x) {
   rank_tricot(dat, 
@@ -58,13 +75,13 @@ R = lapply(trait_list, function(x) {
 mod = lapply(R, PlackettLuce)
 
 plot(mod[[ov]],
-     ref = ref,
+     ref = 1,
      log = TRUE,
      levels = rev(items))
 
 
 ## ----reliability, message = FALSE, eval = TRUE, echo = TRUE-------------------
-rel = reliability(mod[[ov]], ref = "Akpu")
+rel = reliability(mod[[ov]], ref = "TMS1")
 
 ggplot(data = rel,
        aes(x = reliability, 

### 2024-11-28 dev/kendall_permute.R
update-vignette-2

@@ -1,0 +1,50 @@
+# Load necessary library
+library(stats)
+
+# Example data
+data <- data.frame(
+  group = c(rep("Group1", 50), rep("Group2", 50)),
+  trait = rnorm(100),  # Replace with actual trait values
+  overall = rnorm(100) # Replace with actual "Overall" values
+)
+
+# Function to calculate Kendall's Tau for a group
+calculate_tau <- function(group_data) {
+  cor(group_data$trait, group_data$overall, method = "kendall")
+}
+
+# Observed Kendall's Tau for each group
+tau_group1 <- calculate_tau(data[data$group == "Group1", ])
+tau_group2 <- calculate_tau(data[data$group == "Group2", ])
+observed_diff <- abs(tau_group1 - tau_group2)
+
+# Permutation test
+set.seed(123)  # For reproducibility
+n_permutations <- 1000
+permuted_diffs <- numeric(n_permutations)
+
+for (i in 1:n_permutations) {
+  # Shuffle group labels
+  shuffled_data <- data
+  shuffled_data$group <- sample(shuffled_data$group)
+  
+  # Recalculate Kendall's Tau for permuted groups
+  tau_perm_group1 <- calculate_tau(shuffled_data[shuffled_data$group == "Group1", ])
+  tau_perm_group2 <- calculate_tau(shuffled_data[shuffled_data$group == "Group2", ])
+  
+  # Calculate absolute difference in permuted Kendall's Tau
+  permuted_diffs[i] <- abs(tau_perm_group1 - tau_perm_group2)
+}
+
+# P-value: Proportion of permuted differences greater than or equal to observed difference
+p_value <- mean(permuted_diffs >= observed_diff)
+
+# Output results
+cat("Observed Difference in Kendall's Tau:", observed_diff, "\n")
+cat("P-value from Permutation Test:", p_value, "\n")
+
+# Visualize permutation test results
+hist(permuted_diffs, breaks = 30, main = "Permutation Test Results",
+     xlab = "Difference in Kendall's Tau")
+abline(v = observed_diff, col = "red", lwd = 2, lty = 2)
+legend("topright", legend = c("Observed Difference"), col = "red", lwd = 2, lty = 2)

### 2024-11-28 vignettes/vignette-2-gari-eba-consumer-testing.R

@@ -50,19 +50,6 @@ table(unlist(dat[pack]), rep(dat$country, 3))
 
 table(unlist(dat[pack]), rep(dat$gender, 3))
 
-# rmv = data.frame(table(unlist(dat[pack]), rep(dat$country, 3)))
-# 
-# rmv = as.vector(rmv[rmv$Freq == 0, "Var1"])
-# 
-# dat[pack] = lapply(dat[pack], function(x){
-#   x[x %in% rmv] = NA
-#   x
-# })
-# 
-# keep = colSums(apply(dat[pack], 1, is.na)) < 2
-# 
-# dat = dat[keep, ]
-
 ## ----rank, message = FALSE, eval = TRUE, echo = TRUE--------------------------
 R = lapply(trait_list, function(x) {
   rank_tricot(dat, 

### 2024-11-28 vignettes/vignette-2-gari-eba-consumer-testing.R
add-kendall_permute()

@@ -42,13 +42,10 @@ pack = c("option_a", "option_b", "option_c")
 
 items = sort(unique(unlist(dat[pack])))
 
-ref = "Akpu"
+check = "Obasanjo-2"
 
 ov = which(traits %in% "Overall")
 
-table(unlist(dat[pack]), rep(dat$country, 3))
-
-table(unlist(dat[pack]), rep(dat$gender, 3))
 
 ## ----rank, message = FALSE, eval = TRUE, echo = TRUE--------------------------
 R = lapply(trait_list, function(x) {
@@ -62,16 +59,22 @@ R = lapply(trait_list, function(x) {
 mod = lapply(R, PlackettLuce)
 
 plot(mod[[ov]],
-     ref = 1,
+     ref = check,
      log = TRUE,
      levels = rev(items))
 
 
 ## ----reliability, message = FALSE, eval = TRUE, echo = TRUE-------------------
-rel = reliability(mod[[ov]], ref = "TMS1")
+rel = reliability(mod[[ov]], ref = check)
+
+rel$improvement = round((rel$reliability / 0.5 - 1), 2)
+
+rel = rel[order(rel$improvement), ]
+
+rel$item = factor(rel$item, levels = rel$item)
 
 ggplot(data = rel,
-       aes(x = reliability, 
+       aes(x = improvement, 
            y = item,
            fill = "white")) +
   geom_bar(stat = "identity",
@@ -79,11 +82,11 @@ ggplot(data = rel,
            position = "dodge", 
            show.legend = FALSE) +
   scale_fill_manual(values = "#5aae61") +
-  geom_vline(xintercept = 0.5,
+  geom_vline(xintercept = 0,
              colour = "grey40",
              linewidth = 1) +
   theme_classic() +
-  labs(x = "Probability of outperforming",
+  labs(x = "Probability of outperforming the check",
        y = "")
 
 
@@ -155,9 +158,35 @@ for (i in seq_along(slice_lvs)) {
 trait_plot[[1]] + trait_plot[[2]] + plot_layout(ncol = 1)
 
 
+## ----kendall1, message = FALSE, eval = TRUE, echo = TRUE----------------------
+
+kendall = lapply(R, function(x){
+  kendallTau(x, R[[ov]])
+})
+
+kendall = do.call("rbind", kendall)
+
+kendall$trait = traits
+
+kendall
+
+
[…]

### 2024-11-30 R/worth_map.R
update-vignette

@@ -8,6 +8,8 @@
 #'  an object of class \code{PlackettLuce} or
 #'  a list objects of class \code{PlackettLuce}
 #' @param labels a vector with the name of models in \var{object}
+#' @param labels.order optional, a vector to determine the order of labels
+#' @param items.order optional, a vector to determine the order of items
 #' @param ... additional arguments passed to methods
 #' @examples 
 #' library("psychotree")
@@ -106,12 +108,30 @@ worth_map.default = function(object, ...) {
 #' @method worth_map list
 #' @rdname worth_map
 #' @export
-worth_map.list = function(object, labels, ...) {
+worth_map.list = function(object, 
+                          labels, 
+                          labels.order = NULL, 
+                          items.order = NULL, ...) {
   
-  winprobs = .combine_coeffs(object, log = TRUE, vcov = FALSE, ...)
+  if (is.null(labels.order)) {
+    lvls = labels
+  } else {
+    lvls = labels.order
+  }
+  
+  winprobs = .combine_coeffs(object, log = TRUE, ...)
   
   # add name of features
   names(winprobs) = labels
+  
+  if (is.null(items.order)) {
+    order_items = order(exp(winprobs[,lvls[[1]]]))
+    order_items = rownames(winprobs)[order_items]
+  } else {
+    order_items = items.order
+  }
+  
+  
   
   winprobs = data.frame(items = rep(dimnames(winprobs)[[1]], 
                                     times = ncol(winprobs)),
@@ -119,7 +139,10 @@ worth_map.list = function(object, labels, ...) {
                                      each = nrow(winprobs)),
                         winprob = as.numeric(unlist(winprobs)))
   
-  winprobs$labels = factor(winprobs$labels, levels = labels)
+  winprobs$labels = factor(winprobs$labels, levels = rev(lvls))
+  
+  winprobs$items = factor(winprobs$items, 
+                          levels = rev(order_items))
   
   items = winprobs$items
   winprob = winprobs$winprob
@@ -142,16 +165,13 @@ worth_map.list = function(object, labels, ...) {
                                   direction = 1,
                                   na.value = "white",
                                   name = "") +
-    ggplot2::theme_bw() +
-    theme(axis.text = ggplot2::element_text(color = "grey20"),
-          strip.text.x = ggplot2::element_text(color = "grey20"),
[…]

### 2024-11-30 dev/dominance-analysis.R

@@ -1,0 +1,82 @@
+library(dominanceanalysis)
+library(caTools)
+library("gosset")
+library("ClimMobTools")
+library("PlackettLuce")
+library("ggplot2")
+library("patchwork")
+
+data("cassava", package = "gosset")
+
+dat = cassava
+
+head(dat[, 1:11])
+
+keep = unlist(lapply(dat[1:ncol(dat)], function(x) sum(is.na(x))))
+
+keep = keep == 0
+
+dat = dat[, keep]
+
+names(dat)
+
+# extract list of traits from the data
+trait_list = getTraitList(dat, pattern = c("_pos", "_neg"))
+
+# trait names extracted from the function 
+traits = unlist(lapply(trait_list, function(x) x$trait_label))
+
+# clean trait names and put them title case
+traits = gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", traits, perl = TRUE)
+
+traits
+
+pack = c("option_a", "option_b", "option_c")
+
+items = sort(unique(unlist(dat[pack])))
+
+check = "Obasanjo-2"
+
+ov = which(traits %in% "Overall")
+
+R = lapply(trait_list, function(x) {
+  rank_tricot(dat, 
+              items = pack, 
+              input = x$string,
+              validate.rankings = TRUE)
+})
+
+mod = lapply(R, PlackettLuce)
+
+worth = lapply(mod, function(x){
+  z = resample(x, log = TRUE, bootstrap = TRUE, n1 = 10, seed = 1432)$estimate
+  #log(z)
+})
+
+worth = as.data.frame(do.call("cbind", worth))
+
+names(worth) = traits
+
+head(worth)
+
+plot(worth$Colour, worth$Overall)
+plot(worth$Stretchability, worth$Overall)
+plot(worth$Taste, worth$Overall)
+
+modpres = lm(Overall ~ Taste + Stretchability + Colour, data = worth)
+
+summary(modpres)
+
+dapres = dominanceAnalysis(modpres)
+
+getFits(dapres,"r2")
+
+dominanceMatrix(dapres, type="complete",fit.functions = "r2", ordered=TRUE)
+
+contributionByLevel(dapres,fit.functions="r2")
+
+plot(dapres, which.graph ="conditional",fit.function = "r2") + theme_minimal()
+
+dom = averageContribution(dapres,fit.functions = "r2")
+
+

### 2024-11-30 vignettes/vignette-2-gari-eba-consumer-testing.R

@@ -91,12 +91,9 @@ ggplot(data = rel,
 
 
 ## ----worth, message = FALSE, eval = TRUE, echo = TRUE-------------------------
-worth_map(mod, labels = traits) +
-  labs(x = "", y = "") +
-  scale_fill_distiller(palette = "BrBG", 
-                       direction = 1, 
-                       na.value = "white", 
-                       name = "")
+worth_map(mod, 
+          labels = traits, 
+          labels.order = c("Overall", "Taste", "Stretchability", "Colour"))
 
 ## ----llr, message = FALSE, eval = TRUE, echo = TRUE---------------------------
 # by gender
@@ -130,8 +127,10 @@ slice = dat$country
 # and get the unique values
 slice_lvs = unique(slice)
 
-
 trait_plot = list()
+
+# order of varieties from best to worst in the full dataset
+items_lvls = names(sort(rank(coef(mod[[ov]], log = FALSE) * -1)))
 
 for (i in seq_along(slice_lvs)) {
   
@@ -141,17 +140,11 @@ for (i in seq_along(slice_lvs)) {
   })
   
   # plot the worth map
-  trait_plot[[i]] = worth_map(mod_i, 
-            labels = traits) +
-    labs(x = "", 
-         y = "",
-         title = slice_lvs[i]) +
-    scale_fill_distiller(palette = "BrBG", 
-                         direction = 1, 
-                         na.value = "white", 
-                         name = "")
+  trait_plot[[i]] = worth_map(mod_i,
+                              labels = traits,
+                              labels.order = c("Overall", "Taste", "Stretchability", "Colour"),
+                              items.order = items_lvls)
   
-    
 }
 
 # plot the two maps using patchwork

### 2024-12-05 R/kendallTau.R
update-bean-vignette

@@ -13,9 +13,13 @@
 #' @family goodness-of-fit functions
 #' @param x a numeric vector, matrix or data frame
 #' @param y a vector, matrix or data frame with compatible dimensions to \code{x}
+#' @param nboot integer, the size of the bootstrap sample
 #' @param null.rm logical, to remove zeros from \code{x} and \code{y} 
 #' @param average logical, if \code{FALSE} returns the kendall and N-effective for each entry
-#' @param na.omit logical, if \code{TRUE} ignores entries with kendall = NA when computing the average
+#' @param na.omit logical, if \code{TRUE} ignores entries with kendall = NA 
+#'  when computing the average
+#' @param seed integer, the seed for random number generation. If NULL (the default),
+#'  gosset will set the seed randomly
 #' @param ... further arguments affecting the Kendall tau produced. See details 
 #' @return The Kendall correlation coefficient and the Effective N, which 
 #' is the equivalent N needed if all items were compared to all items. 
@@ -272,3 +276,34 @@ kendallTau.paircomp = function(x, y, ...) {
   kendallTau(X, Y, ...)
   
 }
+
+#' @rdname kendallTau
+#' @export
+kendallTau_bootstrap = function(x, y, nboot = 100, seed = NULL, ...) {
+  
+  if(is.null(seed)) {
+    seed = as.integer(runif(1, 1, 100000))
+  }
+  
+  k = 4
+  n = nrow(x)
+  set.seed(seed)
+  folds = sample(rep(1:k, times = ceiling(n / k), length.out = n))
+  
+  result = data.frame()
+  
+  # run over folds
+  for (i in seq_len(k)) {
+    
+    kt = kendallTau(x[folds != i, ], y[folds != i, ], ...)
+    
+    result = rbind(result, kt)
+    
+  }
+  
+  set.seed(seed)
+  result = bayes_boot(result$kendallTau, mean, n1 = nboot, ...)
+  
+}
+
+

### 2024-12-05 R/kendall_permute.R
namespace

@@ -33,7 +33,7 @@
 #'                 split = rep(c("Group1", "Group2", "Group3"), length.out = nrow(breadwheat)), 
 #'                 n.permutations = 100)
 #'                 
-#' @importFrom parallel mclapply
+#' @importFrom parallel mclapply detectCores
 #' @rdname kendallTau
 #' @export
 kendallTau_permute = function(x, y, split, n.permutations = 500) {
@@ -90,7 +90,9 @@ kendallTau_permute = function(x, y, split, n.permutations = 500) {
   }
   
   # Parallelize permutation computation
-  results_list = parallel::mclapply(group_pairs, compute_permuted_diff, mc.cores = detectCores() - 1)
+  results_list = parallel::mclapply(group_pairs, 
+                                    compute_permuted_diff, 
+                                    mc.cores = parallel::detectCores() - 1)
   
   # Combine results into a data frame
   results = do.call(rbind, lapply(seq_along(group_pairs), function(i) {

### 2024-12-05 vignettes/vignette-1-trait-prioritization-and-crop-performance.R
docs

@@ -138,28 +138,3 @@ node_rules(tree)
 
 top_items(tree, top = 3)
 
-## ----node_info2, message=FALSE, eval=TRUE,echo=TRUE---------------------------
-plot(tree, ref = "Amadeus 77")
-
-## ----rel1, message=FALSE, eval=FALSE, echo=TRUE-------------------------------
-#  reliability(tree, ref = "Amadeus 77")
-
-## ----rel2, message=FALSE, eval=TRUE, echo=FALSE-------------------------------
-
-rel = reliability(tree, ref = "Amadeus 77")
-
-rel = rel[rel$reliability >= 0.5, ]
-
-rel = rel[c(1:5)]
-
-rel
-
-
-## ----compare, message=FALSE, echo=TRUE, eval=TRUE, out.width="50%"------------
-Overall = PlackettLuce(R[[baseline]])
-Yield = PlackettLuce(R[[yield]])
-
-compare(Overall, Yield) +
-  labs(x = "Average log(worth)",
-       y = "Difference (Overall appreciation - Yield)")
-

### 2025-10-01 R/rank_numeric.R
add-rank_tricot2()

@@ -44,12 +44,23 @@
 #' 
 #' @importFrom tidyr gather spread
 #' @importFrom PlackettLuce as.rankings group
+#' @importFrom lifecycle deprecate_soft 
 #' @export
 rank_numeric = function(data, items, input, 
                         id = NULL, 
                         group = FALSE, 
                         ascending = FALSE, 
                         ...) {
+  
+  lifecycle::deprecate_soft(
+    when = "1.5.1",         
+    what  = "gosset::rank_numeric()",
+    with  = "gosset::rank_tricot2()",
+    details = paste(
+      "rank_numeric() is deprecated for tricot data.",
+      "Use rank_tricot2(), which optmizes tricot block/item order and sparse inputs."
+    )
+  )
   
   
   if (class(data)[1] != "data.frame") {

### 2025-10-01 R/rank_tricot2.R

@@ -1,0 +1,81 @@
+#' Build Plackett-Luce rankings from tricot ranking dataset 
+#'
+#' Create an object of class "rankings" from tricot ranking data. 
+#'
+#' @author Kauê de Sousa and Jacob van Etten, with ideas from Heather Turner
+#' @family rank functions
+#' @param id an index of \code{data} indicating the id column
+#' @inheritParams rank_tricot
+#' @param ... additional arguments passed to methods. See details
+#' @return a PlackettLuce "rankings" object, which is a matrix of dense rankings 
+#' @seealso \code{\link[PlackettLuce]{rankings}}, \code{\link{nicabean}}
+#' @references 
+#' 
+#' van Etten J., et al. (2016). Experimental Agriculture, 55(S1), 275–296.
+#' \doi{https://doi.org/10.1017/S0014479716000739}
+#' 
+#' @examples
+#' 
+#' library("gosset")
+#' 
+#' R = nicabean$trial
+#' 
+#' R = R[R$trait == "Vigor", ]
+#' 
+#' names(R)
+#' 
+#' R = rank_tricot2(R, id = "id", items = "item", input = "rank")
+#' 
+#' @importFrom PlackettLuce as.rankings group
+#' @export
+rank_tricot2 = function(data,
+                        id,
+                        items, 
+                        input, 
+                        ...) {
+  
+  # columns
+  x = data
+  id_col     = id
+  item_col   = items
+  value_col  = input
+  
+  # order-preserving unique (same as your code)
+  ids        = x[[id_col]][!duplicated(x[[id_col]])]
+  items      = x[[item_col]][!duplicated(x[[item_col]])]
+  
+  # map each row to matrix coordinates
+  i = match(x[[id_col]],   ids)
+  j = match(x[[item_col]], items)
+  
+  # basic sanity checks (optional but helpful)
+  bad_rows = which(is.na(i) | is.na(j) | is.na(x[[value_col]]))
+  if (length(bad_rows)) {
+    warning(sprintf("Skipping %d rows with NA in id/item/value.", length(bad_rows)))
+  }
+  
+  # detect duplicates of the same (id, item)
+  dupes = duplicated(cbind(i, j)) | duplicated(cbind(i, j), fromLast = TRUE)
+  if (any(dupes, na.rm = TRUE)) {
+    d = x[dupes, c(id_col, item_col, value_col)]
[…]

### 2025-10-01 vignettes/vignette-1-trait-prioritization-and-crop-performance.R

@@ -23,11 +23,10 @@ for (i in seq_along(traits)) {
 
   dat_i = subset(dat, dat$trait == traits[i])
 
-  R[[i]] = rank_numeric(data = dat_i,
-                         items = "item",
-                         input = "rank",
-                         id = "id",
-                         ascending = TRUE)
+  R[[i]] = rank_tricot2(data = dat_i,
+                        items = "item",
+                        input = "rank",
+                        id = "id")
 }
 
 
